### 2016-02-14 doc/paper.tex
fixed and improved method to compute neighborhoods

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{fix-cm}
%\documentclass[twocolumn]{svjour3}          % twocolumn
\documentclass[smallextended]{svjour3}          % twocolumn

\smartqed  % flush right qed marks, e.g. at end of proof

\usepackage{graphicx} % adjustbox loads it
\usepackage{epstopdf} % loads eps
\usepackage[normalem]{ulem}

% insert here the call for the packages your document requires
\usepackage{geometry}
\usepackage[export]{adjustbox}
\usepackage[labelfont=bf, labelsep=space]{caption}
%\usepackage{subcaption}
\usepackage{subfig}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}

\usepackage{natbib} % enables author year and other citation styles
\usepackage[bookmarks,bookmarksopen,bookmarksdepth=2]{hyperref} % active links
\hypersetup{backref,
	colorlinks=true,
	citecolor=blue,
	linkcolor=blue}

\usepackage{tikz} % graphics,
\usetikzlibrary{fit,positioning} % tikz elements positioning
\usepackage{soul} % annotations

\DeclareMathOperator*{\argmin}{arg\,min}

\newcommand\alberto[1]{\textcolor{red}{#1}}

%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\graphicspath{ {./img/} }
\DeclareMathOperator*{\argmax}{arg\,max}
\journalname{Comput Stat}

\begin{document}
\title{Structural posts neighborhoods to characterize users in online forums}

\author{Alberto Lumbreras \and
        Julien Velcin  \and\\
        Marie Guégan \and
        Bertrand Jouve
}

\institute{Alberto Lumbreras \and Marie Guégan \at
		   Technicolor\\
           975 Avenue des Champs Blancs, \\35576 Cesson-Sévigné,\\ France\\
           \email{alberto.lumbreras@technicolor.com}\\
           \email{marie.guegan@technicolor.com}
           \and
           Julien Velcin \at
           Laboratoire ERIC, Université de Lyon,\\
           5, avenue Pierre Mendès France, 69676 Bron,\\ France\\
           \email{julien.velcin@univ-lyon2.fr}
		  \and
           Bertrand Jouve \at
           Université de Toulouse; UT2; FRAMESPA/IMT; 5 allée Antonio Machado, 31058 Toulouse, cedex 9\\
           CNRS; FRAMESPA; F-31000 Toulouse\\
           CNRS; IMT; F-31000 Toulouse\\ France\\     
           \email{jouve@univ-tlse2.fr}
}

\date{Draft date: \today}
% The correct dates will be entered by the editor

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:introduction}

The popularization of online forums has brought a growing interest on their underlying dynamics. As any other complex system, an the dynamic of online forums can be studied at different levels, from the more macro to the most micro. Macro dynamics are, for instance, the evolution of some global properties such as its diameter, or its distribution degree. Micro dynamics are, for instance, the triadic motifs that represent local phenomena such as transitivity (friends of my friends are also my friends).

An interesting question in online communities is that concerning roles. In sociology, roles are generally seen as the set of expected behaviors that are attached  to a position in the community. Extrapoliting the notion of role, some researchers have looked for roles in online forums. Some others have tried to detect the roles and the users who hold that roles.

Roles can also be studied from the macro or the micro perspective. If studied from the macro, we can analyze the number of users, the percentage of replied post, its centrality in the network, and  so forth.

In this paper, we focus on the anamyse of roles at a micro level, and more specifically at the discussion level. We would like to answer the following question: are there different types of users in terms of the kind of conversation they participate in?

The remaining of the paper is as follows. First we will introduce our representation of online discussions. Then we will explain our concept of structural neighborhood. Then we will apply this to the clustering of users and we will analyze the clusters. Finally, we will attemp to study whether these roles have predictive power.

\section{Discussion trees}
A very natural way to represent a conversation thread is a tree graph where vertex represent posts and edges represent replies from some post to another. The tree is rooted at the post who started the thread. This representation allows to apply the mathematical toolbox of graph theory. Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.

\begin{figure}
	\centering
	\includegraphics[width=0.32\textwidth]{tree1}
	\includegraphics[width=0.32\textwidth]{tree2}
	\includegraphics[width=0.32\textwidth]{tree3}
	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
	\label{fig:trees}
\end{figure}
\section{Posts graph neighborhoods}

Our intuition is that some users like participating in some kind of discussion rather than other. Certainly, an analysis of the textual content will tell us much about a discussion. However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humor, irony, or changing contexts, we turn our attention towards the structure of the discussions. More precisely, we analyze the local graph structure in which a user post in embedded, in the hope that this structure will be meaningful since it also reflects the kind of conversation in that part of the thread.

Formally, these local graphs are known as neighborhoods. Let us consider first the most basic notion of neighborhood in order to understand its limitations. Let $p_i$ be a post in the tree graph $G$. The \textit{neighborhood of $p_i$ at radius $r$} $N_G(p_i,r)$ is the induced subgraph of $G$ consisting of all posts $j \in G$ at distance $d_{ij}\leq r$. In order to analyze the type of discussion where a user participates in, we might count the different neighborhoods in which her posts are embedded. However, the number of possible neighborhoods even with radius 1 is unbounded since, although a post can only have a parent, it can have an infinite number of children (see Figure~ \ref{fig:large_neighboorhood}).

In practice, if we create a matrix $M$ where rows correspond to users, columns correspond to neighborhoods and $M_{ij}$ indicates how many times user $i$ has a post in the neighborhood $j$,  we obtain a very sparse matrix. Moreover, many of the columns correspond to similar type of neighborhoods. Think, for instance, of a post with 1 reply, a post with 10 replies and another post with 15 replies. While the type of conversation that embeds the first posts is different from the other two, it does not make sense to say that the former two are categorically different.

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{large_neighborhood}
	\includegraphics[width=0.4\textwidth]{small_neighborhood}
	\caption{The size of the neighborhood at distance $R$ is unbounded. These two graph represent frequent neighborhoods of a post (red) that replied to the root (white). However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
	\label{fig:large_neighboorhood}
\end{figure}

In order to reduce the space of possible neighborhoods we introduce the notion of \textit{temporal neighborhood}. Let $t_i$ the time where post $p_i$ was added to the thread $G$. Let $\tau_{ij}=|t_i-t_j|$ be the \textit{temporal distance} between two posts $p_i$ and $p_j$. The \textit{temporal neighborhood} $N_{G}^{(t)}(i, r, n)$of spacial radius $r$ and order $n$ of a post $p_i$ is the set of $n$ posts in its spacial neighborhood $N_{G}(i, r)$ that are temporally closest to $i$ and that can reach $i$ trough posts that are also in the temporal neighborhood.

Finally, we give their own colors the post $i$ and to the root, and we consider that two neighborhoods are equivalent if they are color-isomorphic.

With the above definitions in mind, and for a given forum, we can build a matrix $M$ with the counts of neighborhoods for every user.

\section{Application to Reddit forums}

We analyze the Podemos forum\footnote{https://www.reddit.com/r/podemos} in Reddit. It was was first used as a tool of internal democracy. Today, its members use it also to share and discuss about political news.
The Podemos forum contains 995 discussion threads that started on January. These threads contain a total of 12912 posts written by 1218 users.

\subsection{Counting neighorhoods}
In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighborhood structures.

If we set the a spatial radius $r=1$ and an order $n=4$, the temporal neighborhood $N_{G}^{(t)}(i, 1, 4)$ recovers the four adjacent posts that were written temporally closest to a given post. Figure~\ref{fig:neighborhoods_1_4} shows the four detected neighborhoods and their frequency.

In order to detect more complex types of neighborhoods we enlarge the spatial radius to $r=4$. Figure~\ref{fig:neighborhoods_4_4} shows the eight detected neighbors and their frequency. It is interesting to see different structures like chains or stars where the root node migh or might no appear and the ego post is placed at different positions.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{neighborhoods_1_4}
	\caption{Neighborhood counts with $r=1$ and $n=4$}
	\label{fig:neighborhoods_1_4}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{neighborhoods_4_4}
	\caption{Neighborhood counts with $r=4$ and $n=4$}
	\label{fig:neighborhoods_4_4}
\end{figure}

Note that some neighbors are still very similar. For instance, \textit{common answer 3} is very close to \textit{common answer 2} in the sense that they both represent the ego post replying to a root post that has one and two replies respectively. It seems a good idea to merge them both since they seem to represent the same type of discussion.

\subsection{Clustering}

Given the neighborhoods in which each user participates, we will analyze whether there exists different types of users or all users look similar.

We want to make the analyze independent of the number of posts, and for that we normalize each user feature vector so that features indicate the percentage of posts in this kind of neighborhood. Moreover, some neighborhoods are much more common than others due to the nature of the forums. Thus, we normalize and scale the features so that every feature has a global mean 0 and variance 1. User features now represent z-scores, that is, how many standard deviations is this user feature away from the mean.

We use a simple k-means to find the clusters. To decide the number of clusters, we run k-means for k=2,...,25 clusters and look at the Within-Cluster Sum of Squares (Figure~\ref{fig:elbow}) and we chose $k=5$ so that the results are more interpretable.  Figure~\ref{fig:PCA} shows a PCA projection of the users colored by cluster and the distribution of the clusters in every dimension.
\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{elbow}
	\caption{Within-Sum of Squares under k-means}
	\label{fig:elbow}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{PCA}
	\includegraphics[width=0.5\textwidth]{whiskers}
	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
	\label{fig:PCA}
\end{figure}

\subsection{Predictive roles}
One interesting use of roles is to make predictions on users behaviors. Here, we will analyze whether the initial composition of roles in a thread is a good predictor of the final length of the thread.

\section{Conclusions}
\bibliographystyle{spbasic} 
%\bibliographystyle{chicago} 
%\bibliographystyle{plainnat} 
\bibliography{library}
\end{document}

### 2016-02-14 R/clustering.r

@@ -1,0 +1,95 @@
+cluster <- function(df.user_features){
+    wss <- (nrow(df.user_features)-1)*sum(apply(df.user_features,2,var))
+    for (i in 2:25){
+      wss[i] <- sum(kmeans(df.user_features, i)$withinss)
+    }
+    plot(1:25, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")
+    fit <- kmeans(df.user_features, 5)
+    z <- fit$cluster
+    
+    # relabel cluster by size (to keep color consistency between executions)
+    k.order <- order(table(z), decreasing=TRUE)
+    z.temp <- z
+    for(i in 1:max(z)){
+      z.temp[z==k.order[i]] <- i
+    }
+    z <- z.temp
+    
+    return(z)
+}
+
+# Complementary
+#library(mclust)
+#BIC <-  mclustBIC(df.scaled)
+#plot(BIC)
+#fit2 <-Mclust(df.scaled)
+#summary(fit2, parameters=TRUE)
+
+plot.clusters <- function(df.features, clusters, sizes, colors){
+  par(mfrow=c(1,1))
+  pca <- princomp(df.features)
+  
+  # Colored PCA (base graphics)
+  ######################################
+  plot(pca$scores[,1], pca$scores[,2], 
+       col  = colors, 
+       cex  = sizes, 
+       pch  = 19,
+       xlab = "Dimension 1", ylab = "Dimension 2")
+  text(pca$scores[,1], pca$scores[,2], 
+       labels = rownames(df.features), 
+       cex    = 0.7)
+  title("Individual factor map (PCA)")
+  dev.copy(png, paste0('2016-01-15-PCA.png'), width = 800, height = 800)
+  dev.off()
+  
+  # Colored PCA (another library)
+  ################################
+  p1 <- ggbiplot(pca, 
+                 obs.scale = 1, 
+                 var.scale    = 1, 
+                 groups       = factor(z), 
+                 ellipse      = TRUE, 
+                 circle       = TRUE,
+                 labels       = rownames(df.features),
+                 labels.size  = 3.5,
+                 varname.size = 4) 
+  p1 <- p1 + scale_color_manual(values = alpha(cluster.colors,0.75))
+  p1 <- p1 + theme_bw() +
+    theme(text            = element_text(size = 15),
+          aspect.ratio    = 1,
+          legend.position = "none") +
+    ggtitle("Individual factor map (PCA)")
+  print(p1) 
[…]

### 2016-02-14 R/count_motifs.r

@@ -1,0 +1,200 @@
+count_motifs <- function(threads, database='reddit'){
+  con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
+  
+  # assign and id to each user
+  users <- df.users$user
+  users.ids <- new.env()
+  for(i in 1:length(users)){
+    users.ids[[users[i]]] <- i
+  }
+  
+  # Now we can create a list with user entries in order to count.
+  user.motifs <- matrix(0, nrow=length(users), ncol=100)
+  rownames(user.motifs) <- users
+  
+  posts.id <- vector()
+  posts.motifs.id <- vector()
+  
+  # For every thread, recontruct its graph. Then count 
+  # if neighborhood at distance 1 and binary tree, max nodes is 4
+  # if neighborhood at distance 1 and binary tree, max nodes is 10
+  # 4   10   22   30   46   62   93  123...
+  
+  # the maximum egonet will be a binary tree
+  rad <- 4
+  max.neighbors <- 4
+  motifs <- list()
+  plotted.trees <- 1
+  
+  nthreads <- length(threads)
+  for(i in 1:nthreads){
+    cat('\n', i, '/', nthreads)
+    
+    #Extract the egos of every post
+    g <- database.to.graph(threads[i], con, database)
+    gp <- g$gp
+    egos <- make_ego_graph(gp, rad, nodes=V(gp))
+    
+    if (FALSE){
+      if(vcount(gp)>50){
+        la <- layout_with_fr(gp)
+        plot(as.undirected(gp),
+             layout = la, 
+             vertex.label = "",
+             vertex.color = "black",
+             vertex.size = 1.5 + 1.5 * log( graph.strength(gp), 3 ), 
+             edge.width = 1.5)  
+        dev.copy(png, paste0('2015-01-15-tree', plotted.trees, '.png'))
+        dev.off()
+        plotted.trees <- plotted.trees +1
+      }
+    }
+    
+    for(j in 1:length(egos)){
+      # Reduce the ego to a maximum of N neighbors (the N closest in time)
+      # and where the neighborhood is fully connected.
+      # Set a different color (only) to the ego post
+      user.name <- V(gp)[j]$user
+      # TODO: postid <- V(gp)[j]$postid
+      if (user.name == "root") 
+        next
+      user.id <- users.ids[[user.name]]
+      post.id <- V(gp)[j]$name
[…]

### 2016-02-14 R/forest.r

@@ -1,0 +1,106 @@
+library(igraph)
+library(rgexf)
+library(dplyr)
+forest <- function(df.posts){
+  df.edges <- select(df.posts, postid, parent, date)
+  df.vertices <- select(df.posts, postid, user, date)
+  df.edges <- filter(df.edges, parent%in%df.vertices$postid) # only edges with known vertices
+  gforest <- graph.data.frame(df.edges, directed=TRUE, vertices=df.vertices)
+  gforest
+}
+
+
+plot.forest <- function(gforest){
+  #V(gforest)$date <- as.numeric(V(gforest)$date) - min(as.numeric(V(gforest)$date))
+  #E(gforest)$date <- as.numeric(E(gforest)$date) - min(as.numeric(E(gforest)$date))
+  
+  la <- layout_with_lgl(gforest)
+  plot(gforest,
+       layout = la, 
+       vertex.size = 1,
+       vertex.label = NA,
+       edge.width = 0.2, 
+       edge.arrow.size=0.02,
+       asp=9/16,
+       margin=-0.15)
+  
+  # Dynamic graph
+  df.edges$date <- as.numeric(df.edges$date) - as.numeric(min(df.edges$date))
+  df.vertices$date <- as.numeric(df.vertices$date) - as.numeric(min(df.vertices$date))
+  df.nodeDyn <- data.frame(start=df.vertices$date, end=max(df.vertices$date))
+  df.edgeDyn <- data.frame(start=df.edges$date, end=max(df.edges$date))
+  df.nodeDyn <- data.frame(start=df.vertices$date, end=df.vertices$date)
+  df.edgeDyn <- data.frame(start=df.edges$date, end=df.edges$date)
+  write.gexf(select(df.vertices, postid, user), select(df.edges, postid, parent), 
+             nodeDynamic=df.nodeDyn, 
+             edgeDynamic=df.edgeDyn,
+             defaultedgetype='directed',
+             output='podemos_dynamic.gexf')
+}
+
+socialnetwork <- function(df.posts){
+  
+  df.posts$date <- as.numeric(df.posts$date) - min(as.numeric(df.posts$date))
+  df.posts <- filter(df.posts, user!='[deleted]')
+  
+  # Use this to make every root as a different user
+  #idx.root <- which(df.posts$user=='root') # give every root its own identity
+  #df.posts[idx.root,]$user <- sapply(1:length(idx.root), function(x) paste0("root.",x))
+  
+  # Edges and vertices in the posts graph
+  ########################################
[…]

### 2016-02-14 R/load_participations.r

@@ -1,0 +1,19 @@
+library(RSQLite)
+source('R/extract_from_db.r')
+
+load_posts <- function(database='reddit', forum='podemos'){
+  con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
+  
+  # Threads in forum
+  thread.ids <- dbGetQuery(con,  paste0("select threadid from threads where forum='", forum, "'"))
+  
+  # Dataframe of users and threads where they participated
+  df.posts <- data.frame()
+  for(i in 1:nrow(thread.ids)){ 
+    #thread.posts <- dbGetQuery(con,  paste0("select user,date from posts where thread='", thread.ids[i,], "'"))
+    thread.posts <- dbGetQuery(con,  paste0("SELECT postid, parent, user, date FROM posts WHERE thread='", thread.ids[i,], "'"))
+    thread.posts$thread <- thread.ids[i,]
+    df.posts <- rbind(df.posts, thread.posts)
+  }
+  return(df.posts)
+}

### 2016-02-14 R/normalize_counts.r

@@ -1,0 +1,17 @@
+normalize_counts <- function(df.user.motifs){
+  # Normalize the matrix with the user counts
+  # Summary  matrix
+  df.user.motifs <- df.user.motifs[,colSums(df.user.motifs)>0]
+  cat("\nMotifs counts:", colSums(df.user.motifs))
+  
+  df.user.motifs <- df.user.motifs[rowSums(df.user.motifs)>0,]
+  # Normalize user activity (make the analysis independent of number of posts)
+  df.user.motifs <- as.data.frame(t(apply(df.user.motifs, 1, function(x) x/sum(x))))
+  cat("\nMotifs % of active users:", colSums(df.user.motifs))
+  
+  # Center and scale data
+  df.user.motifs <- as.data.frame(scale(df.user.motifs))
+  cat("\nMotifs % of active users (z-score):", colSums(df.user.motifs))
+  
+  return(df.user.motifs)
+}

### 2016-02-14 R/plotPCA.r

@@ -1,0 +1,61 @@
+library(gplots)
+library(ggbiplot)
+
+plot.clusters <- function(df.features, clusters, sizes, colors){
+    par(mfrow=c(1,1))
+    pca <- princomp(df.features)
+    
+    # Colored PCA (base graphics)
+    ######################################
+    plot(pca$scores[,1], pca$scores[,2], 
+         col  = colors, 
+         cex  = sizes, 
+         pch  = 19,
+         xlab = "Dimension 1", ylab = "Dimension 2")
+    text(pca$scores[,1], pca$scores[,2], 
+         labels = rownames(df.features), 
+         cex    = 0.7)
+    title("Individual factor map (PCA)")
+    dev.copy(png, paste0('2016-01-15-PCA.png'), width = 800, height = 800)
+    dev.off()
+    
+    # Colored PCA (another library)
+    ################################
+    p1 <- ggbiplot(pca, 
+                   obs.scale = 1, 
+                   var.scale    = 1, 
+                   groups       = factor(z), 
+                   ellipse      = TRUE, 
+                   circle       = TRUE,
+                   labels       = rownames(df.features),
+                   labels.size  = 3.5,
+                   varname.size = 4) 
+    p1 <- p1 + scale_color_manual(values = alpha(cluster.colors,0.75))
+    p1 <- p1 + theme_bw() +
+      theme(text            = element_text(size = 15),
+            aspect.ratio    = 1,
+            legend.position = "none") +
+      ggtitle("Individual factor map (PCA)")
+    print(p1) 
+    dev.copy(png, paste0('2016-01-15-PCA2.png'), width=800, height=800)
+    dev.off()
+    
+    # Boxplots
+    #########################
+    # Clusters profile
+    points <- df.features
+    points$cluster <- factor(z)
+    points <- melt(points, id='cluster')
+    p <- ggplot(points, aes(x=variable, y=value)) + 
+      scale_fill_manual(values = cluster.colors) + 
+      geom_boxplot(aes(fill = cluster), position = position_dodge(width = 0.75))
+    p <- p + theme_bw() +
+      theme(text = element_text(size = 15),
+            legend.key = element_blank()) +
+      ggtitle("Clusters means and variances")
+    print(p)
+    dev.copy(png, paste0('2016-01-15-whiskers.png'), width=800)
+    dev.off()
+    
+}
+

### 2016-02-14 R/plot_thread_cluster.r

@@ -1,0 +1,73 @@
+plot.whiskers_cluster_length <- function(df.posts, df.users, df.threads){
+    
+    # Prepare the data
+    df.cluster_length <- merge(merge(df.posts, df.users), df.threads) %>% 
+      mutate(cluster=factor(cluster)) %>%
+      select(cluster, length)
+    
+    # Plot
+    p <- ggplot(df.cluster_length, aes(x=cluster, y=length)) + 
+      scale_fill_manual(values = c("white", cluster.colors)) + # no role -> white color
+      geom_boxplot(aes(fill=cluster)) +
+      theme_bw() +
+      theme(text        = element_text(size = 15),
+            legend.key  = element_blank()) +
+      ggtitle("Participations vs threads length")
+    print(p)
+    
+    # Save to file
+    dev.copy(png, paste0('2016-01-15-whiskers_roles_vs_length.png'), width=800)
+    dev.off()
+}
+
+plot.by_thread_cluster <- function(df.posts, df.users){
+  
+  df.threads <- plyr::count(df.posts, "thread")
+  names(df.threads)[2] <- "length"
+  
+  df.participations <- merge(merge(df.posts, df.users), df.threads)
+  df.participations$cluster <- factor(df.participations$cluster)
+  
+  by_thread_cluster <- acast(df.participations, thread~cluster)
+  by_thread_cluster.perc <- as.data.frame(t(apply(by_thread_cluster, 1, function(x) x/sum(x))))
+  
+  
+  # Add length and sort by length
+  df.lengths <- plyr::count(df.participations, "thread")
+  names(df.lengths) <- c('thread', 'length')
+  by_thread_cluster.perc$thread <-rownames(by_thread_cluster.perc)
+  by_thread_cluster.perc <- merge(by_thread_cluster.perc, df.lengths) # add lengths
+  by_thread_cluster.perc <- by_thread_cluster.perc[order(by_thread_cluster.perc$length),] # sort by length
+  by_thread_cluster.perc <- subset(by_thread_cluster.perc, select=-c(thread)) # drop again thread column
+  by_thread_cluster.norm <- by_thread_cluster.perc
+  by_thread_cluster.norm$length <-  by_thread_cluster.norm$length/max(by_thread_cluster.perc$length)
+  
+  # Plot
+  heatmap(as.matrix(by_thread_cluster.norm), 
+          Colv=NA, Rowv=NA, labRow=NA, ylab="Threads", xlab="Roles",
[…]

### 2016-02-14 R/pred_regression.r

@@ -1,0 +1,13 @@
+pred_regression <- function(df.thread_cluster){
+  X <- t(as.matrix(df.thread_cluster[,1:6]))
+  X <- rbind(rep(1, dim(X)[2]),X)
+  y <- as.matrix(df.thread_cluster[,7])
+  
+  par(mfrow=c(1,1))
+  alphaI <- diag(dim(X)[1])*0.01
+  b_mle <- (solve(X%*%t(X)+alphaI)%*%X)%*%y
+  preds <- t(X)%*%b_mle
+  plot(preds, ylim=c(0,max(y)))
+  lines(y, col='red')
+  title(paste("Predictions with linear regression. SSE:", (1/length(y))*sqrt(sum((preds-y)^2))))
+}

### 2016-02-14 pipeline_clustering.r

@@ -1,0 +1,111 @@
+# Extract list of triads where a user participates.
+#
+# author: Alberto Lumbreras
+#
+#https://www.reddit.com/r/europe/CasualConversation
+#https://www.reddit.com/r/europe/
+#https://www.reddit.com/r/datascience
+#https://www.reddit.com/r/science/
+#https://www.reddit.com/r/france
+#https://www.reddit.com/r/catalunya
+#https://www.reddit.com/r/es
+
+library(ggplot2)
+library(ggbiplot)
+library(gplots)
+
+library(reshape2)
+library(igraph)
+
+#library(RSQLite)
+#library(dplyr)
+#library(GGally)
+
+source('R/load_participations.r')
+source('R/count_motifs.r')
+source('R/normalize_counts.r')
+source('R/clustering.r')
+
+###################################################
+# Load data
+###################################################
+df.posts <- load_posts(database='reddit', forum='podemos')
+df.threads <- plyr::count(df.posts, "thread")
+df.users <- plyr::count(df.posts, 'user')
+names(df.threads)[2] <- "length"
+names(df.users)[2] <- "posts"
+
+######################################################
+# Count motifs in which each user appears
+######################################################
+# deprecated
+#res <- count_motifs(df.threads$thread, database='reddit')
+#df.user.motifs <- as.data.frame(res$user.motifs)
+#df.post.motif <- res$posts.motifs
+#motifs <- res$motifs
+
+# more elegant
+res <-count_motifs_by_post(df.threads$thread, database='reddit')
+df.post.motif  <- res$posts.motifs
+df.posts <- merge(df.posts, df.post.motif, all.x=TRUE, sort=FALSE)
+df.user.motifs <- acast(df.posts, user~motif)
+df.user.motifs <- df.user.motifs[rownames(df.user.motifs)!='root',] # remove root
+motifs <- res$motifs
+
+# sort by frequency
+idx <- order(colSums(df.user.motifs), decreasing = TRUE)
+df.user.motifs <- df.user.motifs[,idx]
+motifs <- motifs[idx]
+
+# Put names to the features
+#colnames(df.user.motifs) <- c("spot", "lonely answer",  "chain init answer 1", "chain init answer 2", 
+#                              "common answer 1", "common answer 2", "common answer 3", "star init answer")
+
+# Plot motifs with names
+mypalette <- c("black", "red", "white")
+par(mfrow=c(3,3))
+for(i in 1:length(motifs)){
[…]

### 2016-02-14 pipeline_coparticipations.r

@@ -1,0 +1,84 @@
+#########################
+# Role co-participation
+#########################
+m.thread_cluster <- merge(df.posts, df.users, by='user') %>% 
+  acast(thread~cluster) %>%
+  apply(1, function(x) x/sum(x)) %>%
+  t
+
+role.cooccur <- t(m.thread_cluster)%*%m.thread_cluster
+image(role.cooccur)
+
+##############################
+# User co-participations
+##############################
+# TODO: It might be interesting to do blockmodeling over the cooccur matrix
+user_thread_matrix <- group_by(df.posts, user, thread)  %>% 
+  tally() %>%
+  acast(user ~ thread, fun.aggregate = length)
+user_thread_matrix <- ifelse(user_thread_matrix>0, 1,0)
+cooccur <- user_thread_matrix%*%t(user_thread_matrix)
+cooccur <- ifelse(cooccur>0, 1,0)
+image(cooccur)
+title(paste("density:", mean(cooccur)*100))
+
+# Co-participations of active users
+user_thread_matrix <- user_thread_matrix[df.users$cluster!=0,]
+cooccur.active <- user_thread_matrix%*%t(user_thread_matrix)
+cooccur.active <- ifelse(cooccur.active>0, 1,0)
+image(cooccur.active)
+title(paste("density:", mean(cooccur.active)*100))
+
+heatmap(cooccur.active, 
+        Colv=NA, Rowv=NA, labRow=NA)
+
+# Graph
+#############################################################
+par(mfrow=c(1,1))
+g <- graph.adjacency(cooccur, mode = "undirected", diag = FALSE)
+V(g)$cluster <- df.users$cluster # add clusters to colors users
+g <- delete.vertices(g, "root") # remove root since this is an artificial user
+
+# Remove micro-components (probably threads that have just started at the end of January)
+memberships <- components(g)$membership
+g <- induced_subgraph(g, memberships==1)
+# remove non-active users (those that have no role)
+g <- induced_subgraph(g, V(g)$cluster!=0)
+
+# Save to file for Gephi
+fname <- paste0('forum', '.edges')
+write_graph(g, fname, format = c("edgelist"))
+
+# Plot graph, only for active users
+degrees <- degree(g)
+q_degree <- quantile(degrees, 0.5)
+V(g)$label <- V(g)$name
+#V(g)$label[degree(g)<800] = ""
+la <- layout_with_fr(g)
[…]

### 2016-02-14 pipeline_prediction.r

@@ -1,0 +1,129 @@
+library(dplyr)
+source('R/plot_thread_cluster.r')
+
+# Some plots and setting up the matrix to make predictions
+#############################################################
+# Plot individual relation between participation by cluster vs length
+plot.whiskers_cluster_length(df.posts, df.users, df.threads)
+
+# Plot cluster census by thread
+plot.by_thread_cluster(df.posts, df.users)
+
+# Create a dataset where features are the % of users within every role in the first posts
+# of a thread, and output is the length of a thread
+df.thread_cluster <- by_thread_cluster_nfirst(df.posts, nfirst=10) %>%
+                     merge(df.users, by=c('user')) %>%
+                     acast(thread~cluster) %>%
+                     apply(1, function(x) x/sum(x)) %>%
+                     t %>%
+                     as.data.frame %>%
+                     merge(df.threads, by.x="row.names", by.y="thread") %>% # add length
+                     arrange(length) %>% # sort by length 
+                     select(-Row.names) #drop thread names
+
+# Plot it with length column normalized to stay in the same color range than the rest  
+df.thread_cluster %>% 
+  mutate(length=length/max(length)) %>%
+  as.matrix %>%
+  heatmap(Colv=NA, Rowv=NA, labRow=NA, ylab="Threads", xlab="Roles", 
+          main = "Participations by role (10 first posts)")  
+  
+dev.copy(png, paste0('2016-01-15-thread_role_composition_nfirst.png'), width=600)
+dev.off()
+
+
+##########################################
+# Length prediction
+###########################################
+source('R/pred_regression.r')
+# pred_regression(df.thread_cluster)
+
+# Binary lengths 
+df.thread_cluster <- mutate(df.thread_cluster, length =factor(length > quantile(length,0.5)))
+
+
+# SVM
+######################
+library('e1071')
+if(FALSE){
+  # Feature selection
+  features <- 1:(ncol(df.thread_cluster)-1)
+  idx <- sample(length(y)) 
+  cut <- floor(length(idx)*0.7)
+  idx.train <- idx[1:cut]
+  idx.test <- idx[cut:length(idx)]
+  y.train <- df.thread_cluster$length[idx.train]
+  y.test <- df.thread_cluster$length[idx.test]
[…]

### 2016-02-14 sandbox/SO_isomorphism_bug.r

@@ -1,0 +1,35 @@
+"""
+<?xml version="1.0" encoding="UTF-8"?>
+  <graphml xmlns="http://graphml.graphdrawing.org/xmlns"
+xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
+xsi:schemaLocation="http://graphml.graphdrawing.org/xmlns
+         http://graphml.graphdrawing.org/xmlns/1.0/graphml.xsd">
+  <!-- Created by igraph -->
+  <key id="v_color" for="node" attr.name="color" attr.type="double"/>
+  <key id="v_id" for="node" attr.name="id" attr.type="string"/>
+  <graph id="G" edgedefault="directed">
+  <node id="n0">
+  <data key="v_color">3</data>
+  <data key="v_id">n0</data>
+  </node>
+  <node id="n1">
+  <data key="v_color">1</data>
+  <data key="v_id">n1</data>
+  </node>
+  <node id="n2">
+  <data key="v_color">1</data>
+  <data key="v_id">n2</data>
+  </node>
+  <node id="n3">
+  <data key="v_color">1</data>
+  <data key="v_id">n3</data>
+  </node>
+  <edge source="n1" target="n0">
+  </edge>
+  <edge source="n2" target="n0">
+  </edge>
+  <edge source="n3" target="n0">
+  </edge>
+  </graph>
+  </graphml>
+"""

### 2016-02-14 sandbox/isomorphs.r

@@ -1,0 +1,105 @@
+g1 <- graph.empty(1)
+g2 <- make_graph(c(2,1, 3,1, 4,1))
+V(g1)$color <- 2
+V(g2)$color <- c(3,1,1,1)
+is_isomorphic_to(g1,g2, method='vf2')
+
+mypalette <- c("black", "red", "white")
+
+par(mfrow=c(1,2))
+la <- layout_with_fr(g1)
+plot(g1,
+     layout = la, 
+     vertex.color = mypalette[V(g1)$color],
+     vertex.label = V(g1)$color,
+     vertex.size = 5, 
+     edge.width = 1.5)
+title("g1")
+
+la <- layout_with_fr(g2)
+plot(g2,
+     layout = la, 
+     vertex.color = mypalette[V(g2)$color],
+     vertex.label = V(g2)$color,
+     vertex.size = 5, 
+     edge.width = 1.5)
+title("g2")
+
+g1 <- make_ring(3, directed=TRUE)
+g2 <- make_ring(3, directed=TRUE)
+V(g1)$color <- c(1,2,3)
+V(g2)$color <- c(1,2,3)
+isomorphic(g1,g2)
+isomorphic(g1,g2, method='vf2')
+
+
+
+mypalette <- c("black", "red", "white")
+
+par(mfrow=c(1,2))
+la <- layout_with_fr(g1)
+plot(g1,
+     layout = la, 
+     vertex.color = mypalette[V(g1)$color],
+     vertex.label = V(g1)$color,
+     vertex.size = 5, 
+     edge.width = 1.5)
+title("g1")
+
+la <- layout_with_fr(g2)
+plot(g2,
+     layout = la, 
+     vertex.color = mypalette[V(g2)$color],
+     vertex.label = V(g2)$color,
+     vertex.size = 5, 
+     edge.width = 1.5)
+title("g2")
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+#write.graph(g1, "g1", format='graphml')
+#write.graph(g2, "g2", format='graphml')
+#write.graph(eg, "eg", format='graphml')
+#write.graph(gmotif, "gmotif", format='graphml')
+
+#g1 <- read.graph("g1",  format='graphml')
+#g2 <- read.graph("g2",  format='graphml')
+#eg <- read.graph("eg",  format='graphml')
+#gmotif <- read.graph("gmotif",  format='graphml')
+#is_isomorphic_to(eg, gmotif, method='vf2')
+
[…]

### 2016-02-14 sandbox/order_by_group.r

@@ -1,0 +1,21 @@
+# For each Species, change Sepal.Length by their order in the group
+library(dplyr)
+library(plyr)
+data(iris)
+by_species <- iris %>% arrange(Species, Sepal.Length) %>% group_by(Species) %>% mutate(rank=row_number())
+# Error
+
+detach("package:plyr", unload=TRUE)
+by_species <- iris %>% arrange(Species, Sepal.Length) %>% group_by(Species) %>% mutate(rank=row_number())
+by_species %>% filter(rank <= 3)
+
+
+# With Base R
+iris <- iris[with(iris, order(Species, Sepal.Length)), ]
+
+# RUN A ROW COUNT FOR RANK BY SPECIES GROUP
+iris$rank <- sapply(1:nrow(iris), 
+                    function(i) sum(iris[1:i, c('Species')]==iris$Species[i]))
+
+# FILTER DATA FRAME BY TOP 3
+iris <- iris[iris$rank <= 3,]

### 2016-02-14 user_triads.r

@@ -1,352 +1,0 @@
-# Extract list of triads where a user participates.
-#
-# author: Alberto Lumbreras
-#
-#https://www.reddit.com/r/europe/CasualConversation
-#https://www.reddit.com/r/europe/
-#https://www.reddit.com/r/datascience
-#https://www.reddit.com/r/science/
-#https://www.reddit.com/r/france
-#https://www.reddit.com/r/catalunya
-#https://www.reddit.com/r/es
-####################################################
-#  TODO: I can use a color for the ego user so that its position in the egonet is taken into account by the iso algorithm
-library(RSQLite)
-library(dplyr)
-library(ggplot2)
-library(gplots)
-library(igraph)
-library(GGally)
-library(ggbiplot)
-library(reshape2)
-
-source('./R/extract_from_db.r')
-
-# Load data
-dataset <- "reddit"
-forum <- 'podemos'
-con <- dbConnect(dbDriver("SQLite"), 
-                 dbname = paste0("./data/", dataset, ".db"))
-
-# Threads in forum
-thread.ids <- dbGetQuery(con,  paste0("select threadid from threads where forum='", forum, "'"))
-
-# Dataframe of users and threads where they participated
-df.user_thread <- data.frame()
-for(i in 1:nrow(thread.ids)){ 
-  thread.users <- dbGetQuery(con,  paste0("select user from posts where thread='", thread.ids[i,], "'"))
-  thread.users$thread <- thread.ids[i,]
-  #thread.users$length <- nrow(thread.users)
-  df.user_thread <- rbind(df.user_thread, thread.users)
-}
-
-df.threads <- count(df.user_thread, "thread")
-names(df.threads)[2] <- "length"
-df.users <- count(df.user_thread, 'user')
-names(df.users)[2] <- "posts"
-
-# assign and id to each user
-users <- df.users$user
-users.ids <- new.env()
-for(i in 1:length(users)){
-  users.ids[[users[i]]] <- i
-}
-
-
-######################################################
-# Count motifs in which each user appears
-######################################################
-
-# Now we can create a list with user entries in order to count.
-users.motifs <- matrix(0, nrow=length(users), ncol=100)
-rownames(users.motifs) <- users
-
-# For every thread, recontruct its graph. Then count 
-# if neighborhood at distance 1 and binary tree, max nodes is 4
[…]

### 2016-02-18 doc/presentation_MIT_Paris_Fev_2016.tex
backup after Paris

\documentclass{beamer}

\usepackage{default}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}

\DeclareGraphicsExtensions{{.pdf},{.png},{.jpg}}
\graphicspath{ {./img_presentation/} }

\hypersetup{colorlinks,urlcolor=blue}
\AtBeginSection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection]
		\end{NoHyper}
	\end{frame}
}
\begin{document}
\addtobeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\title{(temporal) Motifs in discussions (trees)}
\author{Alberto Lumbreras}
\date{February 17, 2016}
\maketitle

\begin{frame}\frametitle{Overview} 
\begin{NoHyper}
\tableofcontents
\end{NoHyper}
\end{frame}

\section{Introduction}
\subsection{Reddit dataset}

\begin{frame}{Reddit dataset}{A forum of forums}
	Download from:
	\href{http://couch.whatbox.ca:36975/reddit/comments/monthly/}{http://couch.whatbox.ca:36975/reddit/comments/monthly/}

	\vfill
	Extract forum of interest:\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}science}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}france}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}sociology}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}complexsystems}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}podemos}}\\
	...
\end{frame}

\begin{frame}{Graph representations}{Graph of user interactions (a social network)}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{sna}	
	\end{figure}
\end{frame}

\subsection{Graph representations}
\begin{frame}{Graph representations}{Trees of posts}
\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{forest}	
\end{figure}
\end{frame}

\subsection{Dynamics of conversations}
\begin{frame}{Dynamics of conversations}{Triads are not enough}
% Dibujar un arbol a la derecha
Triads in \textbf{trees of posts}:
	\begin{itemize}
		\item Only 3 possible triads (dyad, chain and star)
	\end{itemize}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{triads}
	\end{figure}
Triads in graph of \textbf{user interactions}:
	\begin{itemize}
		\item Need of time in edges.
	\end{itemize}
\end{frame}

\begin{frame}{Temporal neighborhoods in trees}{Definition}
\begin{itemize}
	\item $N_{G}(i, d)$: neighborhood of post $i$ at distance $d$.
	\item $N_{G}(i, d, n)$: keep only the $n$ neighbors in $N_{G}(i, d)$ that are \textit{temporally} closest to post $i$ (computed as $|t-t_i|$)
	\item Keep only those posts in $N_{G}(i, d, n)$ that have a path to $i$. 
\end{itemize}

 $N_{G}(i, d, n)$ is the \textbf{temporal neighborhood} of post $i$ with distance $d$ and order $n$. \footnote{sexier names accepted}

	\begin{figure}
		\centering
		\includegraphics[width=0.5\textwidth]{large_neighborhood}
	\end{figure}

\end{frame}

\section{Neighborhood census}
\begin{frame}{Neighborhood census}{A real example}
Distance $d=4$ and order $n=4$. 41 discussion patterns:

\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{motifs1}	
\end{figure}

\end{frame}

\begin{frame}{Neighborhood census}{Cyclic dynamics}

\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{motifs_by_hour}	
\end{figure}
\end{frame}

\begin{frame}{Neighborhood census}{Proportions}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{census_proportions}	
	\end{figure}
\end{frame}

\begin{frame}{Neighborhood census}{Census and thread growth}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{census_length}	
	\end{figure}
\end{frame}

\begin{frame}{Neighborhood census}{Predictions}
	\textbf{Q}: Can we predict whether a thread will succeed based on its initial structure (neighborhoods)?\\
	\textbf{A}: ...\textit{the answer in a few hours}
\end{frame}
\section{You are the way you (structurally) talk?}
\begin{frame}{Neighborhood-based clustering}{Overview}
\begin{itemize}
	\item Create a user $\times$ neighborhood matrix of counts.
	\item Z-normalize (users characterized by their deviation from the mean)
	\item Cluster!
\end{itemize}
\end{frame}

\begin{frame}{Neighborhood-based clustering}{Uncolored neighborhoods}
	\begin{figure}
		\centering
		\includegraphics[width=0.75\textwidth]{PCA_nocolor}
	\end{figure}	
\end{frame}

\begin{frame}{Neighborhood-based clustering}{Uncolored neighborhoods}
	$k$-means suggests 5 groups:
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{PCA}%
	\end{figure}	
\end{frame}

\begin{frame}{Neighborhood-based clustering }
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{whiskers}	
	\end{figure}
\end{frame}

\section{Conclusions}
\begin{frame}{Conclusions}
	\begin{itemize}
			\item Temporal neighborhoods richer than triads to analyze the structure of conversations.
			\item Users can be characterized in terms of what type of neighborhood they participate in (or they trigger).
	\end{itemize}
	Future work:
	\begin{itemize}
		\item Do users jump from cluster to cluster (paths of roles)
		\item Are initial census predictive of the success of a discussion?.
	\end{itemize}

\end{frame}
\end{document}

### 2016-02-18 R/clustering.r

@@ -67,7 +67,7 @@ plot.clusters <- function(df.features, clusters, sizes, colors){
   # Boxplots
   #########################
   # Clusters profile
-  points <- df.features
+  points <- as.data.frame(df.features)
   points$cluster <- factor(z)
   points <- melt(points, id='cluster')
   p <- ggplot(points, aes(x=variable, y=value)) + 

### 2016-02-18 R/count_motifs.r

@@ -1,110 +1,3 @@
-count_motifs <- function(threads, database='reddit'){
-  con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
-  
-  # assign and id to each user
-  users <- df.users$user
-  users.ids <- new.env()
-  for(i in 1:length(users)){
-    users.ids[[users[i]]] <- i
-  }
-  
-  # Now we can create a list with user entries in order to count.
-  user.motifs <- matrix(0, nrow=length(users), ncol=100)
-  rownames(user.motifs) <- users
-  
-  posts.id <- vector()
-  posts.motifs.id <- vector()
-  
-  # For every thread, recontruct its graph. Then count 
-  # if neighborhood at distance 1 and binary tree, max nodes is 4
-  # if neighborhood at distance 1 and binary tree, max nodes is 10
-  # 4   10   22   30   46   62   93  123...
-  
-  # the maximum egonet will be a binary tree
-  rad <- 4
-  max.neighbors <- 4
-  motifs <- list()
-  plotted.trees <- 1
-  
-  nthreads <- length(threads)
-  for(i in 1:nthreads){
-    cat('\n', i, '/', nthreads)
-    
-    #Extract the egos of every post
-    g <- database.to.graph(threads[i], con, database)
-    gp <- g$gp
-    egos <- make_ego_graph(gp, rad, nodes=V(gp))
-    
-    if (FALSE){
-      if(vcount(gp)>50){
-        la <- layout_with_fr(gp)
-        plot(as.undirected(gp),
-             layout = la, 
-             vertex.label = "",
-             vertex.color = "black",
-             vertex.size = 1.5 + 1.5 * log( graph.strength(gp), 3 ), 
-             edge.width = 1.5)  
-        dev.copy(png, paste0('2015-01-15-tree', plotted.trees, '.png'))
-        dev.off()
-        plotted.trees <- plotted.trees +1
-      }
-    }
-    
-    for(j in 1:length(egos)){
-      # Reduce the ego to a maximum of N neighbors (the N closest in time)
-      # and where the neighborhood is fully connected.
-      # Set a different color (only) to the ego post
-      user.name <- V(gp)[j]$user
-      # TODO: postid <- V(gp)[j]$postid
-      if (user.name == "root") 
-        next
-      user.id <- users.ids[[user.name]]
-      post.id <- V(gp)[j]$name
[…]

### 2016-02-18 R/forest.r

@@ -11,9 +11,9 @@ forest <- function(df.posts){
 
 
 plot.forest <- function(gforest){
-  #V(gforest)$date <- as.numeric(V(gforest)$date) - min(as.numeric(V(gforest)$date))
-  #E(gforest)$date <- as.numeric(E(gforest)$date) - min(as.numeric(E(gforest)$date))
-  
+  # Plot the threads of a forum
+  # Arguments:
+  #   gforest: and igraph object
   la <- layout_with_lgl(gforest)
   plot(gforest,
        layout = la, 
@@ -23,6 +23,17 @@ plot.forest <- function(gforest){
        edge.arrow.size=0.02,
        asp=9/16,
        margin=-0.15)
+}
+
+plot.forest.gephi <- function(df.posts){
+  # Creates a gexf file that represents the graph of threes (forest) of a set of posts
+  # extracted from a forum
+  # Arguments:
+  #   df.posts: dataframe with posts in rows and at least columns postid, parent, and date
+  
+  df.edges <- select(df.posts, postid, parent, date)
+  df.vertices <- select(df.posts, postid, user, date)
+  df.edges <- filter(df.edges, parent%in%df.vertices$postid) # only edges with known vertices
   
   # Dynamic graph
   df.edges$date <- as.numeric(df.edges$date) - as.numeric(min(df.edges$date))
@@ -38,21 +49,21 @@ plot.forest <- function(gforest){
              output='podemos_dynamic.gexf')
 }
 
+
 socialnetwork <- function(df.posts){
   
   df.posts$date <- as.numeric(df.posts$date) - min(as.numeric(df.posts$date))
   df.posts <- filter(df.posts, user!='[deleted]')
   
   # Use this to make every root as a different user
-  #idx.root <- which(df.posts$user=='root') # give every root its own identity
-  #df.posts[idx.root,]$user <- sapply(1:length(idx.root), function(x) paste0("root.",x))
+  # idx.root <- which(df.posts$user=='root') # give every root its own identity
+  # df.posts[idx.root,]$user <- sapply(1:length(idx.root), function(x) paste0("root.",x))
   
   # Edges and vertices in the posts graph
   ########################################
   df.edges <- select(df.posts, postid, parent, date)
   df.vertices <- select(df.posts, postid, user, date)
   df.edges <- filter(df.edges, parent%in%df.vertices$postid) # only edges with known vertices
-  
 
   # Edges and vertices in the users graph
[…]

### 2016-02-18 R/normalize_counts.r

@@ -1,17 +1,21 @@
-normalize_counts <- function(df.user.motifs){
+normalize_counts <- function(user.motifs){
   # Normalize the matrix with the user counts
   # Summary  matrix
-  df.user.motifs <- df.user.motifs[,colSums(df.user.motifs)>0]
-  cat("\nMotifs counts:", colSums(df.user.motifs))
+  features <- user.motifs
+  cat("\nMotifs counts:", colSums(features))
   
-  df.user.motifs <- df.user.motifs[rowSums(df.user.motifs)>0,]
   # Normalize user activity (make the analysis independent of number of posts)
-  df.user.motifs <- as.data.frame(t(apply(df.user.motifs, 1, function(x) x/sum(x))))
-  cat("\nMotifs % of active users:", colSums(df.user.motifs))
+  features <- t(apply(features, 1, function(x) x/sum(x)))
+  
+  #as.data.frame(t(apply(user.motifs, 1, function(x) x/sum(x))))
+  cat("\nMotifs % of active users:", colSums(features))
   
   # Center and scale data
-  df.user.motifs <- as.data.frame(scale(df.user.motifs))
-  cat("\nMotifs % of active users (z-score):", colSums(df.user.motifs))
+  features <- scale(features)
+  cat("\nMotifs % of active users (z-score):", colSums(features))
   
-  return(df.user.motifs)
+  # Some NA because no active user has these neighborhoods. 
+  #user.motifs[is.na(user.motifs)] <- 0
+  #features[,is.na(colSums(features))] <- 0
+  return(features)
 }

### 2016-02-18 R/plotPCA.r

@@ -1,61 +1,0 @@
-library(gplots)
-library(ggbiplot)
-
-plot.clusters <- function(df.features, clusters, sizes, colors){
-    par(mfrow=c(1,1))
-    pca <- princomp(df.features)
-    
-    # Colored PCA (base graphics)
-    ######################################
-    plot(pca$scores[,1], pca$scores[,2], 
-         col  = colors, 
-         cex  = sizes, 
-         pch  = 19,
-         xlab = "Dimension 1", ylab = "Dimension 2")
-    text(pca$scores[,1], pca$scores[,2], 
-         labels = rownames(df.features), 
-         cex    = 0.7)
-    title("Individual factor map (PCA)")
-    dev.copy(png, paste0('2016-01-15-PCA.png'), width = 800, height = 800)
-    dev.off()
-    
-    # Colored PCA (another library)
-    ################################
-    p1 <- ggbiplot(pca, 
-                   obs.scale = 1, 
-                   var.scale    = 1, 
-                   groups       = factor(z), 
-                   ellipse      = TRUE, 
-                   circle       = TRUE,
-                   labels       = rownames(df.features),
-                   labels.size  = 3.5,
-                   varname.size = 4) 
-    p1 <- p1 + scale_color_manual(values = alpha(cluster.colors,0.75))
-    p1 <- p1 + theme_bw() +
-      theme(text            = element_text(size = 15),
-            aspect.ratio    = 1,
-            legend.position = "none") +
-      ggtitle("Individual factor map (PCA)")
-    print(p1) 
-    dev.copy(png, paste0('2016-01-15-PCA2.png'), width=800, height=800)
-    dev.off()
-    
-    # Boxplots
-    #########################
-    # Clusters profile
-    points <- df.features
-    points$cluster <- factor(z)
-    points <- melt(points, id='cluster')
-    p <- ggplot(points, aes(x=variable, y=value)) + 
-      scale_fill_manual(values = cluster.colors) + 
-      geom_boxplot(aes(fill = cluster), position = position_dodge(width = 0.75))
-    p <- p + theme_bw() +
-      theme(text = element_text(size = 15),
-            legend.key = element_blank()) +
-      ggtitle("Clusters means and variances")
-    print(p)
-    dev.copy(png, paste0('2016-01-15-whiskers.png'), width=800)
-    dev.off()
-    
-}
-

### 2016-02-18 pipeline_clustering.r

@@ -14,11 +14,11 @@ library(ggplot2)
 library(ggbiplot)
 library(gplots)
 
+library(dplyr)
 library(reshape2)
 library(igraph)
-
+library(ggbiplot)
 #library(RSQLite)
-#library(dplyr)
 #library(GGally)
 
 source('R/load_participations.r')
@@ -26,6 +26,7 @@ source('R/count_motifs.r')
 source('R/normalize_counts.r')
 source('R/clustering.r')
 
+MIN_POSTS <- 30 # number of post to consider a user as active
 ###################################################
 # Load data
 ###################################################
@@ -35,35 +36,29 @@ df.users <- plyr::count(df.posts, 'user')
 names(df.threads)[2] <- "length"
 names(df.users)[2] <- "posts"
 
-######################################################
-# Count motifs in which each user appears
-######################################################
-# deprecated
-#res <- count_motifs(df.threads$thread, database='reddit')
-#df.user.motifs <- as.data.frame(res$user.motifs)
-#df.post.motif <- res$posts.motifs
-#motifs <- res$motifs
+# Compute neighborhood around every post
+#res <-count_motifs_by_post(df.threads$thread, database='reddit')
 
-# more elegant
-res <-count_motifs_by_post(df.threads$thread, database='reddit')
 df.post.motif  <- res$posts.motifs
-df.posts <- merge(df.posts, df.post.motif, all.x=TRUE, sort=FALSE)
-df.user.motifs <- acast(df.posts, user~motif)
-df.user.motifs <- df.user.motifs[rownames(df.user.motifs)!='root',] # remove root
 motifs <- res$motifs
 
-# sort by frequency
-idx <- order(colSums(df.user.motifs), decreasing = TRUE)
-df.user.motifs <- df.user.motifs[,idx]
+# Sort by frequency (and relabel: 1 for the most frequent and so forth)
+idx <- order(table(df.post.motif$motif), decreasing = TRUE)
+df.post.motif$motif <- match(df.post.motif$motif, idx)
 motifs <- motifs[idx]
 
-# Put names to the features
-#colnames(df.user.motifs) <- c("spot", "lonely answer",  "chain init answer 1", "chain init answer 2", 
-#                              "common answer 1", "common answer 2", "common answer 3", "star init answer")
+# Add motif info to posts dataframe
+df.posts <- merge(df.posts, df.post.motif, all.x=TRUE, sort=FALSE)
 
-# Plot motifs with names
+# Check that df.posts gets the motifs relabeled
[…]

### 2016-02-18 pipeline_neighborhoods_census_evolution.r

@@ -1,0 +1,143 @@
+# Extract the neighborhood census of a forum
+# and plot its evolution
+# author: Alberto Lumbreras
+#################################################
+library(dplyr)
+library(reshape2)
+library(ggplot2)
+source('R/load_participations.r')
+source('R/count_motifs.r')
+select <- dplyr::select # avoid confusion with MASS function
+
+###################################################
+# Load data
+###################################################
+df.posts <- load_posts(database='reddit', forum='podemos')
+df.threads <- plyr::count(df.posts, "thread")
+df.users <- plyr::count(df.posts, 'user')
+names(df.threads)[2] <- "length"
+names(df.users)[2] <- "posts"
+
+######################################################
+# Extract neighborhood around every post
+######################################################
+res <- count_motifs_by_post(df.threads$thread, database='reddit')
+df.post.motif  <- res$posts.motifs
+motifs <- res$motifs
+
+# Sort by frequency
+idx <- order(table(df.post.motif$motif), decreasing = TRUE)
+motifs <- motifs[idx]
+df.post.motif$motif <- match(df.post.motif$motif, idx)
+
+df.posts <- merge(df.posts, df.post.motif, all.x=TRUE, sort=FALSE)
+
+####################################################
+# Plot neighborhoods by frequency
+####################################################
+mypalette <- c("black", "red", "white")
+par(mfrow=c(3,5))
+for(i in 1:length(motifs)){
+  gmotif <- as.undirected(motifs[[i]])
+  la = layout_as_tree(gmotif, mode='out', root=which.min(V(gmotif)$date))
+  plot(gmotif,
+       layout = la,
+       vertex.color=mypalette[V(motifs[[i]])$color],
+       vertex.label = "",
+       edge.arrow.size=0.6)
+  title(paste(i),sub=sum(df.post.motif$motif==i))  
+}
+
+
+################################################
+# Plot cumsum with sliding window
+# Note: lay with both times, real and sequential
+################################################
+df.census <- arrange(df.posts, date) %>%
+             acast(date~motif) %>%
+             as.data.frame
+
+# chronological dates
+df.census$date <- rownames(df.census)
+
+# sequential dates
+df.census$date <- 1:nrow(df.census)
+
+steps <- seq(min(as.numeric(df.census$date)),
+         max(as.numeric(df.census$date)), 
+         by=3600)
+
+df.census.window <- vector()
+for(i in 1:(length(steps)-1)){
[…]

### 2016-02-18 sandbox/time_series.r

@@ -1,0 +1,18 @@
+head(beav1, 100)
+data(beav1)
+beav1$day <- as.factor(beav1$day)
+beav1[beav1$day==346,]$time <- 1:sum(beav1$day==346)
+beav1[beav1$day==347,]$time <- 1:sum(beav1$day==347)
+beav1 <- filter(beav1, time<23)
+ggplot(beav1, aes(x=time, y=temp, group=day, fill=day, color=day)) + stat_summary(fun.y = mean, na.rm=TRUE, geom='line')
+
+ggplot(beav1, aes(x=time, y=temp, group=day, fill=day, color=day)) + 
+  geom_line()
+
+data(beav1)
+beav1$day <- as.factor(beav1$day)
+beav1[beav1$day==346,]$time <- 1:sum(beav1$day==346)
+beav1[beav1$day==347,]$time <- 1:sum(beav1$day==347)
+beav1 <- filter(beav1, time<23)
+ggplot(beav1, aes(x=time, y=temp, group=day, fill=day, color=day)) + 
+  stat_summary(fun.y = mean, na.rm=TRUE, geom='line')

### 2016-02-18 R/clustering.r
minor fix

@@ -56,10 +56,10 @@ plot.clusters <- function(df.features, clusters, sizes, colors){
                  varname.size = 4) 
   p1 <- p1 + scale_color_manual(values = alpha(cluster.colors,0.75))
   p1 <- p1 + theme_bw() +
-    theme(text            = element_text(size = 15),
-          aspect.ratio    = 1,
-          legend.position = "none") +
-    ggtitle("Individual factor map (PCA)")
+        theme(text            = element_text(size = 15),
+              aspect.ratio    = 1,
+              legend.position = "none") +
+        ggtitle("Individual factor map (PCA)")
   print(p1) 
   dev.copy(png, paste0('2016-01-15-PCA2.png'), width=800, height=800)
   dev.off()

### 2016-02-18 pipeline_clustering.r

@@ -98,15 +98,15 @@ z <- cluster(features)
 # Update users df with their cluster
 df.users <-  data.frame(user=rownames(features), cluster=z) %>%
              merge(df.users, all=TRUE)
-df.users[is.na(df.users$cluster),]$cluster <- 0
+df.users[is.na(df.users$cluster),]$cluster <- 0 # non-active users go to cluster 0
 
 # Assign colors to clusters by size so that we use the same colors all along
 cluster.colors <- palette()[2:length(palette())] # 2: to avoid black
 
 # PCA and whisker plots
 #################################################### 
-#TODO: features should contain also inactive users
+#TODO: features should contain also inactive users?
 plot.clusters(features, 
               clusters = z,
-              sizes    = 0.5*log(rowSums(user.motifs)), 
+              sizes    = 0.5*log(rowSums(user.motifs[active.mask,])), 
               colors   = alpha(cluster.colors[z], 0.5))

### 2016-03-03 R/count_motifs.r
parallel neighborhood census

@@ -1,3 +1,6 @@
+library(RSQLite)
+source('R/extract_from_db.r')
+
 count_motifs_by_post <- function(threads, database='reddit'){
 
   con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
@@ -22,12 +25,14 @@ count_motifs_by_post <- function(threads, database='reddit'){
     #Extract the egos of every post
     g <- database.to.graph(threads[i], con, database)
     gp <- g$gp
+    root.post <- V(gp)[which(degree(gp, mode='out')==0)]$name
     egos <- make_ego_graph(gp, rad, nodes=V(gp))
     
-    if (FALSE){
+    if (TRUE){
       V(gp)$color <- "black"
-      if(vcount(gp)>10 && vcount(gp)<30){
-        par(mfrow=c(3,3))
+      #if(vcount(gp)>10 && vcount(gp)<30){
+      if(TRUE){
+        par(mfrow=c(1,1))
         la <- layout_with_fr(gp)
         plot(as.undirected(gp),
              layout = la, 
@@ -55,8 +60,9 @@ count_motifs_by_post <- function(threads, database='reddit'){
       V(eg)$color <- 1 
       V(eg)[u]$color <- 2 
       #V(eg)[V(eg)$user==user.name]$color <- 2 # any post from the user
-      V(eg)[V(eg)$user=="root"]$color <- 3
-    
+      #V(eg)[V(eg)$user=="root"]$color <- 3
+      V(eg)[V(eg)$name==root.post]$color <- 3
+      
       # Drop nodes that are not connected to the post
       eg <- delete_vertices(eg, which(distances(eg, u)==Inf))
 
@@ -88,6 +94,100 @@ count_motifs_by_post <- function(threads, database='reddit'){
   } # end threads
   
   posts.motifs <- data.frame(postid=posts.id, motif=posts.motifs.id)
+  
+  
+  # Sort by frequency (and relabel: 1 for the most frequent and so forth)
+  idx <- order(table(posts.motifs$motif), decreasing = TRUE)
+  posts.motifs$motif <- match(posts.motifs$motif, idx)
+  motifs <- motifs[idx]
+  
   return(list(posts.motifs = posts.motifs,
               motifs = motifs))
 }
+
+
+merge.motif.counts.pair <- function(res1, res2){
[…]

### 2016-03-03 R/load_participations.r

@@ -1,4 +1,5 @@
 library(RSQLite)
+library(data.table)
 source('R/extract_from_db.r')
 
 load_posts <- function(database='reddit', forum='podemos'){
@@ -13,7 +14,13 @@ load_posts <- function(database='reddit', forum='podemos'){
     #thread.posts <- dbGetQuery(con,  paste0("select user,date from posts where thread='", thread.ids[i,], "'"))
     thread.posts <- dbGetQuery(con,  paste0("SELECT postid, parent, user, date FROM posts WHERE thread='", thread.ids[i,], "'"))
     thread.posts$thread <- thread.ids[i,]
-    df.posts <- rbind(df.posts, thread.posts)
+    
+    #df.posts <- rbind(df.posts, thread.posts)
+    df.posts <- rbindlist(list(df.posts,thread.posts))
+    
+    if(i%%1000==0){
+      cat('\n', i, '/', nrow(thread.ids))
+    }
   }
-  return(df.posts)
+  return(data.frame(df.posts))
 }

### 2016-03-03 pipeline_clustering.r

@@ -9,7 +9,9 @@
 #https://www.reddit.com/r/france
 #https://www.reddit.com/r/catalunya
 #https://www.reddit.com/r/es
-
+library(parallel)
+library(doParallel)
+library(foreach)
 library(ggplot2)
 library(ggbiplot)
 library(gplots)
@@ -26,52 +28,66 @@ source('R/count_motifs.r')
 source('R/normalize_counts.r')
 source('R/clustering.r')
 
-MIN_POSTS <- 30 # number of post to consider a user as active
+MIN_POSTS <- 100 # number of post to consider a user as active
 ###################################################
 # Load data
 ###################################################
-df.posts <- load_posts(database='reddit', forum='podemos')
+#df.posts <- load_posts(database='reddit', forum='podemos')
+#save(df.posts,file="dfposts.Rda")
+load('dfposts.Rda')
+df.posts <- data.frame(df.posts)
+
 df.threads <- plyr::count(df.posts, "thread")
-df.users <- plyr::count(df.posts, 'user')
+df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
 names(df.threads)[2] <- "length"
 names(df.users)[2] <- "posts"
 
 # Compute neighborhood around every post
-#res <-count_motifs_by_post(df.threads$thread, database='reddit')
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/8000))
+
+##############################"
+# sequential
+#par(mfrow=c(1,1))
+#res.seq <- count_motifs_by_post(as.vector(unlist(chunks[1025])), database='reddit')
+#res <- res.seq
+#plot.motif.counts(res.seq)
+
+# parallel
+ncores <- detectCores() - 2
+cl<-makeCluster(ncores, outfile="", port=11439)
+registerDoParallel(cl)
+pck <- c('RSQLite')
+res.parallel <- foreach(i=1:4, .packages = pck)%dopar%{
+          source('R/extract_from_db.r')
+          count_motifs_by_post(chunks[[i]], database='reddit')
+}
+stopCluster(cl)
+res.parallel <- merge.motif.counts(res.parallel)
+res <- res.parallel
+plot.motif.counts(res)
+dev.copy(png, paste0('2016-01-15-motifs_4_4.png'))
+dev.off()
+
+#########################################
+
+
+
 
 df.post.motif  <- res$posts.motifs
 motifs <- res$motifs
 
[…]

### 2016-03-03 sandbox/parallel.r

@@ -1,0 +1,18 @@
+# 
+
+process_words <- function(words){
+  sum(length(words))
+  cbind(words)
+}
+
+words <- c("a", "b", "c", "d", "e", "f", "g", "h")
+
+
+ncores <- detectCores() - 2
+cl<-makeCluster(ncores, outfile="", port=11439)
+registerDoParallel(cl)
+
+res <- foreach(i=1:3)%dopar%{
+  process_words(words[i:(i+2)])
+}
+stopCluster(cl)

### 2016-03-04 doc/paper.tex
parallel neighborhood census
data.table to store neighborhoods faster

@@
 \DeclareMathOperator*{\argmax}{arg\,max}
-\journalname{Comput Stat}
+\journalname{}
 
 \begin{document}
-\title{Structural posts neighborhoods to characterize users in online forums}
+\title{Structural-temporal neighborhoods of posts to characterize users in online forums}
 
@@
 An interesting question in online communities is that concerning roles.
-In sociology, roles are generally seen as the set of expected behaviors that are attached  to a position in the community.
+In sociology, roles are generally seen as the set of expected behaviours that are attached  to a position in the community.
 Extrapoliting the notion of role, some researchers have looked for roles in online forums.
@@
 Roles can also be studied from the macro or the micro perspective.
-If studied from the macro, we can analyze the number of users, the percentage of replied post, its centrality in the network, and  so forth.
+If studied from the macro, we can analyse the number of users, the percentage of replied post, its centrality in the network, and  so forth.
 
-In this paper, we focus on the anamyse of roles at a micro level, and more specifically at the discussion level.
+In this paper, we focus on the analyse of roles at a micro level, and more specifically at the discussion level.
 We would like to answer the following question: are there different types of users in terms of the kind of conversation they participate in?
+
+Our intuition is that some users like participating in some kind of discussion rather than other.
+Certainly, an analysis of the textual content will tell us much about a discussion.
+However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humor, irony, or changing contexts, we turn our attention towards the structure of the discussions.
+More precisely, we analyze the local graph structure in which a user post in embedded, in the hope that this structure will be meaningful since it also reflects the kind of conversation in that part of the thread.
+
+Formally, these local graphs are known as neighborhoods.
 
@@
 First we will introduce our representation of online discussions.
-Then we will explain our concept of structural neighborhood.
+Then we will explain our concept of structural neighbourhood.
 Then we will apply this to the clustering of users and we will analyze the clusters.
@@
 \end{figure}
-\section{Posts graph neighborhoods}
 
-Our intuition is that some users like participating in some kind of discussion rather than other.
-Certainly, an analysis of the textual content will tell us much about a discussion.
-However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humor, irony, or changing contexts, we turn our attention towards the structure of the discussions.
-More precisely, we analyze the local graph structure in which a user post in embedded, in the hope that this structure will be meaningful since it also reflects the kind of conversation in that part of the thread.
+\section{Neighborhoods in discussion trees}
+In this section, we discuss about the convenience of classic neighbourhood definition in discussion trees.
+Then we introduce a new definition to overcome some of the limitations of the classic definition when dealing with temporal trees.
 
-Formally, these local graphs are known as neighborhoods.
-Let us consider first the most basic notion of neighborhood in order to understand its limitations.
-Let $p_i$ be a post in the tree graph $G$.
-The \textit{neighborhood of $p_i$ at radius $r$} $N_G(p_i,r)$ is the induced subgraph of $G$ consisting of all posts $j \in G$ at distance $d_{ij}\leq r$.
-In order to analyze the type of discussion where a user participates in, we might count the different neighborhoods in which her posts are embedded.
-However, the number of possible neighborhoods even with radius 1 is unbounded since, although a post can only have a parent, it can have an infinite number of children (see Figure~ \ref{fig:large_neighboorhood}).
+\begin{definition}
+Given a tree graph $G$, the \textit{structural neighborhood} of distance $d$ of post $i$ is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$
+\end{definition}
 
-In practice, if we create a matrix $M$ where rows correspond to users, columns correspond to neighborhoods and $M_{ij}$ indicates how many times user $i$ has a post in the neighborhood $j$,  we obtain a very sparse matrix.
-Moreover, many of the columns correspond to similar type of neighborhoods.
-Think, for instance, of a post with 1 reply, a post with 10 replies and another post with 15 replies.
-While the type of conversation that embeds the first posts is different from the other two, it does not make sense to say that the former two are categorically different.
+From this definition a structuarl neighborhood at any distance has an unbounded number of posts.
+Figure \ref{fig:large_neighboorhood}, for instance, shows a typical case where a post has a large neighborhood even at distance two since a post can only one parent but infinite children (replies).
+
+\subsection{Structural neighborhoods in discussion trees}
 
@@
 \end{figure}
-
+\subsection{Structuro-temporal neighborhoods in discussion trees}
 In order to reduce the space of possible neighborhoods we introduce the notion of \textit{temporal neighborhood}.
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{PCA}
+	\includegraphics[width=0.5\textwidth]{PCA_nocolor}
 	\includegraphics[width=0.5\textwidth]{whiskers}

### 2016-03-04 R/count_motifs.r

@@ -1,4 +1,5 @@
 library(RSQLite)
+library(data.table)
 source('R/extract_from_db.r')
 
 count_motifs_by_post <- function(threads, database='reddit'){
@@ -19,6 +20,7 @@ count_motifs_by_post <- function(threads, database='reddit'){
   plotted.trees <- 1
   
   nthreads <- length(threads)
+  posts.motifs <- data.frame()
   for(i in 1:nthreads){
     cat('\n', i, '/', nthreads)
     
@@ -88,13 +90,17 @@ count_motifs_by_post <- function(threads, database='reddit'){
         motifs[[motif.id]] <- eg
       }
 
-      posts.id <- c(posts.id, post.id)
-      posts.motifs.id <- c(posts.motifs.id, motif.id)
+      #posts.id <- c(posts.id, post.id)
+      #posts.motifs.id <- c(posts.motifs.id, motif.id)
+      
+      posts.motifs <- rbindlist(list(posts.motifs, 
+                                     data.frame(postid=post.id,motif=motif.id)))
     } # end egos
   } # end threads
   
-  posts.motifs <- data.frame(postid=posts.id, motif=posts.motifs.id)
-  
+  #posts.motifs <- data.frame(postid=posts.id, motif=posts.motifs.id)
+  posts.motifs <- data.frame(posts.motifs)
+  names(posts.motifs) <- c('postid', 'motif')
   
   # Sort by frequency (and relabel: 1 for the most frequent and so forth)
   idx <- order(table(posts.motifs$motif), decreasing = TRUE)

### 2016-03-04 pipeline_clustering.r

@@ -43,8 +43,8 @@ names(df.threads)[2] <- "length"
 names(df.users)[2] <- "posts"
 
 # Compute neighborhood around every post
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/8000))
-
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/3000))
+length(chunks)
 ##############################"
 # sequential
 #par(mfrow=c(1,1))
@@ -56,20 +56,20 @@ chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/8000))
 ncores <- detectCores() - 2
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
-pck <- c('RSQLite')
-res.parallel <- foreach(i=1:4, .packages = pck)%dopar%{
-          source('R/extract_from_db.r')
-          count_motifs_by_post(chunks[[i]], database='reddit')
+pck <- c('RSQLite', 'data.table')
+res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
+  source('R/extract_from_db.r')
+  count_motifs_by_post(chunks[[i]], database='reddit')
 }
 stopCluster(cl)
-res.parallel <- merge.motif.counts(res.parallel)
-res <- res.parallel
+res <- merge.motif.counts(res.parallel)
+save(res,file="res.Rda")
+
 plot.motif.counts(res)
 dev.copy(png, paste0('2016-01-15-motifs_4_4.png'))
 dev.off()
 
 #########################################
-
 
 
 

### 2016-03-06 doc/paper.tex
paper avanzado con la idea bayesiana de detectar cortes de dinamica

@@
 \begin{document}
-\title{Structural-temporal neighborhoods of posts to characterize users in online forums}
+\title{You are the way you structurally talk: Structural-temporal neighborhoods of posts to characterize users in online forums}
 
@@
 \section{Discussion trees}
-A very natural way to represent a conversation thread is a tree graph where vertex represent posts and edges represent replies from some post to another.
+A natural way to represent a conversation thread is a tree graph where vertex represent posts and edges represent replies from some post to another.
 The tree is rooted at the post who started the thread.
@@
 \section{Neighborhoods in discussion trees}
-In this section, we discuss about the convenience of classic neighbourhood definition in discussion trees.
-Then we introduce a new definition to overcome some of the limitations of the classic definition when dealing with temporal trees.
+
+Our goal is to characterize conversations by their local structure.
+By local structure we mean the structure of the graph that surrounds a given post, which we call \textit{neighbourhood}.
+Given a set of posts written by a user, analysing the neighbourhoods where user posts are embedded will give us an insight of what kind of user she is.
+It is important to note that, if the space of possible neighbourhoods is too small, we might not be able to say anything interesting, while if it is too large we might have neighbourhoods that correspond to the same type of conversation.
+
+In this section, we discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
+Then we introduce a new definition to overcome some of the limitations of the classic one when dealing with temporal trees.
+
+\subsection{Structural neighborhoods in discussion trees}
 
 \begin{definition}
-Given a tree graph $G$, the \textit{structural neighborhood} of distance $d$ of post $i$ is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$
+Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$.
 \end{definition}
+%The above definition is an extension of the classic definition of \textit{neighbourhood} which only considers adjacent vertices.
 
-From this definition a structuarl neighborhood at any distance has an unbounded number of posts.
-Figure \ref{fig:large_neighboorhood}, for instance, shows a typical case where a post has a large neighborhood even at distance two since a post can only one parent but infinite children (replies).
-
-\subsection{Structural neighborhoods in discussion trees}
+% pros and cons
+The above definition has two drawbacks when used in the context of conversation trees.
+%%% not all dynamics are captured
+First, the dynamics of the conversation (time or order in which posts are attached to the tree) are not entirely captured in the structure of a tree representation.
+We know, for instance, that the time in which a node was attached to the tree is always posterior to that of its parent.
+But it impossible to say, by just looking at the structure of the tree, the order in which a set of sibling posts replied to their common parent post.
+Since these siblings may get more replies themselves, a single \textit{structural neighbourhood} may sometimes correspond to very different dynamics (see Figure).
+%TODO: ejemplo artificial para demostrar que la definicion de vecindario estructural puede corresponder a dinamicas muy diferentes.
+%% unbounded size
+Second, a structural neighbourhood at a given radius $r$ has an unbounded number of posts, and therefore the space of possible neighbourhoods is infinite.
+This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent (see Figure)
 
@@
 	\includegraphics[width=0.4\textwidth]{small_neighborhood}
-	\caption{The size of the neighborhood at distance $R$ is unbounded.
-These two graph represent frequent neighborhoods of a post (red) that replied to the root (white).
+	\caption{The size of the neighbourhood with radius $r$ is unbounded.
+These two graph represent frequent neighbourhoods of a post (red) that replied to the root (white).
 However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
-	\label{fig:large_neighboorhood}
+	\label{fig:large_neighbourhood}
 \end{figure}
+
 \subsection{Structuro-temporal neighborhoods in discussion trees}
-In order to reduce the space of possible neighborhoods we introduce the notion of \textit{temporal neighborhood}.
-Let $t_i$ the time where post $p_i$ was added to the thread $G$.
-Let $\tau_{ij}=|t_i-t_j|$ be the \textit{temporal distance} between two posts $p_i$ and $p_j$.
-The \textit{temporal neighborhood} $N_{G}^{(t)}(i, r, n)$of spacial radius $r$ and order $n$ of a post $p_i$ is the set of $n$ posts in its spacial neighborhood $N_{G}(i, r)$ that are temporally closest to $i$ and that can reach $i$ trough posts that are also in the temporal neighborhood.
+\begin{definition}
+Given a tree graph $G$, the \textit{structural-temporal neighborhood} of distance $d$ of post $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighborhood composed the $n$ vertices that are closer to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
+\end{definition}
 
-Finally, we give their own colors the post $i$ and to the root, and we consider that two neighborhoods are equivalent if they are color-isomorphic.
+The above definition has two advantages over the previous one.
+First, dynamics are better taken into account since the neighbourhood only includes posts that are structurally and temporally close.
+Second, the size of the neighborhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.
 
-With the above definitions in mind, and for a given forum, we can build a matrix $M$ with the counts of neighborhoods for every user.
+Note that this definition does not consider the timestamp of posts but only the order in which they arrived.
+An alternative definition of structural-temporal neighborhood might include not the $o$  posts that are closest to the ego post, but every post $j$ at distance $|t_i-t_j|<\tau$ from the ego post $i$.
+However, both options consider that the local dynamic ends at a fixed horizon ($o$ or $\tau$).
+This is specially problematic for the case of $\tau$ since different conversation threads may grow at different speeds, as well as different parts of the same thread or even the same part of the thread at different moments (e.g.
+threads barely grow at night).
+If we were to use $\tau$ as temporal horizon, we need a mechanism to automatically decide the proper $\tau$ for every post.
+
+\begin{definition}
+Let $s={p^{(1)},...p^{(m)}}$ be the sequence of posts in the structural neighbourhood $\mathcal{N}_i(r)$ ordered by their temporal distance to $i$.
+Let $\pi=\{\Delta t_1,...,\Delta t_M\}$ be the sequence of time differences between a post in $s$ and its predecessor in $s$.
+The \textit{temporal radius} $\tau$ of the neighbourhood is the timestamp of the last post $k$ whose $\Delta t_k$ comes from the same distribution than its predecessors $\Delta t_{k-1},..., \Delta t_{1}$ (an intuition for this definition is given in Figure~\ref{fig:cutpoints})
+\end{definition}
+
+According to the above definition, to find the temporal radius we need to find the first outlier in a sequence of $\Delta t$.
+Assuming the $\Delta t_i$ comes from an Exponential distribution, and that its $\lambda$ parameter comes itself from a Gamma distribution, we can exploit this conjugacy and obtain the analytical form of the probability of a given $\Delta t_k$ of belonging to the same distribution than its predecessors:
+\begin{equation}
+p(\Delta t_k | \Delta t_1,...\Delta t_{k-1})
+=
+\int p(\Delta t_k | \lambda_{k-1})p(\lambda_{k-1} |\Delta t_1,...\Delta t_{k-1})
+=\frac{(\beta + n\overline{x})^{\alpha+n}}{(\beta + n\overline{x}+\Delta t_k)^{\alpha+n+1}}(\alpha+n)
+\end{equation}
+Details of this derivation are given in the Appendix.
+
+\begin{figure}
+\centering
+\includegraphics[width=1\textwidth]{temporal_cutpoint}
+\caption{Intuition to choice the temporal cutpoint of a neighbourhood.
+The two posts out of the box have a distance (time) to their parents clearly bigger than the average, so we consider that they are not part of the same local dynamic.}
+\label{fig:cutpoints}
+\end{figure}
+
+\subsection{Colored neighborhoods}
+In conversation trees, the root vertex plays an important role since it represents the post that opened the conversation and thus set the topic of the thread.
+Therefore, a reply to the root post is clearly different from a reply to any other post in the tree.
+We distinguish root posts from non-root posts by assigning root posts a unique color (white).
+We also assign a unique color to the ego post (red).
 
 \section{Application to Reddit forums}
+In this section, we use our definition of structural-temporal neighbourhood to detect different types of users according to the neighbourhoods in which their posts are embedded.
 
-We analyze the Podemos forum\footnote{https://www.reddit.com/r/podemos} in Reddit.
-It was was first used as a tool of internal democracy.
-Today, its members use it also to share and discuss about political news.
-The Podemos forum contains 995 discussion threads that started on January. These threads contain a total of 12912 posts written by 1218 users.
+\subsection{Podemos forum}
+Our dataset consists of all posts from March 2014 to May 2015 of the Podemos forum in Reddit\footnote{https://www.reddit.com/r/podemos}\footnote{I have also the data for \textit{gameofthrones, france, datascience, machinelearning complexsystems, philosophy, twoxchromosomes, trees, sex}}.
+The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014.
+Nowadays, its members use it mainly to share and discuss about political news.
+The forum contains 83,6119 posts spread over 47,803 threads and written by 26,193 users. (see Figure~\ref{fig:podemos_distributions})
 
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{threads_length}%
+\includegraphics[width=0.5\textwidth]{users_posts}
+\includegraphics[width=0.5\textwidth]{threads_length_cum}%
+\includegraphics[width=0.5\textwidth]{users_posts_cum}
+\caption{Log-log and cumulative distributions and of thread lengths and posts per user.}
+\label{fig:podemos_distributions}
+\end{figure}
+
+    \noindent\textcolor{red}{\rule{16cm}{1mm}}
+(Il fait que j'implemente l'équation 1 pour chercher les neighborhoods. Puis il faut voir si elle sert tout à fait à bien detecter la baisse du rythme)
 \subsection{Counting neighorhoods}
-In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighborhood structures.
+In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighbourhood structures.
 
@@
 	\includegraphics[width=0.8\textwidth]{neighborhoods_4_4}
-	\caption{Neighborhood counts with $r=4$ and $n=4$}
+	\caption{Most frequent neighborhoods with $r=4$ and $n=4$}
 	\label{fig:neighborhoods_4_4}

### 2016-03-06 pipeline_clustering.r

@@ -35,15 +35,35 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 #df.posts <- load_posts(database='reddit', forum='podemos')
 #save(df.posts,file="dfposts.Rda")
 load('dfposts.Rda')
-df.posts <- data.frame(df.posts)
-
+df.posts <- data.frame(df.posts) %>% arrange(date)
+df.posts <- df.posts[1:100000,]
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
 names(df.threads)[2] <- "length"
 names(df.users)[2] <- "posts"
 
+
+# Plot general overview of forum
+n <- hist(df.threads$length, breaks=0:max(df.threads$length))$counts
+plot(1:max(df.threads$length), n, log='xy', 
+     ylab='Number of threads', xlab='Length')
+title(main='Threads length')
+
+plot(cumsum(table(df.threads$length)), pch=19, cex=0.5,
+     ylab='Length (cum)', xlab='Thread')
+title('Threads length (cumulative)')
+
+n <- hist(df.users$posts, breaks=0:max(df.users$posts))$counts
+plot(1:max(df.users$posts), n, log='xy', 
+     ylab='Number of users', xlab='Posts')
+title(main='Users posts')
+
+plot(cumsum(table(df.users$posts)), pch=19, cex=0.5,
+     ylab='Posts (cum)', xlab='User')
+title('Users posts (cumulative)')
+
 # Compute neighborhood around every post
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/3000))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/1000))
 length(chunks)
 ##############################"
 # sequential
@@ -63,7 +83,7 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
 }
 stopCluster(cl)
 res <- merge.motif.counts(res.parallel)
-save(res,file="res.Rda")
+save(res,file="res_4_4_dyn.Rda")
 
 plot.motif.counts(res)
 dev.copy(png, paste0('2016-01-15-motifs_4_4.png'))
@@ -82,7 +102,9 @@ motifs <- res$motifs
 #motifs <- motifs[idx]
 
 # Add motif info to posts dataframe
-df.posts <- merge(df.posts, df.post.motif, all.x=TRUE, sort=FALSE)
[…]

### 2016-03-29 doc/paper.tex
neighborhoods bt order and time

@@
 Figure~\ref{fig:neighborhoods_4_4} shows the eight detected neighbors and their frequency.
-It is interesting to see different structures like chains or stars where the root node migh or might no appear and the ego post is placed at different positions.
+It is interesting to see different structures like chains or stars where the root node might or might no appear and the ego post is placed at different positions.
 
@@
 \section{Conclusions}
+
+\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%%
+% APPENDIX
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+\noindent \textbf{\Large Appendix}
+\vspace{-0.1cm}
+
+\begin{align}
+p(\Delta t_k | \Delta t_1,...\Delta t_{k-1})
+&=
+\int p(\Delta t_k | \lambda_{k-1})
+p(\lambda_{k-1} |\Delta t_1,...\Delta t_{k-1})\\
+&=\frac{(\beta + n\overline{x})^{\alpha+n}}{(\beta + n\overline{x}+\Delta t_k)^{\alpha+n+1}}(\alpha+n)
+\end{align}
+First we work on the posterior distribution over $\lambda_{k-1}$:
+\begin{align}
+p(\lambda_{k-1} |\Delta t_1,...\Delta t_{k-1})
+&\propto
+p(\Delta t_1,...\Delta t_{k-1} | \lambda_{k-1} |)p(\lambda_{k-1})\\
+&\propto 
+\prod \lambda e^{-\lambda x}  \lambda^{\alpha-1}e^{-\beta \lambda}\\
+&\propto
+\lambda^{\alpha+k-1}
+e^{-(\beta+k\overline{x})\lambda}\\
+&\propto\mathcal{G}(\alpha+k, \beta + n\overline{x})
+\end{align}
+
+And then the predictive posterior is:
+\begin{align}
+p(\Delta t_k | \Delta t_1,...\Delta t_{k-1})&=
+\int
+p(\Delta t_k | \lambda_{k-1})
+\mathcal{G}(\lambda_{k-1} | \alpha+k, \beta + n\overline{x})\\
+&=
+\frac{(\beta+n\overline{x})^{\alpha+k}}{\Gamma(\alpha+k)}
+\int
+\lambda e^{-\lambda \Delta_k}
+\lambda^{\alpha+k-1}
+e^{-(\beta+n\overline{x})\lambda}
+\text{d}\lambda\\
+&=
+\frac{(\beta+n\overline{x})^{\alpha+k}}{\Gamma(\alpha+k)}
+\int
+\lambda^{(\alpha+k+1)-1}
+e^{-(\beta+n\overline{x}+\Delta_k)\lambda}
+\text{d}\lambda\\
+&=
+\frac{(\beta+n\overline{x})^{\alpha+k}}{\Gamma(\alpha+k)}
+\frac{\Gamma(\alpha+k+1)}{(\beta+n\overline{x}+\Delta_k)^{\alpha+k+1}}\\
+&=
+\frac
+{(\beta+n\overline{x})^{\alpha+k}}{(\beta+n\overline{x}+\Delta_k)^{\alpha+k+1}}
+(\alpha+k)
+\end{align}
+
 \bibliographystyle{spbasic} 

### 2016-03-29 R/count_motifs.r

@@ -1,8 +1,12 @@
 library(RSQLite)
 library(data.table)
 source('R/extract_from_db.r')
-
-count_motifs_by_post <- function(threads, database='reddit'){
+source('R/tree_changepoints.r')
+count_motifs_by_post <- function(threads, 
+                                 database='reddit', 
+                                 neighbourhood=c('order', 'time'),
+                                 rad=3,
+                                 max.neighbors=4){
 
   con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
   
@@ -10,10 +14,6 @@ count_motifs_by_post <- function(threads, database='reddit'){
   # if neighborhood at distance 1 and binary tree, max nodes is 4
   # if neighborhood at distance 1 and binary tree, max nodes is 10
   # 4   10   22   30   46   62   93  123...
-  
-  # the maximum egonet will be a binary tree
-  rad <- 4
-  max.neighbors <- 4
   motifs <- list()
   posts.id <- vector()
   posts.motifs.id <- vector()
@@ -22,53 +22,43 @@ count_motifs_by_post <- function(threads, database='reddit'){
   nthreads <- length(threads)
   posts.motifs <- data.frame()
   for(i in 1:nthreads){
-    cat('\n', i, '/', nthreads)
+    cat('\n', i, '/', nthreads, '/', threads[i])
     
     #Extract the egos of every post
     g <- database.to.graph(threads[i], con, database)
     gp <- g$gp
+    
+    # cast dates to numeric
+    dates <- as.numeric(V(gp)$date)
+    gp <- remove.vertex.attribute(gp, "date")
+    V(gp)$date <- dates
+
     root.post <- V(gp)[which(degree(gp, mode='out')==0)]$name
-    egos <- make_ego_graph(gp, rad, nodes=V(gp))
     
-    if (TRUE){
-      V(gp)$color <- "black"
-      #if(vcount(gp)>10 && vcount(gp)<30){
-      if(TRUE){
-        par(mfrow=c(1,1))
-        la <- layout_with_fr(gp)
-        plot(as.undirected(gp),
-             layout = la, 
-             vertex.label = "",
-             vertex.size = 1.5 + 1.5 * log( graph.strength(gp), 3 ), 
-             edge.width = 1.5)  
-        dev.copy(png, paste0('2015-01-15-tree', plotted.trees, '.png'))
-        dev.off()
[…]

### 2016-03-29 R/forest.r

@@ -1,6 +1,7 @@
 library(igraph)
 library(rgexf)
 library(dplyr)
+
 forest <- function(df.posts){
   df.edges <- select(df.posts, postid, parent, date)
   df.vertices <- select(df.posts, postid, user, date)

### 2016-03-29 R/tree_changepoints.r

@@ -1,0 +1,343 @@
+# Functions to extract the neighborhood of a post in a tree
+# Function to detect the changepoints of a tree w.r.t time
+# between posts
+# author: Alberto Lumbreras
+##############################################################
+
+#t3_2eh5kz    141
+#t3_2dyzfs     36
+#t3_2dz84u     11
+#t3_2dz884     81
+#t3_2dz88i     84
+#t3_2e049q      7
+#t3_2e04x4     11
+#t3_2e05bb     14
+
+changepoints <- function(gp, vertical=FALSE, horizontal=FALSE){
+  # Detect the posts that took a time too long to the previous post
+  # - to its parent (vertical changepoint)
+  # - to its siblings (horizontal changepoints)
+  if(!is.numeric(V(gp)$date)){
+    stop('Dates are not numeric')
+  }
+  
+  breakpoints.v <- c()
+  breakpoints.h <- c()
+  
+  # Vertical changepoints (from root to every leaf)
+  ##################################################
+  if(vertical==TRUE){
+    leafs <- which(ego_size(gp, 1, mode='in', mindist=1)==0)
+    root <-  which(ego_size(gp, 1, mode='out', mindist=1)==0)
+    for(j in 1:length(leafs)){
+      #j<-17
+      # Path and its deltas 
+      path <- unlist(shortest_paths(gp, root, to=leafs[j], mode='all')$vpath)
+      if(length(path)<5){
+        next
+      }
+      dates <- as.numeric(V(gp)[path]$date)
+      deltas <-  diff(dates) 
+      scores <- deltas/quantile(deltas, 0.5)
+      breakpoints.v <- c(breakpoints.v, path[which(scores>2)+1]) # deltas had an offset w.r.t dates
+      #scores <- c(NA, scores)
+      #probs[path] <- scores
+      
+      # Debug
+      #par(mfrow=c(2,1))
+      #plot(dates, cex=1, type='b')
+      #cols <- rep('black', length(dates))
+      #cols[which(scores>2)] <- 'red'
+      #plot(scores, type='b', col=cols)
+    }
+    breakpoints.v <- names(breakpoints.v)
+    
+  }
+  
+  # Horizontal changepoints (between siblings)
+  ############################################
+  if(horizontal==TRUE){
+    candidates.h <- c()
+    dist.matrix <- distances(gp, mode = 'in')
[…]

### 2016-03-29 pipeline_clustering.r

@@ -34,13 +34,17 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 ###################################################
 #df.posts <- load_posts(database='reddit', forum='podemos')
 #save(df.posts,file="dfposts.Rda")
-load('dfposts.Rda')
+load('dfposts.Rda') # 836119 posts, 47803 threads
 df.posts <- data.frame(df.posts) %>% arrange(date)
-df.posts <- df.posts[1:100000,]
+df.posts <- df.posts[1:300000,]
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
 names(df.threads)[2] <- "length"
 names(df.users)[2] <- "posts"
+
+# Print dates range
+start.date <- as.POSIXct(min(as.numeric(df.posts$date)), origin = "1970-01-01") 
+end.date <- as.POSIXct(max(as.numeric(df.posts$date)), origin = "1970-01-01")
 
 
 # Plot general overview of forum
@@ -68,7 +72,14 @@ length(chunks)
 ##############################"
 # sequential
 #par(mfrow=c(1,1))
-#res.seq <- count_motifs_by_post(as.vector(unlist(chunks[1025])), database='reddit')
+res.seq <- count_motifs_by_post(as.vector(unlist(chunks[1])), 
+                                database='reddit',
+                                neighbourhood='order')
+threads <- chunks[[1]]
+res.seq.dyn <- count_motifs_by_post(threads[1:50], 
+                                    database='reddit',
+                                    neighbourhood='time')
+
 #res <- res.seq
 #plot.motif.counts(res.seq)
 
@@ -79,15 +90,21 @@ registerDoParallel(cl)
 pck <- c('RSQLite', 'data.table')
 res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   source('R/extract_from_db.r')
-  count_motifs_by_post(chunks[[i]], database='reddit')
+  count_motifs_by_post(chunks[[i]], 
+                       database='reddit',
+                       neighbourhood='order')
 }
 stopCluster(cl)
 res <- merge.motif.counts(res.parallel)
-save(res,file="res_4_4_dyn.Rda")
+save(res,file="res_3_4_dyn.Rda")
 
+# Plot found motifs and their frequency
 plot.motif.counts(res)
[…]

### 2016-03-29 pipeline_description.r

@@ -1,0 +1,150 @@
+library(dplyr)
+library(reshape2)
+library(igraph)
+library(ggplot2)
+
+source('R/load_participations.r')
+
+MIN_POSTS <- 100 # number of post to consider a user as active
+###################################################
+# Load data
+###################################################
+#df.posts <- load_posts(database='reddit', forum='podemos')
+#save(df.posts,file="dfposts.Rda")
+load('dfposts.Rda')
+df.posts <- as.data.frame(df.posts)
+df.posts <- data.frame(df.posts) %>% arrange(date)
+df.threads <- plyr::count(df.posts, "thread")
+df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
+names(df.threads)[2] <- "length"
+names(df.users)[2] <- "posts"
+df.posts$date <- as.numeric(df.posts$date)
+
+# Plot general overview of forum
+n <- hist(df.threads$length, breaks=0:max(df.threads$length), plot=FALSE)$counts
+plot(1:max(df.threads$length), n, log='xy', 
+     ylab='Number of threads', xlab='Length')
+title(main='Threads length')
+
+plot(cumsum(table(df.threads$length)), pch=19, cex=0.5,
+     ylab='Length (cum)', xlab='Thread')
+title('Threads length (cumulative)')
+
+n <- hist(df.users$posts, breaks=0:max(df.users$posts), plot=FALSE)$counts
+plot(1:max(df.users$posts), n, log='xy', 
+     ylab='Number of users', xlab='Posts')
+title(main='Users posts')
+
+plot(cumsum(table(df.users$posts)), pch=19, cex=0.5,
+     ylab='Posts (cum)', xlab='User')
+title('Users posts (cumulative)')
+
+# How are the delta times between posts?
+#######################################
+# delta to previous post in thread
+df.posts <- arrange(df.posts, thread, date) %>% 
+                              group_by(thread) %>% 
+                              mutate(rank = rank(date, ties.method = "first"), 
+                                     delta = c(NA, diff(date))) %>% as.data.frame
+# delta to parent
[…]

### 2016-03-29 sandbox/change_detection.r

@@ -1,0 +1,174 @@
+# Not use, but standard method
+tsoutliers <- function(x)
+{
+  # Standard method based on IQR
+  # http://stats.stackexchange.com/questions/1142/simple-algorithm-for-online-outlier-detection-of-a-generic-time-series
+  x <- as.ts(x)
+  tt <- 1:length(x)
+  resid <- residuals(loess(x ~ tt))
+  resid.q <- quantile(resid,prob=c(0.25,0.75))
+  iqr <- diff(resid.q)
+  limits <- resid.q + 1.5*iqr*c(-1,1)
+  score <- abs(pmin((resid-limits[1])/iqr,0) + pmax((resid - limits[2])/iqr,0))
+  return(score)
+}
+
+outliers.iqr <- function(x){ 
+  #deltas <- c(1,1,1,1,1,1,10,1,1,1,1,10,1,1,1)
+  # ok
+  
+  #deltas <- c(27175,  1613,  2075,  9607,  1037,  497)
+  # only detects the first one :(
+  
+  sequence <- deltas
+  q.25.75 <- quantile(sequence, c(0.25, 0.75)) 
+  iqr <- diff(q.25.75)
+  limits <- q.25.75 + 1.5*iqr*c(-1,1)
+  score.outlier.sup <- pmax((sequence - limits[2])/iqr,0)
+  return(score.outlier.sup)
+}
+
+outliers.bertrand.seq <- function(x){
+  # Outlier if it is 1.5 times higher than the maximum previous value 
+  # in the window from the last outlier until now
+  x <- c(1,1,1,1,1,1,10,1,1,1,1,10,1,1,1)
+  x <- c(27175,  1613,  2075,  9607,  1037,  497)
+  # does not detect the first and therefore sets the maximum too high :(
+  
+  outliers <- c()
+  scores<- rep(NA, length(x))
+  window.init <- 1
+  pos <- 2
+  while(pos<length(x)){
+    sequence <- x[window.init:(pos-1)]
+    score <- x[pos]/(1.5*max(sequence))
+    scores[pos] <- score
+    
+    cat("\ninit position:", window.init)
+    cat("\nlast position: ", pos-1)
+    cat("\nposition:", pos)
+    cat("\nin-window sequence: ", sequence)
+    cat("\nnew sample:", x[pos])
+    cat("\nmax delta:", 1.5*max(sequence))
+    cat("\nscore: ", score)
+    if(score>1){
+      cat('\n*******outlier')
[…]

### 2016-03-29 sandbox/posterior_expgamma.r

@@ -1,0 +1,25 @@
+posterior <- function(point=1, alpha=1, beta=1){
+    m <- 27175
+    n <- 1
+    
+    log(alpha+n) +
+    (alpha+n)*log(beta + n*m) -
+    (alpha+n+1)*log(beta + n*m + point)
+}
+
+x <- seq(1,10000, by=0.1)
+y <- exp(posterior(point=x, alpha=1, beta=1))
+plot(x, y, pch=19, cex=0.2, type='b')
+title(paste('alpha:', alpha, 'beta:', beta))
+
+posterior(point=1613, alpha=7000, beta=1)
+
+
+alpha <- 10
+beta <- 1
+nsamples <- 100
+x <- rep(NA,nsamples)
+for(i in 1:nsamples){
+  lam <- rgamma(1, shape=alpha, rate=beta)
+  x[i] <- rexp(1, rate=lam)
+}

### 2016-03-29 doc/paper.tex
minor

@@
 
-\subsection{Structural neighborhoods in discussion trees}
-
 \begin{definition}
@@
 
-\subsection{Structuro-temporal neighborhoods in discussion trees}
+\section{Structuro-temporal neighborhoods in discussion trees}
+\subsection{Order-based}
 \begin{definition}
@@
 
+\subsection{Time-based}
 Note that this definition does not consider the timestamp of posts but only the order in which they arrived.
@@
 \begin{definition}
+Given the set of posts from the root to any leaf, a post $p_i$ is a \textit{vertical breakpoint} if the time between $p_i$ and its parent is at least twice the median of the time lapses in the chain  $\Delta$ is   $\Delta_i < 2 \Delta_{median}$
+\end{definition}
+
+\begin{definition}
+Given the set of posts from the root to any leaf, a post is a \textit{horizontal breakpoint} if its $\Delta$ is   $\frac{\Delta_i}{\Delta_{median}}>2$
+\end{definition}
+
+\begin{definition}
 Let $s={p^{(1)},...p^{(m)}}$ be the sequence of posts in the structural neighbourhood $\mathcal{N}_i(r)$ ordered by their temporal distance to $i$.
-Let $\pi=\{\Delta t_1,...,\Delta t_M\}$ be the sequence of time differences between a post in $s$ and its predecessor in $s$.
-The \textit{temporal radius} $\tau$ of the neighbourhood is the timestamp of the last post $k$ whose $\Delta t_k$ comes from the same distribution than its predecessors $\Delta t_{k-1},..., \Delta t_{1}$ (an intuition for this definition is given in Figure~\ref{fig:cutpoints})
+The temporal neighborhood is the subgraph that belongs to the same dynamic than the ego post
+(an intuition for this definition is given in Figure~\ref{fig:cutpoints})
 \end{definition}
@@
 According to the above definition, to find the temporal radius we need to find the first outlier in a sequence of $\Delta t$.
-Assuming the $\Delta t_i$ comes from an Exponential distribution, and that its $\lambda$ parameter comes itself from a Gamma distribution, we can exploit this conjugacy and obtain the analytical form of the probability of a given $\Delta t_k$ of belonging to the same distribution than its predecessors:
-\begin{equation}
-p(\Delta t_k | \Delta t_1,...\Delta t_{k-1})
-=
-\int p(\Delta t_k | \lambda_{k-1})p(\lambda_{k-1} |\Delta t_1,...\Delta t_{k-1})
-=\frac{(\beta + n\overline{x})^{\alpha+n}}{(\beta + n\overline{x}+\Delta t_k)^{\alpha+n+1}}(\alpha+n)
-\end{equation}
-Details of this derivation are given in the Appendix.
+Assuming the $\Delta t_i$ comes from an Exponential distribution, and that its $\lambda$ parameter comes itself from a Gamma distribution, we can exploit this conjugacy and obtain the analytical form of the probability of a given $\Delta t_k$ of belonging to the same distribution than its predecessors.
 

### 2016-03-29 pipeline_clustering.r

@@ -37,6 +37,7 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 load('dfposts.Rda') # 836119 posts, 47803 threads
 df.posts <- data.frame(df.posts) %>% arrange(date)
 df.posts <- df.posts[1:300000,]
+df.posts <- df.posts[1:100000,]
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
 names(df.threads)[2] <- "length"
@@ -45,7 +46,8 @@ names(df.users)[2] <- "posts"
 # Print dates range
 start.date <- as.POSIXct(min(as.numeric(df.posts$date)), origin = "1970-01-01") 
 end.date <- as.POSIXct(max(as.numeric(df.posts$date)), origin = "1970-01-01")
-
+print(paste("Start date:", start.date))
+print(paste("End date:", end.date))
 
 # Plot general overview of forum
 n <- hist(df.threads$length, breaks=0:max(df.threads$length))$counts

### 2016-04-06 doc/IEEEtran/paper.tex
major

--- 2016-03-06 doc/paper.tex
@@
 \RequirePackage{fix-cm}
-%\documentclass[twocolumn]{svjour3}          % twocolumn
-\documentclass[smallextended]{svjour3}          % twocolumn
-
-\smartqed  % flush right qed marks, e.g. at end of proof
-
+\documentclass[9pt,technote]{IEEEtran}
+\newtheorem{definition}{Definition}
 \usepackage{graphicx} % adjustbox loads it
@@
 \usepackage[export]{adjustbox}
-\usepackage[labelfont=bf, labelsep=space]{caption}
+%\usepackage[labelfont=bf, labelsep=space]{caption}
 %\usepackage{subcaption}
-\usepackage{subfig}
+%\usepackage{subfig}
 \usepackage[utf8]{inputenc}
+\usepackage[]{algorithm2e}
 \usepackage{amssymb}
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
-\journalname{}
+%\journalname{}
 
 \begin{document}
-\title{You are the way you structurally talk: Structural-temporal neighborhoods of posts to characterize users in online forums}
+\title{You are the way you structurally talk: Structural-temporal neighbourhoods of posts to characterize users in online forums}
 
@@
         Bertrand Jouve
-}
-
-\institute{Alberto Lumbreras \and Marie Guégan \at
-		   Technicolor\\
-           975 Avenue des Champs Blancs, \\35576 Cesson-Sévigné,\\ France\\
-           \email{alberto.lumbreras@technicolor.com}\\
-           \email{marie.guegan@technicolor.com}
-           \and
-           Julien Velcin \at
-           Laboratoire ERIC, Université de Lyon,\\
-           5, avenue Pierre Mendès France, 69676 Bron,\\ France\\
-           \email{julien.velcin@univ-lyon2.fr}
-		  \and
-           Bertrand Jouve \at
-           Université de Toulouse; UT2; FRAMESPA/IMT; 5 allée Antonio Machado, 31058 Toulouse, cedex 9\\
-           CNRS; FRAMESPA; F-31000 Toulouse\\
-           CNRS; IMT; F-31000 Toulouse\\ France\\     
-           \email{jouve@univ-tlse2.fr}
 }
@@
 The popularization of online forums has brought a growing interest on their underlying dynamics.
-As any other complex system, an the dynamic of online forums can be studied at different levels, from the more macro to the most micro.
-Macro dynamics are, for instance, the evolution of some global properties such as its diameter, or its distribution degree.
+As any other complex system, the dynamic of online forums can be studied at different levels, from the most macro to the most micro.
+Macro dynamics are, for instance, the evolution of some global properties of the social graph such as its diameter, or its distribution degree.
 Micro dynamics are, for instance, the triadic motifs that represent local phenomena such as transitivity (friends of my friends are also my friends).
@@
 In sociology, roles are generally seen as the set of expected behaviours that are attached  to a position in the community.
-Extrapoliting the notion of role, some researchers have looked for roles in online forums.
+Extrapolating the notion of role, some researchers have looked for roles in online forums.
 Some others have tried to detect the roles and the users who hold that roles.
@@
 Certainly, an analysis of the textual content will tell us much about a discussion.
-However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humor, irony, or changing contexts, we turn our attention towards the structure of the discussions.
-More precisely, we analyze the local graph structure in which a user post in embedded, in the hope that this structure will be meaningful since it also reflects the kind of conversation in that part of the thread.
-
-Formally, these local graphs are known as neighborhoods.
+However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humour, irony, or changing contexts, we turn our attention towards the structure of the discussions.
+More precisely, we analyse the local graph structure in which a user post in embedded, in the hope that this structure will be meaningful since it also reflects the kind of conversation in that part of the thread.
+Formally, these local graphs are known as neighbourhoods.
 
 The remaining of the paper is as follows.
-First we will introduce our representation of online discussions.
-Then we will explain our concept of structural neighbourhood.
-Then we will apply this to the clustering of users and we will analyze the clusters.
-Finally, we will attemp to study whether these roles have predictive power.
-
-\section{Discussion trees}
-A natural way to represent a conversation thread is a tree graph where vertex represent posts and edges represent replies from some post to another.
-The tree is rooted at the post who started the thread.
-This representation allows to apply the mathematical toolbox of graph theory.
-Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.
+We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
+Then we introduce two new definitions that takes into account the time to overcome some of the limitations of the structural neighbourhood.
+We will finally apply these time-based neighbourhoods to detect clusters of users than tend to appear in the same time-based neighbourhoods.
 
@@
 
-\section{Neighborhoods in discussion trees}
+\section{Discussion trees}
+We represent a conversation thread as a tree graph where vertex represent posts and edges represent replies from some post to another.
+The tree is rooted at the post who started the thread.
+Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.
 
-Our goal is to characterize conversations by their local structure.
-By local structure we mean the structure of the graph that surrounds a given post, which we call \textit{neighbourhood}.
-Given a set of posts written by a user, analysing the neighbourhoods where user posts are embedded will give us an insight of what kind of user she is.
-It is important to note that, if the space of possible neighbourhoods is too small, we might not be able to say anything interesting, while if it is too large we might have neighbourhoods that correspond to the same type of conversation.
+The \textit{root} post of the tree is the post that starts the discussion. It is the only post that has no parent. A \textit{leaf} is a post with no replies.
+A \textit{branch} of the tree is the shortest path between the root and some of the leafs. Two branches may share their first posts.
 
-In this section, we discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
-Then we introduce a new definition to overcome some of the limitations of the classic one when dealing with temporal trees.
-
-\subsection{Structural neighborhoods in discussion trees}
+\section{Structural neighbourhoods in discussion trees}
+Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, in this paper we define the \textit{structural neighbourhood} as follows:
 
@@
 \end{definition}
-%The above definition is an extension of the classic definition of \textit{neighbourhood} which only considers adjacent vertices.
-
 % pros and cons
-The above definition has two drawbacks when used in the context of conversation trees.
+This definition has two drawbacks when used in the context of conversation trees.
 %%% not all dynamics are captured
@@
 But it impossible to say, by just looking at the structure of the tree, the order in which a set of sibling posts replied to their common parent post.
-Since these siblings may get more replies themselves, a single \textit{structural neighbourhood} may sometimes correspond to very different dynamics (see Figure).
+Thus, a single \textit{structural neighbourhood} may sometimes correspond to very different dynamics.
 %TODO: ejemplo artificial para demostrar que la definicion de vecindario estructural puede corresponder a dinamicas muy diferentes.
@@
 Second, a structural neighbourhood at a given radius $r$ has an unbounded number of posts, and therefore the space of possible neighbourhoods is infinite.
-This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent (see Figure)
+This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent (see Figure \ref{fig:large_neighbourhood}).
 
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{large_neighborhood}
-	\includegraphics[width=0.4\textwidth]{small_neighborhood}
+	\includegraphics[width=0.4\textwidth]{large_neighbourhood}
+	\includegraphics[width=0.4\textwidth]{small_neighbourhood}
 	\caption{The size of the neighbourhood with radius $r$ is unbounded.
@@
 
-\subsection{Structuro-temporal neighborhoods in discussion trees}
+\section{Structuro-temporal neighbourhoods in discussion trees}
+In this section, we propose two new time-based definition of neighbourhoods.
+
+\subsection{Order-based}
+Our first definition is based on the order in which posts are attached to the tree. 
 \begin{definition}
-Given a tree graph $G$, the \textit{structural-temporal neighborhood} of distance $d$ of post $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighborhood composed the $n$ vertices that are closer to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
+Given a tree graph $G$, the \textit{structural-temporal neighbourhood} of distance $d$ of post $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closer to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
 \end{definition}
 
-The above definition has two advantages over the previous one.
-First, dynamics are better taken into account since the neighbourhood only includes posts that are structurally and temporally close.
-Second, the size of the neighborhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.
+This definition has two advantages over the \textit{structural neighbourhood}.
+First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are structurally and temporally close.
+Second, the size of the neighbourhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.
 
-Note that this definition does not consider the timestamp of posts but only the order in which they arrived.
-An alternative definition of structural-temporal neighborhood might include not the $o$  posts that are closest to the ego post, but every post $j$ at distance $|t_i-t_j|<\tau$ from the ego post $i$.
-However, both options consider that the local dynamic ends at a fixed horizon ($o$ or $\tau$).
-This is specially problematic for the case of $\tau$ since different conversation threads may grow at different speeds, as well as different parts of the same thread or even the same part of the thread at different moments (e.g.
-threads barely grow at night).
-If we were to use $\tau$ as temporal horizon, we need a mechanism to automatically decide the proper $\tau$ for every post.
+\subsection{Time-based}
+Our second definition uses the real timestamps of posts in order to decide where the boundaries of a neighbourhood are.
+We might naively set fixed time-based boundaries for the neighbourhood by only including in the neighbourhood those posts whose timestamp $p_i$ is at distance less than $\tau$ from the ego post $|t_i-t_{ego}|<\tau$.
+However, the pace at which posts are added to the conversation may be very different between conversations, and even between different parts of the same conversation tree.
+Instead, we propose a definition based on the concept of $\textit{local dynamic}$.
 
 \begin{definition}
-Let $s={p^{(1)},...p^{(m)}}$ be the sequence of posts in the structural neighbourhood $\mathcal{N}_i(r)$ ordered by their temporal distance to $i$.
-Let $\pi=\{\Delta t_1,...,\Delta t_M\}$ be the sequence of time differences between a post in $s$ and its predecessor in $s$.
-The \textit{temporal radius} $\tau$ of the neighbourhood is the timestamp of the last post $k$ whose $\Delta t_k$ comes from the same distribution than its predecessors $\Delta t_{k-1},..., \Delta t_{1}$ (an intuition for this definition is given in Figure~\ref{fig:cutpoints})
+The temporal neighbourhood $\mathcal{N}_{i}^T(r)$ is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the posts belong to the same local dynamic than the ego post.
 \end{definition}
+Our definition of \textit{local dynamic} is based on the detection of changepoints in the time stamps of the posts.
+Given a sequence of consecutive posts in the same branch $b$ of a tree, denoted as $p_s,...,p_e$ we say that they belong to the same vertical dynamic if there is no (vertical) changepoint $p_i$ in $b$ such that $p_1 \prec p_i \preceq p_n$.
+Similarly, given a sequence of chronologically sorted  siblings $s$, denoted as $p_s,...,p_e$, we say that they belong to the same horizontal dynamic if there is no (horizontal) changepoint $p_i$ in $s$ such that $p_1 \prec p_i \preceq p_n$.
+Now we can outline the complete algorithm for time-based neighbourhood (Algorithm \ref{alg:temporal_neighbourhood}).
+Figure \ref{fig:cutpoints} illustrate a tree with a set of breakpoints and the temporal neighbourhood for a given node.
 
-According to the above definition, to find the temporal radius we need to find the first outlier in a sequence of $\Delta t$.
-Assuming the $\Delta t_i$ comes from an Exponential distribution, and that its $\lambda$ parameter comes itself from a Gamma distribution, we can exploit this conjugacy and obtain the analytical form of the probability of a given $\Delta t_k$ of belonging to the same distribution than its predecessors:
-\begin{equation}
-p(\Delta t_k | \Delta t_1,...\Delta t_{k-1})
-=
-\int p(\Delta t_k | \lambda_{k-1})p(\lambda_{k-1} |\Delta t_1,...\Delta t_{k-1})
-=\frac{(\beta + n\overline{x})^{\alpha+n}}{(\beta + n\overline{x}+\Delta t_k)^{\alpha+n+1}}(\alpha+n)
-\end{equation}
-Details of this derivation are given in the Appendix.
+\begin{algorithm}
+ \KwData{Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$}
+ \KwResult{Subgraph of $g$ with all vertices in $V(g)$}
+ Compute structural neighbourhood $\mathcal{N}_i(r)$\;
+ ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$\;
+ older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$\;
+ dump\_posts $\leftarrow \varnothing$\;
+ \For{bp $\in$ vertical breakpoints}{
+  \eIf{bp $\in$ ancestors}{
+   dump\_posts $\leftarrow$ dump\_posts $\cup$ ancestors(bp)\;
+   }{
+   dump\_posts $\leftarrow$ dump\_posts $\cup$ descendants(bp) $\cup$ bp\;
+  }
+ }
+  \For{bp $\in$ horizontal breakpoints}{
+  \eIf{bp $\in$ (older\_siblings $\cup$ ancestors)}{
+	dump\_posts $\leftarrow$ dump\_posts $\cup$ older\_siblings(bp)\;
+   }{
+	dump\_posts $\leftarrow$ dump\_posts $\cup$ younger\_siblings(bp) $\cup$ bp\;
+  }
+ }
+ $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$\;
+ \caption{Extraction of time-based neighbourhood}
+ \label{alg:temporal_neighbourhood}
+\end{algorithm}
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=1\textwidth]{temporal_cutpoint}
-\caption{Intuition to choice the temporal cutpoint of a neighbourhood.
-The two posts out of the box have a distance (time) to their parents clearly bigger than the average, so we consider that they are not part of the same local dynamic.}
+\includegraphics[width=1\textwidth]{breakpoints}
+\caption{Time-based neighbourhood.
+Horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.
+Post 6 is the ego, and posts with circular borders are those belonging to the time-based neighbourhood.}
 \label{fig:cutpoints}
-\end{figure}
+\end{figure*}
 
-\subsection{Colored neighborhoods}
+\subsection{Colored neighbourhoods}
 In conversation trees, the root vertex plays an important role since it represents the post that opened the conversation and thus set the topic of the thread.
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
@@
 \label{fig:podemos_distributions}
-\end{figure}
+\end{figure*}
 
     \noindent\textcolor{red}{\rule{16cm}{1mm}}
-(Il fait que j'implemente l'équation 1 pour chercher les neighborhoods. Puis il faut voir si elle sert tout à fait à bien detecter la baisse du rythme)
 \subsection{Counting neighorhoods}
@@
 
-If we set the a spatial radius $r=1$ and an order $n=4$, the temporal neighborhood $N_{G}^{(t)}(i, 1, 4)$ recovers the four adjacent posts that were written temporally closest to a given post.
-Figure~\ref{fig:neighborhoods_1_4} shows the four detected neighborhoods and their frequency.
+If we set the a spatial radius $r=1$ and an order $n=4$, the temporal neighbourhood $N_{G}^{(t)}(i, 1, 4)$ recovers the four adjacent posts that were written temporally closest to a given post.
+Figure~\ref{fig:neighbourhoods_1_4} shows the four detected neighbourhoods and their frequency.
 
-In order to detect more complex types of neighborhoods we enlarge the spatial radius to $r=4$.
-Figure~\ref{fig:neighborhoods_4_4} shows the eight detected neighbors and their frequency.
-It is interesting to see different structures like chains or stars where the root node migh or might no appear and the ego post is placed at different positions.
+In order to detect more complex types of neighbourhoods we enlarge the spatial radius to $r=4$.
+Figure~\ref{fig:neighbourhoods_4_4} shows the eight detected neighbors and their frequency.
+It is interesting to see different structures like chains or stars where the root node might or might no appear and the ego post is placed at different positions.
 
-\begin{figure}
+\begin{figure*}
 	\centering
-	\includegraphics[width=0.8\textwidth]{neighborhoods_1_4}
-	\caption{Neighborhood counts with $r=1$ and $n=4$}
-	\label{fig:neighborhoods_1_4}
-\end{figure}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_1_4}
+	\caption{neighbourhood counts with $r=1$ and $n=4$}
+	\label{fig:neighbourhoods_1_4}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
 	\centering
-	\includegraphics[width=0.8\textwidth]{neighborhoods_4_4}
-	\caption{Most frequent neighborhoods with $r=4$ and $n=4$}
-	\label{fig:neighborhoods_4_4}
-\end{figure}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_1}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_2}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_3}
+	\caption{Order-based neighbourhoods with $r=3$ and $n=4$}
+	\label{fig:neighbourhoods_4_4}
+\end{figure*}
 
-Note that some neighbors are still very similar.
+Note that some neighbours are still very similar.
 For instance, \textit{common answer 3} is very close to \textit{common answer 2} in the sense that they both represent the ego post replying to a root post that has one and two replies respectively.
@@
 
-Given the neighborhoods in which each user participates, we will analyze whether there exists different types of users or all users look similar.
+Given the neighbourhoods in which each user participates, we will analyze whether there exists different types of users or all users look similar.
 
-We want to make the analyze independent of the number of posts, and for that we normalize each user feature vector so that features indicate the percentage of posts in this kind of neighborhood.
-Moreover, some neighborhoods are much more common than others due to the nature of the forums.
+We want to make the analyze independent of the number of posts, and for that we normalize each user feature vector so that features indicate the percentage of posts in this kind of neighbourhood.
+Moreover, some neighbourhoods are much more common than others due to the nature of the forums.
 Thus, we normalize and scale the features so that every feature has a global mean 0 and variance 1.
@@
 Figure~\ref{fig:PCA} shows a PCA projection of the users colored by cluster and the distribution of the clusters in every dimension.
-\begin{figure}
+
+\begin{figure*}
 	\centering
-	\includegraphics[width=0.5\textwidth]{elbow}
-	\caption{Within-Sum of Squares under k-means}
-	\label{fig:elbow}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth]{PCA_nocolor}
-	\includegraphics[width=0.5\textwidth]{whiskers}
+	\includegraphics[width=1\textwidth]{PCA_clustering_3_4_order}
+	\includegraphics[width=1\textwidth]{clustering_3_4_order}
 	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
 	\label{fig:PCA}
-\end{figure}
-
-\subsection{Predictive roles}
-One interesting use of roles is to make predictions on users behaviors.
-Here, we will analyze whether the initial composition of roles in a thread is a good predictor of the final length of the thread.
+\end{figure*}
 
 \section{Conclusions}
+
 \bibliographystyle{spbasic} 

### 2016-04-06 doc/IEEEtran/paper_IEEEformat.tex
major

--- 2016-04-06 doc/IEEEtran/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[conference]{IEEEtran}
+% \documentclass[conference]{../sty/IEEEtran}
+
+% *** CITATION PACKAGES ***
 %
-% This is a general template file for the LaTeX package SVJour3
-% for Springer journals.          Springer Heidelberg 2010/09/16
+%\usepackage{cite}
+% cite.sty was written by Donald Arseneau
+% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
+% \cite{} output to follow that of the IEEE. Loading the cite package will
+% result in citation numbers being automatically sorted and properly
+% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
+% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
+% \cite will automatically add leading space, if needed. Use cite.sty's
+% noadjust option (cite.sty V3.8 and later) if you want to turn this off
+% such as if a citation ever needs to be enclosed in parenthesis.
+% cite.sty is already installed on most LaTeX systems. Be sure and use
+% version 5.0 (2009-03-20) and later if using hyperref.sty.
+% The latest version can be obtained at:
+% http://www.ctan.org/pkg/cite
+% The documentation is contained in the cite.sty file itself.
+
+% *** GRAPHICS RELATED PACKAGES ***
 %
-% Copy it to a new file with a new name and use it as the basis
-% for your article. Delete % signs as needed.
+\ifCLASSINFOpdf
+   \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../pdf/}{../jpeg/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+   \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  \graphicspath{ {./img/} }
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
+\graphicspath{ {./img/} }
+% graphicx was written by David Carlisle and Sebastian Rahtz. It is
+% required if you want graphics, photos, etc. graphicx.sty is already
+% installed on most LaTeX systems. The latest version and documentation
+% can be obtained at: 
+% http://www.ctan.org/pkg/graphicx
+% Another good source of documentation is "Using Imported Graphics in
+% LaTeX2e" by Keith Reckdahl which can be found at:
+% http://www.ctan.org/pkg/epslatex
 %
-% This template includes a few options for different layouts and
-% content for various journals. Please consult a previous issue of
-% your journal as needed.
+% latex, and pdflatex in dvi mode, support graphics in encapsulated
+% postscript (.eps) format. pdflatex in pdf mode supports graphics
+% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
+% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
+% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
+% which can result in "jaggedy"/blurry rendering of lines and letters as
+% well as large increases in file sizes.
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% You can find documentation about the pdfTeX application at:
+% http://www.tug.org/applications/pdftex
 
-\RequirePackage{fix-cm}
-\documentclass[9pt,technote]{IEEEtran}
+% *** MATH PACKAGES ***
+%
+%\usepackage{amsmath}
+% A popular package from the American Mathematical Society that provides
+% many useful and powerful commands for dealing with mathematics.
+%
+% Note that the amsmath package sets \interdisplaylinepenalty to 10000
+% thus preventing page breaks from occurring within multiline equations. Use:
+%\interdisplaylinepenalty=2500
+% after loading amsmath to restore such page breaks as IEEEtran.cls normally
+% does. amsmath.sty is already installed on most LaTeX systems. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/pkg/amsmath
+
+% *** SPECIALIZED LIST PACKAGES ***
+%
+\usepackage{algorithmic}
+% algorithmic.sty was written by Peter Williams and Rogerio Brito.
+% This package provides an algorithmic environment fo describing algorithms.
+% You can use the algorithmic environment in-text or within a figure
+% environment to provide for a floating algorithm. Do NOT use the algorithm
+% floating environment provided by algorithm.sty (by the same authors) or
+% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
+% algorithm float types and packages that provide these will not provide
+% correct IEEE style captions. The latest version and documentation of
+% algorithmic.sty can be obtained at:
+% http://www.ctan.org/pkg/algorithms
+% Also of interest may be the (relatively newer and more customizable)
+% algorithmicx.sty package by Szasz Janos:
+% http://www.ctan.org/pkg/algorithmicx
+
+% *** ALIGNMENT PACKAGES ***
+%
+%\usepackage{array}
+% Frank Mittelbach's and David Carlisle's array.sty patches and improves
+% the standard LaTeX2e array and tabular environments to provide better
+% appearance and additional user controls. As the default LaTeX2e table
+% generation code is lacking to the point of almost being broken with
+% respect to the quality of the end results, all users are strongly
+% advised to use an enhanced (at the very least that provided by array.sty)
+% set of table tools. array.sty is already installed on most systems. The
+% latest version and documentation can be obtained at:
+% http://www.ctan.org/pkg/array
+
+% IEEEtran contains the IEEEeqnarray family of commands that can be used to
+% generate multiline equations as well as matrices, tables, etc., of high
+% quality.
+
+% *** SUBFIGURE PACKAGES ***
+%\ifCLASSOPTIONcompsoc
+%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+%\else
+%  \usepackage[caption=false,font=footnotesize]{subfig}
+%\fi
+% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
+% for subfigure.sty, the latter of which is no longer maintained and is
+% incompatible with some LaTeX packages including fixltx2e. However,
+% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
+% which will override IEEEtran.cls' handling of captions and this will result
+% in non-IEEE style figure/table captions. To prevent this problem, be sure
+% and invoke subfig.sty's "caption=false" package option (available since
+% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
+% handling of captions.
+% Note that the Computer Society format requires a larger sans serif font
+% than the serif footnote size font used in traditional IEEE formatting
+% and thus the need to invoke different subfig.sty package options depending
+% on whether compsoc mode has been enabled.
+%
+% The latest version and documentation of subfig.sty can be obtained at:
+% http://www.ctan.org/pkg/subfig
+
+% *** FLOAT PACKAGES ***
+%
+%\usepackage{fixltx2e}
+% fixltx2e, the successor to the earlier fix2col.sty, was written by
+% Frank Mittelbach and David Carlisle. This package corrects a few problems
+% in the LaTeX2e kernel, the most notable of which is that in current
+% LaTeX2e releases, the ordering of single and double column floats is not
+% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
+% single column figure to be placed prior to an earlier double column
+% figure.
+% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
+% corrections already built into the system in which case a warning will
+% be issued if an attempt is made to load fixltx2e.sty as it is no longer
+% needed.
+% The latest version and documentation can be found at:
+% http://www.ctan.org/pkg/fixltx2e
+
+%\usepackage{stfloats}
+% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
+% the ability to do double column floats at the bottom of the page as well
+% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
+% LaTeX2e). It also provides a command:
+%\fnbelowfloat
+% to enable the placement of footnotes below bottom floats (the standard
+% LaTeX2e kernel puts them above bottom floats). This is an invasive package
+% which rewrites many portions of the LaTeX2e float routines. It may not work
+% with other packages that modify the LaTeX2e float routines. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/pkg/stfloats
+% Do not use the stfloats baselinefloat ability as the IEEE does not allow
+% \baselineskip to stretch. Authors submitting work to the IEEE should note
+% that the IEEE rarely uses double column equations and that authors should try
+% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
+% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
+% such ways.
+% Do not attempt to use stfloats with fixltx2e as they are incompatible.
+% Instead, use Morten Hogholm'a dblfloatfix which combines the features
+% of both fixltx2e and stfloats:
+%
+% \usepackage{dblfloatfix}
+% The latest version can be found at:
+% http://www.ctan.org/pkg/dblfloatfix
+
+% *** PDF, URL AND HYPERLINK PACKAGES ***
+%
+%\usepackage{url}
+% url.sty was written by Donald Arseneau. It provides better support for
+% handling and breaking URLs. url.sty is already installed on most LaTeX
+% systems. The latest version and documentation can be obtained at:
+% http://www.ctan.org/pkg/url
+% Basically, \url{my_url_here}.
+
+\usepackage[utf8]{inputenc}
+\usepackage{amssymb}
+\usepackage{algorithm}
 \newtheorem{definition}{Definition}
-\usepackage{graphicx} % adjustbox loads it
-\usepackage{epstopdf} % loads eps
-\usepackage[normalem]{ulem}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
 
-% insert here the call for the packages your document requires
-\usepackage{geometry}
-\usepackage[export]{adjustbox}
-%\usepackage[labelfont=bf, labelsep=space]{caption}
-%\usepackage{subcaption}
-%\usepackage{subfig}
-\usepackage[utf8]{inputenc}
-\usepackage[]{algorithm2e}
-\usepackage{amssymb}
-\usepackage{amsmath}
+% *** Do not adjust lengths that control margins, column widths, etc. ***
+% *** Do not use packages that alter fonts (such as pslatex).         ***
+% There should be no need to do such things with IEEEtran.cls V1.6 and later.
+% (Unless specifically asked to do so by the journal or conference you plan
+% to submit to, of course. )
 
-\usepackage{natbib} % enables author year and other citation styles
-\usepackage[bookmarks,bookmarksopen,bookmarksdepth=2]{hyperref} % active links
-\hypersetup{backref,
-	colorlinks=true,
-	citecolor=blue,
-	linkcolor=blue}
-
-\usepackage{tikz} % graphics,
-\usetikzlibrary{fit,positioning} % tikz elements positioning
-\usepackage{soul} % annotations
-
-\DeclareMathOperator*{\argmin}{arg\,min}
-
-\newcommand\alberto[1]{\textcolor{red}{#1}}
-
-%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
-
-\graphicspath{ {./img/} }
-\DeclareMathOperator*{\argmax}{arg\,max}
-%\journalname{}
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
 
 \begin{document}
-\title{You are the way you structurally talk: Structural-temporal neighbourhoods of posts to characterize users in online forums}
+%
+% paper title
+% Titles are generally capitalized except for words such as a, an, and, as,
+% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
+% not capitalized unless they are the first or last word of the title.
+% Linebreaks \\ can be used within to get better formatting as desired.
+% Do not put math or special symbols in the title.
+\title{You are the way you structurally talk: structural-temporal neighbourhoods of posts to characterize users in online forums}
 
-\author{Alberto Lumbreras \and
-        Julien Velcin  \and\\
-        Marie Guégan \and
-        Bertrand Jouve
-}
+% author names and affiliations
+% use a multiple column layout for up to three different
+% affiliations
+%\author{\IEEEauthorblockN{Alberto Lumbreras \\ and Bertrand Jouve}
+%\IEEEauthorblockA{Technicolor\\
+%Georgia Institute of Technology\\
+%Atlanta, Georgia 30332--0250\\
+%Email: alberto.lumbreras@irit.fr}
+%\and
+%\IEEEauthorblockN{Bertrand Jouve}
+%\IEEEauthorblockA{Twentieth Century Fox\\
+%Springfield, USA\\
+%Email: jouve@univ-tlse2.fr}
+%\and
+%\IEEEauthorblockN{Marie Guégan}
+%\IEEEauthorblockA{Technicolor Academy\\
+%San Francisco, California 96678--2391\\
+%Telephone: (800) 555--1212\\
+%Fax: (888) 555--1212}
+%\and
+%\IEEEauthorblockN{Marie Guégan}
+%\IEEEauthorblockA{Technicolor Academy\\
+%San Francisco, California 96678--2391\\
+%Telephone: (800) 555--1212\\
+%Fax: (888) 555--1212}}
 
-\date{Draft date: \today}
-% The correct dates will be entered by the editor
+% conference papers do not typically use \thanks and this command
+% is locked out in conference mode. If really needed, such as for
+% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
+% after \documentclass
 
+% for over three affiliations, or if they all won't fit within the width
+% of the page, use this alternative format:
+% 
+\author{\IEEEauthorblockN{Alberto Lumbreras\IEEEauthorrefmark{1},
+Bertrand Jouve\IEEEauthorrefmark{2},
+Marie Guégan\IEEEauthorrefmark{3}, 
+Julien Velcin\IEEEauthorrefmark{4}}
+\IEEEauthorblockA{\IEEEauthorrefmark{1}Technicolor, France\\Email: alberto.lumbreras@gmail.com}
+\IEEEauthorblockA{\IEEEauthorrefmark{2}Université de Toulouse; UT2; FRAMESPA/IMT,\\ 5 allée Antonio Machado, 31058 Toulouse, cedex 9\\
+Email: jouve@univ-tlse2.fr}
+\IEEEauthorblockA{\IEEEauthorrefmark{3}Technicolor\\975 Avenue des Champs Blancs\\35576 Cesson-Sevigné,\\France\\Email: marie.guegan@technicolor.com}
+\IEEEauthorblockA{\IEEEauthorrefmark{4}Laboratoire ERIC, Université de Lyon,\\ 5 avenue Pierre Mendès France, 69676, Bron\\France\\Email: julien.velcin@univ-lyon2.fr}}
+
+% use for special paper notices
+%\IEEEspecialpapernotice{(Invited Paper)}
+
+% make the title area
 \maketitle
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}\label{sec:introduction}
+% As a general rule, do not put math, special symbols or citations
+% in the abstract
+\begin{abstract}
+The abstract goes here.
+\end{abstract}
 
+% no keywords
+
+% For peer review papers, you can put extra information on the cover
+% page as needed:
+% \ifCLASSOPTIONpeerreview
+% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
+% \fi
+%
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+\IEEEpeerreviewmaketitle
+
+\section{Introduction}
 The popularization of online forums has brought a growing interest on their underlying dynamics.
@@
 	\centering
-	\includegraphics[width=0.32\textwidth]{tree1}
-	\includegraphics[width=0.32\textwidth]{tree2}
-	\includegraphics[width=0.32\textwidth]{tree3}
+	%\includegraphics[width=0.3\textwidth]{tree1}
+	\includegraphics[width=0.3\textwidth]{tree2}
+	%\includegraphics[width=0.3\textwidth]{tree3}
 	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{large_neighbourhood}
-	\includegraphics[width=0.4\textwidth]{small_neighbourhood}
+	\includegraphics[width=0.24\textwidth]{large_neighbourhood}
+	\includegraphics[width=0.24\textwidth]{small_neighbourhood}
 	\caption{The size of the neighbourhood with radius $r$ is unbounded.
@@
 Figure \ref{fig:cutpoints} illustrate a tree with a set of breakpoints and the temporal neighbourhood for a given node.
-
-\begin{algorithm}
- \KwData{Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$}
- \KwResult{Subgraph of $g$ with all vertices in $V(g)$}
- Compute structural neighbourhood $\mathcal{N}_i(r)$\;
- ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$\;
- older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$\;
- dump\_posts $\leftarrow \varnothing$\;
- \For{bp $\in$ vertical breakpoints}{
-  \eIf{bp $\in$ ancestors}{
-   dump\_posts $\leftarrow$ dump\_posts $\cup$ ancestors(bp)\;
-   }{
-   dump\_posts $\leftarrow$ dump\_posts $\cup$ descendants(bp) $\cup$ bp\;
-  }
- }
-  \For{bp $\in$ horizontal breakpoints}{
-  \eIf{bp $\in$ (older\_siblings $\cup$ ancestors)}{
-	dump\_posts $\leftarrow$ dump\_posts $\cup$ older\_siblings(bp)\;
-   }{
-	dump\_posts $\leftarrow$ dump\_posts $\cup$ younger\_siblings(bp) $\cup$ bp\;
-  }
- }
- $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$\;
+\begin{algorithm}[H]
+\begin{algorithmic}
+\REQUIRE Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$
+\ENSURE Subgraph of $g$ with all vertices in $V(g)$
+\STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
+\STATE ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$
+\STATE older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$
+\STATE dump $\leftarrow \varnothing$
+\FOR{bp $\in$ vertical breakpoints}
+ \IF{bp $\in$ ancestors}
+   \STATE dump $\leftarrow$ dump $\cup$ ancestors(bp)
+ \ELSE
+   \STATE dump $\leftarrow$ dump $\cup$ descendants(bp) $\cup$ bp
+  \ENDIF
+\ENDFOR
+\FOR{bp $\in$ horizontal breakpoints}
+  \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
+     \STATE dump $\leftarrow$ dump $\cup$ older\_siblings(bp)
+   \ELSE
+     \STATE dump $\leftarrow$ dump $\cup$ younger\_siblings(bp) $\cup$ bp
+  \ENDIF
+\ENDFOR
+\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$
+\end{algorithmic}
 \caption{Extraction of time-based neighbourhood}
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-\includegraphics[width=1\textwidth]{breakpoints}
+\includegraphics[width=0.45\textwidth]{breakpoints}
 \caption{Time-based neighbourhood.
@@
 \label{fig:cutpoints}
-\end{figure*}
+\end{figure}
 
@@
 
-    \noindent\textcolor{red}{\rule{16cm}{1mm}}
+ %\noindent\textcolor{red}{\rule{16cm}{1mm}}
 \subsection{Counting neighorhoods}
@@
 	\label{fig:neighbourhoods_4_4}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_time_1}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_time_2}
+	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
+	\caption{Time-based neighbourhoods (most frequent)}
+	\label{fig:neighbourhoods_time}
 \end{figure*}
@@
 
-\bibliographystyle{spbasic} 
-%\bibliographystyle{chicago} 
-%\bibliographystyle{plainnat} 
-\bibliography{library}
+% use section* for acknowledgment
+\section*{Acknowledgment}
+
+The authors would like to thank...
+
+% trigger a \newpage just before the given reference
+% number - used to balance the columns on the last page
+% adjust value as needed - may need to be readjusted if
+% the document is modified later
+%\IEEEtriggeratref{8}
+% The "triggered" command can be changed if desired:
+%\IEEEtriggercmd{\enlargethispage{-5in}}
+
+% references section
+
+% can use a bibliography generated by BibTeX as a .bbl file
+% BibTeX documentation can be easily obtained at:
+% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
+% The IEEEtran BibTeX style support page is at:
+% http://www.michaelshell.org/tex/ieeetran/bibtex/
+%\bibliographystyle{IEEEtran}
+% argument is your BibTeX string definitions and bibliography database(s)
+%\bibliography{IEEEabrv,../bib/paper}
+%
+% <OR> manually copy in the resultant .bbl file
+% set second argument of \begin to the number of references
+% (used to reserve space for the reference number labels box)
+\begin{thebibliography}{1}
+
+\bibitem{IEEEhowto:kopka}
+H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
+  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+
+\end{thebibliography}
+
+% that's all folks
 \end{document}

### 2016-04-06 doc/paper.tex
major

@@
 %\documentclass[twocolumn]{svjour3}          % twocolumn
-\documentclass[smallextended]{svjour3}          % twocolumn
-
-\smartqed  % flush right qed marks, e.g. at end of proof
+%\documentclass[smallextended]{svjour3}          % twocolumn
+\documentclass[9pt,technote]{IEEEtran}
+%\smartqed  % flush right qed marks, e.g. at end of proof
 
@@
 \usepackage[utf8]{inputenc}
+\usepackage[]{algorithm2e}
 \usepackage{amssymb}
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
-\journalname{}
+%\journalname{}
 
@@
 More precisely, we analyze the local graph structure in which a user post in embedded, in the hope that this structure will be meaningful since it also reflects the kind of conversation in that part of the thread.
-
 Formally, these local graphs are known as neighborhoods.
@@
 The remaining of the paper is as follows.
-First we will introduce our representation of online discussions.
-Then we will explain our concept of structural neighbourhood.
-Then we will apply this to the clustering of users and we will analyze the clusters.
-Finally, we will attemp to study whether these roles have predictive power.
-
-\section{Discussion trees}
-A natural way to represent a conversation thread is a tree graph where vertex represent posts and edges represent replies from some post to another.
-The tree is rooted at the post who started the thread.
-This representation allows to apply the mathematical toolbox of graph theory.
-Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.
+We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
+Then we introduce two new definitions that takes into account the time to overcome some of the limitations of the structural neighbourhood.
+We will finally apply these time-based neighbourhoods to detect clusters of users than tend to appear in the same time-based neighbourhoods.
 
@@
 
-\section{Neighborhoods in discussion trees}
+\section{Discussion trees}
+We represent a conversation thread as a tree graph where vertex represent posts and edges represent replies from some post to another.
+The tree is rooted at the post who started the thread.
+Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.
 
-Our goal is to characterize conversations by their local structure.
-By local structure we mean the structure of the graph that surrounds a given post, which we call \textit{neighbourhood}.
-Given a set of posts written by a user, analysing the neighbourhoods where user posts are embedded will give us an insight of what kind of user she is.
-It is important to note that, if the space of possible neighbourhoods is too small, we might not be able to say anything interesting, while if it is too large we might have neighbourhoods that correspond to the same type of conversation.
+The \textit{root} post of the tree is the post that starts the discussion. It is the only post that has no parent. A \textit{leaf} is a post with no replies.
+A \textit{branch} of the tree is the shortest path between the root and some of the leafs. Two branches may share their first posts.
 
-In this section, we discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
-Then we introduce a new definition to overcome some of the limitations of the classic one when dealing with temporal trees.
+\section{Structural neighborhoods in discussion trees}
+Extending the classic definition of neighbourhood according to which two vertex are neighbours if the distance between them is one, in this paper we define the \textit{structural neighbourhood} as follows:
 
@@
 \section{Structuro-temporal neighborhoods in discussion trees}
+In this section, we propose two new time-based definition of neighbourhoods.
+
 \subsection{Order-based}
+Our first definition is based on the order in which posts are attached to the tree. 
 \begin{definition}
@@
 
-The above definition has two advantages over the previous one.
-First, dynamics are better taken into account since the neighbourhood only includes posts that are structurally and temporally close.
-Second, the size of the neighborhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.
+This definition has two advantages over the \textit{structural neighbourhood}.
+First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are structurally and temporally close.
+Second, the size of the neighbourhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.
 
 \subsection{Time-based}
-Note that this definition does not consider the timestamp of posts but only the order in which they arrived.
-An alternative definition of structural-temporal neighborhood might include not the $o$  posts that are closest to the ego post, but every post $j$ at distance $|t_i-t_j|<\tau$ from the ego post $i$.
-However, both options consider that the local dynamic ends at a fixed horizon ($o$ or $\tau$).
-This is specially problematic for the case of $\tau$ since different conversation threads may grow at different speeds, as well as different parts of the same thread or even the same part of the thread at different moments (e.g.
-threads barely grow at night).
-If we were to use $\tau$ as temporal horizon, we need a mechanism to automatically decide the proper $\tau$ for every post.
+Our second definition uses the real timestamps of posts in order to decide where the boundaries of a neighbourhood are.
+We might naively set fixed time-based boundaries for the neighbourhood by only including in the neighbourhood those posts whose timestamp $p_i$ is at distance less than $\tau$ from the ego post $|t_i-t_{ego}|<\tau$.
+However, the pace at which posts are added to the conversation may be very different between conversations, and even between different parts of the same conversation tree.
+Instead, we propose a definition based on the concept of $\textit{local dynamic}$.
 
 \begin{definition}
-Given the set of posts from the root to any leaf, a post $p_i$ is a \textit{vertical breakpoint} if the time between $p_i$ and its parent is at least twice the median of the time lapses in the chain  $\Delta$ is   $\Delta_i < 2 \Delta_{median}$
+The temporal neighbourhood $\mathcal{N}_{i}^T(r)$ is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the posts belong to the same local dynamic than the ego post.
 \end{definition}
+Our definition of \textit{local dynamic} is based on the detection of changepoints in the time stamps of the posts.
+Given a sequence of consecutive posts in the same branch $b$ of a tree, denoted as $p_s,...,p_e$ we say that they belong to the same vertical dynamic if there is no (vertical) changepoint $p_i$ in $b$ such that $p_1 \prec p_i \preceq p_n$.
+Similarly, given a sequence of consecutive siblings $s$, denoted as $p_s,...,p_e$, we say that they belong to the same horizontal dynamic if there is no (horizontal) changepoint $p_i$ in $s$ such that $p_1 \prec p_i \preceq p_n$.
+Now we can outline the complete algorithm for time-based neighborhood (Algorithm \ref{alg:temporal_neighbourhood}).
+Figure \ref{fig:cutpoints} illustrate a tree with a set of breakpoints and the temporal neighbourhood for a given node.
 
-\begin{definition}
-Given the set of posts from the root to any leaf, a post is a \textit{horizontal breakpoint} if its $\Delta$ is   $\frac{\Delta_i}{\Delta_{median}}>2$
-\end{definition}
-
-\begin{definition}
-Let $s={p^{(1)},...p^{(m)}}$ be the sequence of posts in the structural neighbourhood $\mathcal{N}_i(r)$ ordered by their temporal distance to $i$.
-The temporal neighborhood is the subgraph that belongs to the same dynamic than the ego post
-(an intuition for this definition is given in Figure~\ref{fig:cutpoints})
-\end{definition}
-
-According to the above definition, to find the temporal radius we need to find the first outlier in a sequence of $\Delta t$.
-Assuming the $\Delta t_i$ comes from an Exponential distribution, and that its $\lambda$ parameter comes itself from a Gamma distribution, we can exploit this conjugacy and obtain the analytical form of the probability of a given $\Delta t_k$ of belonging to the same distribution than its predecessors.
+\begin{algorithm}[H]
+ \KwData{Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$}
+ \KwResult{Subgraph of $g$ with all vertices in $V(g)$}
+ Compute structural neighbourhood $\mathcal{N}_i(r)$\;
+ ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$\;
+ older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$\;
+ posts $\leftarrow \varnothing$\;
+ \For{bp $\in$ vertical breakpoints}{
+  \eIf{bp $\in$ ancestors}{
+   posts $\leftarrow$ posts $\cup$ ancestors(bp)\;
+   }{
+   posts $\leftarrow$ posts $\cup$ descendants(bp) $\cup$ bp\;
+  }
+ }
+  \For{bp $\in$ horizontal breakpoints}{
+  \eIf{bp $\in$ (older\_siblings $\cup$ ascendants)}{
+	posts $\leftarrow$ posts $\cup$ older\_siblings(bp)\;
+   }{
+	posts $\leftarrow$ posts $\cup$ younger\_siblings(bp) $\cup$ bp\;
+  }
+ }
+ $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(posts) from $\mathcal{N}_i(r)$\;
+ \caption{Extraction of time-based neighbourhood}
+ \label{alg:temporal_neighbourhood}
+\end{algorithm}
 
@@
 \centering
-\includegraphics[width=1\textwidth]{temporal_cutpoint}
-\caption{Intuition to choice the temporal cutpoint of a neighbourhood.
-The two posts out of the box have a distance (time) to their parents clearly bigger than the average, so we consider that they are not part of the same local dynamic.}
+\includegraphics[width=1\textwidth]{breakpoints}
+\caption{Time-based neighbourhood.
+Horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.
+Post 6 is the ego, and posts with circular borders are those belonging to the time-based neighborhood.}
 \label{fig:cutpoints}
@@
     \noindent\textcolor{red}{\rule{16cm}{1mm}}
-(Il fait que j'implemente l'équation 1 pour chercher les neighborhoods. Puis il faut voir si elle sert tout à fait à bien detecter la baisse du rythme)
 \subsection{Counting neighorhoods}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{neighborhoods_4_4}
-	\caption{Most frequent neighborhoods with $r=4$ and $n=4$}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_1}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_2}
+	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_3}
+	\caption{Order-based neighborhoods with $r=3$ and $n=4$}
 	\label{fig:neighborhoods_4_4}
@@
 
-Note that some neighbors are still very similar.
+Note that some neighbours are still very similar.
 For instance, \textit{common answer 3} is very close to \textit{common answer 2} in the sense that they both represent the ego post replying to a root post that has one and two replies respectively.
@@
 Figure~\ref{fig:PCA} shows a PCA projection of the users colored by cluster and the distribution of the clusters in every dimension.
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth]{elbow}
-	\caption{Within-Sum of Squares under k-means}
-	\label{fig:elbow}
-\end{figure}
 
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{PCA_nocolor}
-	\includegraphics[width=0.5\textwidth]{whiskers}
+	\includegraphics[width=1\textwidth]{PCA_clustering_3_4_order}
+	\includegraphics[width=1\textwidth]{clustering_3_4_order}
 	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
@@
 \end{figure}
-
-\subsection{Predictive roles}
-One interesting use of roles is to make predictions on users behaviors.
-Here, we will analyze whether the initial composition of roles in a thread is a good predictor of the final length of the thread.
 

### 2016-04-06 R/clustering.r

@@ -1,10 +1,10 @@
-cluster <- function(df.user_features){
-    wss <- (nrow(df.user_features)-1)*sum(apply(df.user_features,2,var))
+cluster <- function(features){
+    wss <- (nrow(features)-1)*sum(apply(features,2,var))
     for (i in 2:25){
-      wss[i] <- sum(kmeans(df.user_features, i)$withinss)
+      wss[i] <- sum(kmeans(features, i)$withinss)
     }
     plot(1:25, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")
-    fit <- kmeans(df.user_features, 5)
+    fit <- kmeans(features, 3)
     z <- fit$cluster
     
     # relabel cluster by size (to keep color consistency between executions)

### 2016-04-06 R/count_motifs.r

@@ -22,11 +22,16 @@ count_motifs_by_post <- function(threads,
   nthreads <- length(threads)
   posts.motifs <- data.frame()
   for(i in 1:nthreads){
-    cat('\n', i, '/', nthreads, '/', threads[i])
-    
     #Extract the egos of every post
     g <- database.to.graph(threads[i], con, database)
     gp <- g$gp
+    
+    size <- vcount(gp)
+    cat('\n', i, '/', nthreads, '/', threads[i], ' size: ', size)
+    if(size>100){
+      cat("\n WARNING: skipping thread of size ", size)
+      next 
+    }
     
     # cast dates to numeric
     dates <- as.numeric(V(gp)$date)
@@ -36,9 +41,17 @@ count_motifs_by_post <- function(threads,
     root.post <- V(gp)[which(degree(gp, mode='out')==0)]$name
     
     if(neighbourhood=='time'){
-      breakpoints <- changepoints(gp, vertical=T, horizontal=T)
+      breakpoints <- changepoints.cp(gp, vertical=T, horizontal=T)
       breakpoints.h <- breakpoints$breakpoints.h
       breakpoints.v <- breakpoints$breakpoints.v
+      breakpoints.vh <- breakpoints$breakpoints.vh
+      
+      V(gp)$bp.v <- FALSE
+      V(gp)$bp.h <- FALSE
+      V(gp)$bp.vh <- FALSE
+      V(gp)[breakpoints.v]$bp.v <- TRUE
+      V(gp)[breakpoints.h]$bp.h <- TRUE
+      V(gp)[breakpoints.vh]$bp.vh <- TRUE
     }
     
     for(j in 1:vcount(gp)){
@@ -55,8 +68,10 @@ count_motifs_by_post <- function(threads,
       # See if it matches any seen motif
       mypalette <- c("black", "red", "white")
       u <- V(eg)[V(eg)$name==post.id] # identify the ego vertex
+      uu <- V(eg)[V(eg)$user==user.name]
       V(eg)$color <- 1 
-      V(eg)[u]$color <- 2 
+      V(eg)[uu]$color <- 4
+      V(eg)[u]$color <- 2
       V(eg)[V(eg)$name==root.post]$color <- 3
       
       is.new <- TRUE
@@ -167,7 +182,7 @@ plot.motif.counts <- function(res){
   ####################################################
   # Plot found neighborhoods
   ####################################################
[…]

### 2016-04-06 R/tree_changepoints.r

@@ -3,6 +3,7 @@
 # between posts
 # author: Alberto Lumbreras
 ##############################################################
+library(changepoint)
 
 #t3_2eh5kz    141
 #t3_2dyzfs     36
@@ -12,8 +13,9 @@
 #t3_2e049q      7
 #t3_2e04x4     11
 #t3_2e05bb     14
-
-changepoints <- function(gp, vertical=FALSE, horizontal=FALSE){
+DEBUG <- FALSE
+changepoints.deltasequence <- function(gp, vertical=FALSE, horizontal=FALSE){
+  # Changepoints based of delta deviations from the median in the sequence
   # Detect the posts that took a time too long to the previous post
   # - to its parent (vertical changepoint)
   # - to its siblings (horizontal changepoints)
@@ -40,15 +42,6 @@ changepoints <- function(gp, vertical=FALSE, horizontal=FALSE){
       deltas <-  diff(dates) 
       scores <- deltas/quantile(deltas, 0.5)
       breakpoints.v <- c(breakpoints.v, path[which(scores>2)+1]) # deltas had an offset w.r.t dates
-      #scores <- c(NA, scores)
-      #probs[path] <- scores
-      
-      # Debug
-      #par(mfrow=c(2,1))
-      #plot(dates, cex=1, type='b')
-      #cols <- rep('black', length(dates))
-      #cols[which(scores>2)] <- 'red'
-      #plot(scores, type='b', col=cols)
     }
     breakpoints.v <- names(breakpoints.v)
     
@@ -103,121 +96,302 @@ changepoints <- function(gp, vertical=FALSE, horizontal=FALSE){
        breakpoints.h = V(gp)[breakpoints.h])
 }
 
+changepoints.bertrand <- function(gp, vertical=FALSE, horizontal=FALSE){
+  # Changepoints based of delta deviations from the last delta
+  # Detect the posts that took a time too long to the previous post
+  # - to its parent (vertical changepoint)
+  # - to its siblings (horizontal changepoints)
+  if(!is.numeric(V(gp)$date)){
+    stop('Dates are not numeric')
+  }
+  
+  breakpoints.v <- c()
+  breakpoints.h <- c()
+  
+  # Vertical changepoints (from root to every leaf)
+  ##################################################
+  if(vertical==TRUE){
+    nascendants <- ego_size(gp, 2, mode='out', mindist=1)
+    candidates <- V(gp)[which(nascendants==2)]
[…]

### 2016-04-06 pipeline_clustering.r

@@ -38,6 +38,11 @@ load('dfposts.Rda') # 836119 posts, 47803 threads
 df.posts <- data.frame(df.posts) %>% arrange(date)
 df.posts <- df.posts[1:300000,]
 df.posts <- df.posts[1:100000,]
+#df.posts <- df.posts[30000:60000,] # test debug
+#df.posts <- df.posts[50000:60000,] # test debug
+df.posts <- df.posts[60000:100000,] # test debug
+
+
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
 names(df.threads)[2] <- "length"
@@ -69,7 +74,7 @@ plot(cumsum(table(df.users$posts)), pch=19, cex=0.5,
 title('Users posts (cumulative)')
 
 # Compute neighborhood around every post
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/1000))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/175))
 length(chunks)
 ##############################"
 # sequential
@@ -78,10 +83,10 @@ res.seq <- count_motifs_by_post(as.vector(unlist(chunks[1])),
                                 database='reddit',
                                 neighbourhood='order')
 threads <- chunks[[1]]
-res.seq.dyn <- count_motifs_by_post(threads[1:50], 
+res.seq.dyn <- count_motifs_by_post(as.vector(unlist(chunks[1])), 
                                     database='reddit',
                                     neighbourhood='time')
-
+res <- res.seq.dyn
 #res <- res.seq
 #plot.motif.counts(res.seq)
 
@@ -89,28 +94,34 @@ res.seq.dyn <- count_motifs_by_post(threads[1:50],
 ncores <- detectCores() - 2
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
-pck <- c('RSQLite', 'data.table')
+pck <- c('RSQLite', 'data.table', 'changepoint')
 res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   source('R/extract_from_db.r')
+  #withTimeout(  count_motifs_by_post(chunks[[i]], 
+  #                                   database='reddit',
+  #                                   neighbourhood='time'),
+  #                                  120, onTimeout='warning')
[…]

### 2016-04-06 sandbox/illustration_breakpoints.r

@@ -1,0 +1,54 @@
+library(igraph)
+empty_graph()
+
+g <- make_empty_graph(n=23)
+g <- add_edges(g, c(2,1,
+                    3,2,
+                    6,2,
+                    5,2,
+                    4,2,
+                    7,3,
+                    8,3,
+                    8,9,
+                    10,6,
+                    11,10,
+                    12,10,
+                    14,10,
+                    15,5,
+                    13,10,
+                    17,10,
+                    18,15,
+                    16,15,
+                    19,2,
+                    20,2,
+                    21,2,
+                    22,12,
+                    23,22)) 
+
+eg.id <- 6
+breakpoints.v <- c(2,12)
+breakpoints.h <- c(4,14,20)
+deleted <- c(1,3,7,8,9,12,22,23,17,20,21)
+V(g)$color <- 'white'
+V(g)$color[eg.id] <- "grey"
+V(g)[breakpoints.v]$color <- 'orange'
+V(g)[breakpoints.h]$color <- 'yellow'
+#V(g)[deleted]$color <- 'red'
+par(mfrow=c(1,1))
+g <- as.undirected(g)
+la = layout_as_tree(g, mode='out', root=1)#which.min(V(g)$date))
+V(g)$size <- 6
+
+frame.colors <- rep("black",22)
+frame.colors[deleted] <- "white" 
+
+plot(g,
+     layout = la, 
+     vertex.label = order(V(g)),
+     vertex.label = NA,
+     edge.width = 1, 
+     edge.arrow.size=0.02,
+     vertex.frame.color= frame.colors,
+     vertex.label.color='black',
+     asp=9/16,
+     margin=-0.15)

### 2016-04-08 R/clustering.r
post presentation SIG

@@ -1,10 +1,10 @@
-cluster <- function(features){
+cluster <- function(features, k){
     wss <- (nrow(features)-1)*sum(apply(features,2,var))
     for (i in 2:25){
       wss[i] <- sum(kmeans(features, i)$withinss)
     }
     plot(1:25, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")
-    fit <- kmeans(features, 3)
+    fit <- kmeans(features, k)
     z <- fit$cluster
     
     # relabel cluster by size (to keep color consistency between executions)
@@ -25,9 +25,9 @@ cluster <- function(features){
 #fit2 <-Mclust(df.scaled)
 #summary(fit2, parameters=TRUE)
 
-plot.clusters <- function(df.features, clusters, sizes, colors){
+plot.clusters <- function(features, clusters, sizes, colors){
   par(mfrow=c(1,1))
-  pca <- princomp(df.features)
+  pca <- princomp(features)
   
   # Colored PCA (base graphics)
   ######################################
@@ -37,7 +37,7 @@ plot.clusters <- function(df.features, clusters, sizes, colors){
        pch  = 19,
        xlab = "Dimension 1", ylab = "Dimension 2")
   text(pca$scores[,1], pca$scores[,2], 
-       labels = rownames(df.features), 
+       labels = rownames(features), 
        cex    = 0.7)
   title("Individual factor map (PCA)")
   dev.copy(png, paste0('2016-01-15-PCA.png'), width = 800, height = 800)
@@ -51,7 +51,7 @@ plot.clusters <- function(df.features, clusters, sizes, colors){
                  groups       = factor(z), 
                  ellipse      = TRUE, 
                  circle       = TRUE,
-                 labels       = rownames(df.features),
+                 labels       = rownames(features),
                  labels.size  = 3.5,
                  varname.size = 4) 
   p1 <- p1 + scale_color_manual(values = alpha(cluster.colors,0.75))
@@ -67,7 +67,7 @@ plot.clusters <- function(df.features, clusters, sizes, colors){
   # Boxplots
   #########################
   # Clusters profile
-  points <- as.data.frame(df.features)
+  points <- as.data.frame(features)
   points$cluster <- factor(z)
   points <- melt(points, id='cluster')
   p <- ggplot(points, aes(x=variable, y=value)) + 

### 2016-04-08 R/count_motifs.r

@@ -5,9 +5,9 @@ source('R/tree_changepoints.r')
 count_motifs_by_post <- function(threads, 
                                  database='reddit', 
                                  neighbourhood=c('order', 'time'),
-                                 rad=3,
+                                 rad=2,
                                  max.neighbors=4){
-
+  
   con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
   
   # For every thread, recontruct its graph. Then count 
@@ -22,13 +22,16 @@ count_motifs_by_post <- function(threads,
   nthreads <- length(threads)
   posts.motifs <- data.frame()
   for(i in 1:nthreads){
+    write.table(threads[i], file="tried_threads.csv", row.names=FALSE, col.names=FALSE, append=TRUE)
+    
+    
     #Extract the egos of every post
     g <- database.to.graph(threads[i], con, database)
     gp <- g$gp
     
     size <- vcount(gp)
     cat('\n', i, '/', nthreads, '/', threads[i], ' size: ', size)
-    if(size>100){
+    if(size>1000){
       cat("\n WARNING: skipping thread of size ", size)
       next 
     }
@@ -98,9 +101,18 @@ count_motifs_by_post <- function(threads,
       posts.motifs <- rbindlist(list(posts.motifs, 
                                      data.frame(postid=post.id,motif=motif.id)))
     } # end egos
+    
+    # keep track of threads that gave no problems
+    write.table(threads[i], file="processed_threads.csv", row.names=FALSE, col.names=FALSE, append=TRUE)
   } # end threads
   
   posts.motifs <- data.frame(posts.motifs)
+  
+  if(nrow(posts.motifs)==0){
+    return(list(posts.motifs = NA,
+                motifs = NA))
+  }
+  
   names(posts.motifs) <- c('postid', 'motif')
   
   # Sort by frequency (and relabel: 1 for the most frequent and so forth)

### 2016-04-08 R/tree_changepoints.r

@@ -235,10 +235,10 @@ changepoints.cp <- function(gp, vertical=FALSE, horizontal=FALSE){
           bp <- cpts(cpt.meanvar(dates, class=TRUE, method="PELT"))+1
           breakpoints.h <- c(breakpoints.h, children[bp])
           
-          colors <- rep("black", length(dates))
-          colors[bp] <- 'red'
-          plot(dates, col=colors, pch=19, cex=0.5)
-          title("debug")
+          #colors <- rep("black", length(dates))
+          #colors[bp] <- 'red'
+          #plot(dates, col=colors, pch=19, cex=0.5)
+          #title("debug")
         }
       }
     }

### 2016-04-10 doc/2016-04 SIG Toulouse/presentation_SIG_Toulouse_Avril_2016.tex
neighbourhoods take more colors into account

\documentclass{beamer}

\usepackage{default}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}

\DeclareGraphicsExtensions{{.pdf},{.png},{.jpg}}
\graphicspath{ {./img/} }

\hypersetup{colorlinks,urlcolor=blue}
\AtBeginSection[]
{
	\begin{frame}
		\begin{NoHyper}
		\tableofcontents[currentsection]
		\end{NoHyper}
	\end{frame}
}
\begin{document}
\addtobeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}

\title{You are the way you (structurally) talk:  Structural-temporal neighbourhoods of posts to characterize users in online forums}
\author{Alberto Lumbreras \\Jouve B., Velcin J., Guégan, M.}
\date{April 8, 2016}
\maketitle

\begin{frame}\frametitle{Overview} 
\begin{NoHyper}
\tableofcontents
\end{NoHyper}
\end{frame}

\section{Introduction}
\subsection{The data}

\begin{frame}{The data}{Reddit. A forum of forums}
	\begin{figure}
		\centering
		\includegraphics[width=0.1\textwidth]{reddit-logo}	
	\end{figure}
	Download monthly dumps from:
	\href{http://couch.whatbox.ca:36975/reddit/comments/monthly/}{http://couch.whatbox.ca:36975/reddit/comments/monthly/}

	\vfill
	Extract forum of interest:\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}science}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}france}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}sociology}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}complexsystems}}\\
	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}podemos}} $\leftarrow$ in this presentation\\
	...
\end{frame}

\subsection{The graph representations of the data}

\begin{frame}{Graph representations}{Graph of user interactions (a social network)}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{sna}	
	\end{figure}
\end{frame}

\begin{frame}{Graph representations}{Trees of posts}
\begin{figure}
	\centering
	\includegraphics[width=1\textwidth]{forest}	
\end{figure}
\end{frame}

\begin{frame}{Graph representations}{Which one is better?}

\begin{itemize}
\item Depends on the task!
\item One might choose multiple representations (multi-level analysis)
\end{itemize}
\vfill
My choice: 
\begin{itemize}
\item Mostly tree representation
\item Because it explicitly represents discussions (and their evolution).
\end{itemize}
\begin{figure}
	\centering
	\includegraphics[width=0.3\textwidth]{tree1}	
	\includegraphics[width=0.3\textwidth]{tree2}	
	\includegraphics[width=0.3\textwidth]{tree4}	
\end{figure}
And sometimes:
\begin{itemize}
\item SNA representation of single conversations.
\end{itemize}
\end{frame}

\section{Structures of conversations}
\subsection{Basic idea}
\begin{frame}{Intuition}
\textit{Hypothesis}: different individuals have tendency towards different types of conversations and these types are reflected in the structure of their interactions.
\vfill
These conversational structures might be observed at two levels (at least):
\begin{itemize}
\item Social graph.
\item Posts graph (tree)
\end{itemize}
\end{frame}

\subsection{Triadic structures}
\begin{frame}{Triadic structures}{Triads are not enough}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{triads}
	\end{figure}

Triads in \textbf{trees of posts}:
	\begin{itemize}
		\item Only 3 possible triads (dyad, chain and star)
	\end{itemize}

Triads in  \textbf{social graph}:
	\begin{itemize}
		\item Order (therefore dynamic) is missing.
	\end{itemize}
\vfill
We need something richer that captures the dynamics of conversations.
\end{frame}

\subsection{Neighbourhood structures}
\begin{frame}{Order-based neighbourhoods}{Definition}
\begin{itemize}
	\item 1. Extract neighbourhood of post $i$ with radius $r$.
	\item 2. Keep only the $n$ posts that are closest (in time) to post $i$. 
\end{itemize}
	\begin{figure}
		\centering
		\includegraphics[width=01\textwidth]{order_neighbourhood}
	\end{figure}
Easy to compute, but how to choose the parameters $r,n$ ?  	
\end{frame}

\begin{frame}{Order-based neighbourhoods}{Census}
We found 129 different motifs:
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{neighbourhoods_2_4_1}
	\end{figure}
\end{frame}

\begin{frame}{Time-based neighbourhoods}{Definition}
\begin{itemize}
	\item 1. Extract neighbourhood of post $i$ with radius $r$.
	\item 2. Detect changes of speed (vertical/horizontal changepoints)
	\item 3. From $i$, get the posts around until a changepoint is found.
\end{itemize}
%\vspace{cm}  
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{breakpoints2}
	\end{figure}
\end{frame}

\begin{frame}{Time-based neighbourhoods}{Census}
We found 165 different motifs:
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{neighbourhoods_time_1}
	\end{figure}
\end{frame}

\section{Conversation-based clustering}
\begin{frame}{Methodology}
	\begin{itemize}
	\item Create a user $\times$ neighborhood matrix of counts.
	\item Z-normalize (users characterized by their deviation from the mean)
	\item Cluster!
\end{itemize}
\end{frame}

\begin{frame}{Conversation-based clustering}{Order-based}
	\begin{figure}
		\centering
		\includegraphics[width=1\textwidth]{PCA_cluster_order_based}
	\end{figure}	
\end{frame}

\section{Conclusions}
\begin{frame}{Conclusions}
	\begin{itemize}
	\item \textbf{Q: Can we use graph structure to characterise users?}
	\item A: Yes!
	\end{itemize}	
	\begin{itemize}
	\item \textbf{Q: By using triads}?
	\item A: No. They are not useful in trees.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: So, what kind of structure?}
	\item A: Posts neighbourhoods that are time/order sensitive.
	\end{itemize}
	\begin{itemize}
	\item \textbf{Q: What about language?}
	\item A: It's ok, but structure is more directly linked to thread dynamics (future work)
	\end{itemize}
	\textbf{Future work:}
	\begin{itemize}
		\item Prune time-based neighbourhoods to reduce dimensionality.
		\item Do users jump from cluster to cluster (paths of roles)
	\end{itemize}	
\end{frame}

\begin{frame}
\begin{center}
\Huge Merci !
	\begin{figure}
		\centering
		%\includegraphics[width=0.1\textwidth]{reddit-logo}
		%\includegraphics[width=0.2\textwidth]{reddit_french}	
		%\includegraphics[width=0.5\textwidth]{reddit_questions}
		\includegraphics[width=0.1\textwidth]{reddit_sans_culotte}
	\end{figure}
\end{center}
\end{frame}
\end{document}

### 2016-04-10 R/count_motifs.r

@@ -2,6 +2,56 @@ library(RSQLite)
 library(data.table)
 source('R/extract_from_db.r')
 source('R/tree_changepoints.r')
+
+
+prune <- function(motif){
+  # Arguments:
+  #   motif: a tree graph
+  # Returns:
+  #   a pruned motif
+  # 4 colors: 1(black) 2(red) 4(orange)
+  # we can only remove black nodes
+  # there can be only two black nodes together
+  # delete 
+  parents <- which(degree(motif, mode='in')>2)
+  to.delete <- c()
+  if(length(parents)==0){
+    return(motif)
+  }
+  # If more than two siblings with no children, keep only the first two
+  # (the most ancient)
+  for (p in 1:length(parents)){
+    siblings <- neighbors(motif, parents[p], mode='in')
+    siblings <- siblings[order(V(motif)$date[siblings])]
+    siblings.indegree <- degree(motif, v=siblings, mode='in')
+    
+    if(length(siblings)<3){
+      next
+    }
+    x <- siblings$color
+    
+    counts <- 1
+    delete <- rep(FALSE, length(x))
+    for(i in 2:length(x)){
+      if(x[i] != x[i-1]){
+        counts <- 1
+      }
+      else if(siblings.indegree[i]>0){
+        counts <- 1
+      }
+      else{
+        counts <- counts + 1
+      }
+      if(counts>2){
+        delete[i] <- TRUE
+      }
+    }
+    to.delete <- union(to.delete, siblings[delete])
+  }
+  motif <- delete.vertices(motif, to.delete)
+  motif
+}
+
 count_motifs_by_post <- function(threads, 
                                  database='reddit', 
                                  neighbourhood=c('order', 'time'),
@@ -9,8 +59,8 @@ count_motifs_by_post <- function(threads,
                                  max.neighbors=4){
   
   con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
-  
-  # For every thread, recontruct its graph. Then count 
+
+    # For every thread, recontruct its graph. Then count 
   # if neighborhood at distance 1 and binary tree, max nodes is 4
   # if neighborhood at distance 1 and binary tree, max nodes is 10
   # 4   10   22   30   46   62   93  123...
@@ -19,91 +69,119 @@ count_motifs_by_post <- function(threads,
   posts.motifs.id <- vector()
[…]

### 2016-04-10 pipeline_clustering.r

@@ -46,7 +46,7 @@ df.posts <- data.frame(df.posts) %>% arrange(date)
 #df.posts <- df.posts[80000:82500,] # 
 
 #df.posts <- df.posts[80000:85000,] # ok
-#df.posts <- df.posts[1:5000,] # # 15 mins
+df.posts <- df.posts[1:5000,] # # 15 mins
 #df.posts <- df.posts[1:25000,] # in progress # 15 mins #fail!
 df.posts <- df.posts[1:75000,] 
 #60
@@ -87,7 +87,11 @@ plot(cumsum(table(df.users$posts)), pch=19, cex=0.5,
 title('Users posts (cumulative)')
 
 # Compute neighborhood around every post
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/800))
+
+# Only long threads
+#df.threads <- filter(df.threads, length>10)
+
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/100))
 length(chunks)
 ##############################"
 # sequential
@@ -95,16 +99,17 @@ length(chunks)
 res.seq <- count_motifs_by_post(as.vector(unlist(chunks[1])), 
                                 database='reddit',
                                 neighbourhood='order')
-threads <- chunks[[1]]
 #"t3_2a5f81" problematic
 #t3_2aga0t
-res.seq.dyn <- count_motifs_by_post("t3_2a5f81", 
+
+res.seq.dyn <- count_motifs_by_post('t3_29yrey', 
                                     database='reddit',
                                     neighbourhood='time')
 
-res.seq.dyn <- count_motifs_by_post(as.vector(unlist(chunks)), 
-                                    database='reddit',
-                                    neighbourhood='time')
+library('lineprof')
+l <- lineprof(count_motifs_by_post(unlist(chunks)[1:20], 
+                                   database='reddit',
+                                   neighbourhood='time'))
 res <- res.seq.dyn
 #res <- res.seq
 #plot.motif.counts(res.seq)
@@ -202,7 +207,7 @@ features <- features[,-idx.motifs.delete]
 
 
 par(mfrow=c(1,1))
-z <- cluster(features, 5)
+z <- cluster(features, 3)
 
 # Update users df with their cluster
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   

### 2016-04-10 sandbox/SO_prunning.r

@@ -1,0 +1,141 @@
+prune <- function(motif){
+  # Arguments:
+  #   motif: a tree graph
+  # Returns:
+  #   a pruned motif
+  # 4 colors: 1(black) 2(red) 4(orange)
+  # we can only remove black nodes
+  # there can be only two black nodes together
+  # delete 
+  parents <- which(degree(motif, mode='in')>2)
+  to.delete <- c()
+  if(length(parents)==0){
+    return(motif)
+  }
+  # If more than two siblings with no children, keep only the first two
+  # (the most ancient)
+  for (p in 1:length(parents)){
+    siblings <- neighbors(motif, parents[p], mode='in')
+    siblings <- siblings[order(V(motif)$date[siblings])]
+    siblings.indegree <- degree(motif, v=siblings, mode='in')
+    
+    if(length(siblings)<3){
+      next
+    }
+    x <- siblings$color
+    
+    counts <- 1
+    delete <- rep(FALSE, length(x))
+    for(i in 2:length(x)){
+      if(x[i] != x[i-1]){
+        counts <- 1
+      }
+      else if(siblings.indegree[i]>0){
+        counts <- 1
+      }
+      else{
+        counts <- counts + 1
+      }
+      cat('\n id:', as.numeric(siblings[i]), '--', counts)
+      if(counts>2){
+        delete[i] <- TRUE
+      }
+    }
+    to.delete <- union(to.delete, siblings[delete])
+  }
+  motif <- delete.vertices(motif, to.delete)
+  motif
+}
+
+i <- 28
+
+par(mfrow=c(4,2))
+par(mfrow=c(1,2))
+
+#for (i in 1:length(motifs)){
+
+#for (i in 1:100){  
+  
+  gmotif <- motifs[[i]]
+  if(!any(degree(gmotif, mode='in')>1)){
+    next
+  }
+  mypalette <- c("grey", "red", "white", 'orange')
+  if(!any(degree(gmotif, mode='in')>3)){
+    next
+  }
+  #root <- V(gmotif)[which(degree(gmotif, mode='out')==0)]
+  
+  gmotif <- as.undirected(gmotif)
+  la = layout_as_tree(gmotif, mode='out', root=which.min(as.numeric(V(gmotif)$date)))
+  plot(gmotif,
+       layout = la,
+       vertex.color=mypalette[V(gmotif)$color],
+       vertex.label = V(gmotif),#"",
[…]

### 2016-04-10 sandbox/debug.r

@@ -1,0 +1,15 @@
+#t3_29yrey check logfile:  log8453.csv
+#CATCHED ERROR in thread:  t3_29b97c
+
+gp <- database.to.graph('t3_29yrey', con, database)$gp
+
+par(mfrow=c(1,1))
+g_ <- as.undirected(gp)
+root <- which.min(V(gp)$date)
+la = layout_as_tree(g_, mode='out', root=root)
+plot(g_,
+     layout = la,
+     vertex.color=mypalette[V(g_)$color],
+     vertex.size = 3,
+     vertex.label = "" ,
+     edge.arrow.size=0.6)

### 2016-04-10 sandbox/illustration_order_neighbourhood.r

@@ -1,0 +1,51 @@
+library(igraph)
+empty_graph()
+
+g <- make_empty_graph(n=23)
+g <- add_edges(g, c(2,1,
+                    3,2,
+                    6,2,
+                    5,2,
+                    4,2,
+                    7,3,
+                    8,3,
+                    8,9,
+                    10,6,
+                    11,10,
+                    12,10,
+                    14,10,
+                    15,5,
+                    13,10,
+                    17,10,
+                    18,15,
+                    16,15,
+                    19,2,
+                    20,2,
+                    21,2,
+                    22,12,
+                    23,22)) 
+
+eg.id <- 5
+deleted <- c(1,3,7,8,9,12,22,23,17,20,21)
+closest <- c(2,3,4,6,7,8)
+V(g)$color <- 'white'
+V(g)$color[eg.id] <- "grey"
+
+par(mfrow=c(1,1))
+g <- as.undirected(g)
+la = layout_as_tree(g, mode='out', root=1)#which.min(V(g)$date))
+V(g)$size <- 6
+
+frame.colors <-  rep("white",22) 
+frame.colors[closest] <-"black"
+
+plot(g,
+     layout = la, 
+     vertex.label = order(V(g)),
+     vertex.label = NA,
+     edge.width = 1, 
+     edge.arrow.size=0.02,
+     vertex.frame.color= frame.colors,
+     vertex.label.color='black',
+     asp=9/16,
+     margin=-0.15)

### 2016-04-11 pipeline_clustering.r
Experiments executed

@@ -46,7 +46,7 @@ df.posts <- data.frame(df.posts) %>% arrange(date)
 #df.posts <- df.posts[80000:82500,] # 
 
 #df.posts <- df.posts[80000:85000,] # ok
-df.posts <- df.posts[1:5000,] # # 15 mins
+#df.posts <- df.posts[1:5000,] # # 15 mins
 #df.posts <- df.posts[1:25000,] # in progress # 15 mins #fail!
 df.posts <- df.posts[1:75000,] 
 #60
@@ -91,7 +91,7 @@ title('Users posts (cumulative)')
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/100))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/300))
 length(chunks)
 ##############################"
 # sequential
@@ -115,7 +115,7 @@ res <- res.seq.dyn
 #plot.motif.counts(res.seq)
 
 # parallel
-ncores <- detectCores() - 2
+ncores <- detectCores() - 9
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
 pck <- c('RSQLite', 'data.table', 'changepoint')
@@ -127,7 +127,7 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   #                                  120, onTimeout='warning')
   count_motifs_by_post(chunks[[i]], 
                        database='reddit',
-                       neighbourhood='time')
+                       neighbourhood='order')
 }
 stopCluster(cl)
 
@@ -144,8 +144,11 @@ if(FALSE){
 }
 
 res <- merge.motif.counts(res.parallel)
-save(res,file="res_time_75000.Rda")
-#save(res, file='res_2_4_order_75000.Rda') 
+
+#save(res,file="res_time_75000.Rda")
+#load("res_time_75000.Rda")
+
+save(res, file='res_2_4_order_75000.Rda') 
 #load('res_2_4_order_75000.Rda') 
 
 # Plot found motifs and their frequency
@@ -195,7 +198,7 @@ features <- normalize_counts(user.motifs[active.mask,])
 # after removing the non-active, some motifs dissapear.
 # remove also motifs that appeared less than 10 times
 idx.motifs.none <- as.vector((which(is.na(colSums(features)))))
-idx.motifs.low <- as.vector((which(colSums(user.motifs[active.mask,])<10)))
[…]

### 2016-04-12 pipeline_clustering.r
plot user participations

@@ -231,3 +231,29 @@ plot.clusters(features,
               clusters = z,
               sizes    = 0.5*log(rowSums(user.motifs[active.mask,])), 
               colors   = alpha(cluster.colors[z], 0.5))
+
+
+####################################################
+# Which are the top neighbourhoods of each cluster?
+####################################################
+centers <- kmeans(features, 3)$centers
+plot(centers[1,])
+plot(centers[2,])
+plot(centers[3,])
+
+k <- 1
+fav.motifs <- order(abs(centers[k,]),  decreasing=TRUE)
+plot(centers[k,fav.motifs])
+cat(fav.motifs)
+
+k <- 2
+fav.motifs <- order(abs(centers[k,]),  decreasing=TRUE)
+plot(centers[k,fav.motifs])
+cat(fav.motifs)
+
+k <- 3
+fav.motifs <- order(abs(centers[k,]),  decreasing=TRUE)
+plot(centers[k,fav.motifs])
+cat(fav.motifs)
+
+## Plot which are those favorite clusters

### 2016-04-12 doc/IEEEtran/paper_IEEEformat.tex
cleaning

@@
 
+\section{Data}
+
 \section{Discussion trees}
@@
 
+\subsection{Neighbourhood pruning}\label{sec:pruning}
+
 \section{Application to Reddit forums}
@@
 
-\subsection{Podemos forum}
 Our dataset consists of all posts from March 2014 to May 2015 of the Podemos forum in Reddit\footnote{https://www.reddit.com/r/podemos}\footnote{I have also the data for \textit{gameofthrones, france, datascience, machinelearning complexsystems, philosophy, twoxchromosomes, trees, sex}}.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{threads_length}%
-\includegraphics[width=0.5\textwidth]{users_posts}
-\includegraphics[width=0.5\textwidth]{threads_length_cum}%
+\includegraphics[width=0.5\textwidth]{users_posts}%
 \includegraphics[width=0.5\textwidth]{users_posts_cum}
-\caption{Log-log and cumulative distributions and of thread lengths and posts per user.}
+\caption{Log-log and cumulative distributions and of posts per user.}
 \label{fig:podemos_distributions}
@@
  %\noindent\textcolor{red}{\rule{16cm}{1mm}}
-\subsection{Counting neighorhoods}
+\subsection{Neighbourhood census}
+In this section we analyse the neighbourhood census of our dataset.
+A counter for every neighbourhood structure is created the first time we detect them.
+For every post, we extract its neighbourhood (order-based or time-based) and prune it as explained above (Section~\ref{sec:pruning}).
+Then check whether it is isomorphic to some of the already seen neighbourhoods.
+If this is the case, then we increment the counter for the neighbourhood previously seen.
+Otherwise we create a new entry for the current neighbourhood.
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
+	%\includegraphics[width=0.8\textwidth]{census_orderbased_2}
+	%\includegraphics[width=0.8\textwidth]{census_orderbased_3}
+	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$}
+	\label{fig:census_orderbased}
+\end{figure*}
+
+The census obtained with both type of neighbourhood are similar for the most frequent structures. That means that both kind of neighbourhoods are isomorphic for the most popular discussion structures.
+
 In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighbourhood structures.
-
-If we set the a spatial radius $r=1$ and an order $n=4$, the temporal neighbourhood $N_{G}^{(t)}(i, 1, 4)$ recovers the four adjacent posts that were written temporally closest to a given post.
-Figure~\ref{fig:neighbourhoods_1_4} shows the four detected neighbourhoods and their frequency.
 
@@
 It is interesting to see different structures like chains or stars where the root node might or might no appear and the ego post is placed at different positions.
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=0.8\textwidth]{neighbourhoods_1_4}
-	\caption{neighbourhood counts with $r=1$ and $n=4$}
-	\label{fig:neighbourhoods_1_4}
-\end{figure*}
 
@@
 
-\subsection{Clustering}
+\subsection{Conversation-based clustering of users}
 
@@
 Figure~\ref{fig:PCA} shows a PCA projection of the users colored by cluster and the distribution of the clusters in every dimension.
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=1\textwidth]{PCA_cluster_timebased}
+	\includegraphics[width=1\textwidth]{PCA_cluster_orderbased}
+	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
+	\label{fig:PCA}
+\end{figure*}
 

### 2016-04-12 pipeline_clustering.r

@@ -49,7 +49,7 @@ df.posts <- data.frame(df.posts) %>% arrange(date)
 #df.posts <- df.posts[1:5000,] # # 15 mins
 #df.posts <- df.posts[1:25000,] # in progress # 15 mins #fail!
 df.posts <- df.posts[1:75000,] 
-#60
+
 
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
@@ -131,7 +131,6 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
 }
 stopCluster(cl)
 
-
 # Debug. get threads not still processed and create chunks to try again
 # until we find the thread that raises the exception
 if(FALSE){
@@ -149,7 +148,7 @@ res <- merge.motif.counts(res.parallel)
 #load("res_time_75000.Rda")
 
 save(res, file='res_2_4_order_75000.Rda') 
-#load('res_2_4_order_75000.Rda') 
+load('res_2_4_order_75000.Rda') 
 
 # Plot found motifs and their frequency
 plot.motif.counts(res)

### 2016-04-12 pipeline_clustering.r
cleaning

@@ -145,10 +145,10 @@ if(FALSE){
 res <- merge.motif.counts(res.parallel)
 
 #save(res,file="res_time_75000.Rda")
-#load("res_time_75000.Rda")
+load("res_time_75000.Rda")
 
-save(res, file='res_2_4_order_75000.Rda') 
-load('res_2_4_order_75000.Rda') 
+#save(res, file='res_2_4_order_75000.Rda') 
+#load('res_2_4_order_75000.Rda') 
 
 # Plot found motifs and their frequency
 plot.motif.counts(res)

### 2016-04-12 plot_user_participations.r

@@ -1,0 +1,82 @@
+# Shows the different conversations (neighbourhoods) in which a user participated
+# 
+# author: Alberto Lumbreras
+library(RSQLite)
+
+plot.user.participations <- function(thread, user, con, database='reddit'){
+  
+  g <- database.to.graph(thread, con, database)
+  gp <- g$gp
+  mask <- V(gp)$user==user
+  posts <- V(gp)[mask]
+  
+  for(i in 1:length(posts)){
+    eg <- neighborhood.temporal.order(gp, posts[i], 2, 4)
+    g <- graph.union(gp,eg)
+    
+    
+    # Color nodes
+    root.post <- V(gp)[which(degree(gp, mode='out')==0)]$name
+    user.name <- V(gp)[posts[i]]$user
+    post.id <- V(gp)[posts[i]]$name
+    
+    # See if it matches any seen motif
+    u <- V(gp)[V(gp)$name==post.id] # identify the ego vertex
+    uu <- V(gp)[V(gp)$user==user.name] # posts writen by ego author
+    
+    # note the <= in case two siblings have the same date
+    V(gp)$color[V(gp)$date<=V(gp)[u]$date] <- 1 
+    V(gp)$color[V(gp)$date>V(gp)[u]$date] <- 2 
+    if(degree(gp, v=u, mode='out')==1){
+      p.user.name <- neighbors(gp, u, mode='out')$user
+      V(gp)[V(gp)$user==p.user.name]$color <- 3 # posts writen by parent of ego
+    }
+    V(gp)[uu]$color <- 4
+    V(gp)[u]$color <- 5
+    V(gp)[V(eg)$name==root.post]$color <- 6
+    V(gp)[! V(gp)$name %in% V(eg)$name]$color <- 7
+    
+    mypalette <- c("grey", "black", "yellow", "orange", "red", "white", 'green')
+    
+    # Plot
+    gmotif <- as.undirected(gp)
+    la = layout_as_tree(gmotif, mode='out', root=which.min(V(gmotif)$date))
+    plot(gmotif,
+         layout = la,
+         vertex.color=mypalette[V(gmotif)$color],
+         vertex.label = order(V(gp)$date),
+         vertex.label.cex = 1,
+         vertex.size= ifelse(V(gp)$color==7, 1, 3),
+         edge.arrow.size=0.6)
+    title(paste(user))
+  }
+}
+
+
[…]

### 2016-04-12 doc/IEEEtran/paper_IEEEformat.tex
paper ++

@@
 
-\section{Data}
-
 \section{Discussion trees}
@@
 
-The census obtained with both type of neighbourhood are similar for the most frequent structures. That means that both kind of neighbourhoods are isomorphic for the most popular discussion structures.
+The census obtained with both type of neighbourhood are similar for the most frequent structures? That means that both kind of neighbourhoods are isomorphic for the most popular discussion structures.
 
-In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighbourhood structures.
-
-In order to detect more complex types of neighbourhoods we enlarge the spatial radius to $r=4$.
-Figure~\ref{fig:neighbourhoods_4_4} shows the eight detected neighbors and their frequency.
-It is interesting to see different structures like chains or stars where the root node might or might no appear and the ego post is placed at different positions.
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_1}
-	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_2}
-	\includegraphics[width=0.8\textwidth]{neighbourhoods_3_4_3}
-	\caption{Order-based neighbourhoods with $r=3$ and $n=4$}
-	\label{fig:neighbourhoods_4_4}
-\end{figure*}
+%In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighbourhood structures.
 
@@
 	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
-	\caption{Time-based neighbourhoods (most frequent)}
+	%\caption{Time-based neighbourhoods (most frequent)}
 	\label{fig:neighbourhoods_time}
@@
 
-Note that some neighbours are still very similar.
-For instance, \textit{common answer 3} is very close to \textit{common answer 2} in the sense that they both represent the ego post replying to a root post that has one and two replies respectively.
-It seems a good idea to merge them both since they seem to represent the same type of discussion.
+\subsection{Conversation-based clustering of users}
+In this section we categorize users based on the neighbourhoods of their posts.
+We limit our study to a set of 100 \textit{active users} who wrote more than 100 posts.
+For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the census, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$.
+We drop those feature columns that are zero for every user (these are neighbourhoods seen only around posts of users with low-level activity).
+To make the feature vector of a user independent on the number of posts, we transform the counts into percentages.
+And since some features have much higher percentages than others for most users, we scale and normalize the matrix so that every feature has mean 0 and variance 1.
+To avoid non-significant scores, we remove also the feature column corresponding to neighbourhoods that have a frequency less than 50 among the active users.
 
-\subsection{Conversation-based clustering of users}
+We use k-means to find the clusters, though one can use any other clustering method.
+Since the plot over the Within-Cluster Sum of Squares for $k=1,...20$ did not show any clear elbow we chose $k=3$ clusters so that clusters are easier to interpret.
+We did the clustering over a feature matrix with order-based neighbourhoods and another feature matrix with time-based neighbourhoods.
+Figures~\ref{fig:PCA} and shows the PCA projections of the users coloured by their assigned cluster.
 
-Given the neighbourhoods in which each user participates, we will analyze whether there exists different types of users or all users look similar.
-
-We want to make the analyze independent of the number of posts, and for that we normalize each user feature vector so that features indicate the percentage of posts in this kind of neighbourhood.
-Moreover, some neighbourhoods are much more common than others due to the nature of the forums.
-Thus, we normalize and scale the features so that every feature has a global mean 0 and variance 1.
-User features now represent z-scores, that is, how many standard deviations is this user feature away from the mean.
-
-We use a simple k-means to find the clusters.
-To decide the number of clusters, we run k-means for k=2,...,25 clusters and look at the Within-Cluster Sum of Squares (Figure~\ref{fig:elbow}) and we chose $k=5$ so that the results are more interpretable.
-Figure~\ref{fig:PCA} shows a PCA projection of the users colored by cluster and the distribution of the clusters in every dimension.
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=1\textwidth]{PCA_cluster_timebased}
-	\includegraphics[width=1\textwidth]{PCA_cluster_orderbased}
-	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
-	\label{fig:PCA}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=1\textwidth]{PCA_clustering_3_4_order}
-	\includegraphics[width=1\textwidth]{clustering_3_4_order}
-	\caption{PCA projection of the clusters found and cluster profile in every dimension.}
-	\label{fig:PCA}
-\end{figure*}
+%Interpretation
 
 \section{Conclusions}
+We presented a method to characterise conversations of users in online threads.
+Due to the tree nature of online threads, traditional patters such as triads are not able to capture much of relevant dynamics of a conversation.
+Our defined order-based and temporal-based neighbourhoods
+are able to capture a very rich variety of structures.
+We used this neighbourhoods to characterise users in terms of the structure of the conversations they participate in and showed that, indeed, there are different types of structural conversationalists
+
+%Future
+The concept of structural-temporal neighbourhood opens the door to some interesting paths of research.
+One might wonder whether other pruning are more pertinent than the proposed here.
+Also, even after pruning some neighbourhoods suggest the same type of conversation, so a manual merge might be convenient.
+Maybe sociologists can help to merge the found neighbourhoods in a meaningful way.
 
 % use section* for acknowledgment
-\section*{Acknowledgment}
-
-The authors would like to thank...
+%\section*{Acknowledgment}
 

### 2016-04-12 R/clustering.r

@@ -51,7 +51,7 @@ plot.clusters <- function(features, clusters, sizes, colors){
                  groups       = factor(z), 
                  ellipse      = TRUE, 
                  circle       = TRUE,
-                 labels       = rownames(features),
+                 #labels       = rownames(features),
                  labels.size  = 3.5,
                  varname.size = 4) 
   p1 <- p1 + scale_color_manual(values = alpha(cluster.colors,0.75))

### 2016-04-13 doc/IEEEtran/paper_IEEEformat.tex
paper ++

@@
 % use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
+\IEEEspecialpapernotice{(Draft version 13/04/2016)}
 
@@
 % in the abstract
-\begin{abstract}
-The abstract goes here.
+\begin{abstract}Users of social networks are often characterised by extracting some relevant features from the graphs associated to that network.
+The structural dynamic of the conversation are usually forgotten due to a lack of proper tools.
+We present a purely graph-based method to cluster users in online forums.
 \end{abstract}
@@
 Macro dynamics are, for instance, the evolution of some global properties of the social graph such as its diameter, or its distribution degree.
-Micro dynamics are, for instance, the triadic motifs that represent local phenomena such as transitivity (friends of my friends are also my friends).
+Micro dynamics are, for instance, the triadic motifs that represent local phenomena such as transitivity (friends of my friends are also my friends) and that may explain some macro phenomena.
 
+The study of user behaviours has also brought the attention of researchers. 
 An interesting question in online communities is that concerning roles.
@@
 If studied from the macro, we can analyse the number of users, the percentage of replied post, its centrality in the network, and  so forth.
+
+In this paper we present two types of structural patterns adapted to discussion trees. These trees can represent e-mails, forums, etc. Here we will use forum data to illustrate out methodology.
 
@@
 We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
-Then we introduce two new definitions that takes into account the time to overcome some of the limitations of the structural neighbourhood.
-We will finally apply these time-based neighbourhoods to detect clusters of users than tend to appear in the same time-based neighbourhoods.
-
+Then we introduce two richer definitions that take into account the order and the time.
+To illustrate the kind of neighbourhoods we obtain in a real forum, we show the neighbourhood census in a Reddit forum.
+Finally, we apply both neighbourhood definitions to detect clusters of users that tend to appear in the same neighbourhoods.
+We close the paper with some suggestions of future research.
 \begin{figure}
@@
 \section{Discussion trees}
-We represent a conversation thread as a tree graph where vertex represent posts and edges represent replies from some post to another.
-The tree is rooted at the post who started the thread.
-Figure~\ref{fig:trees} shows some real examples of trees in a Reddit\footnote{www.reddit.com} forum.
+The interactions like those found in an online forum can be modelled, at least, by two sort of graphs.
+On the one hand, we can create a graph where vertices represent individuals and an edges represent an interaction between two individuals.
+At the individual level, this graph is suitable for a large set of analysis concerning the centrality of users, the detection of communities, the detection of groups that tend to interact with the same other groups.
+At the community level we can analyse properties such as the density of the graph, the clustering coefficient or the degree distribution.
+On the other hand, we can create a separate tree graph for every discussion where posts are represented by the vertices and and edge from one vertex to another indicates that first post is a reply to the second.
+This allows to analyse things like patterns of time delay between posts \cite{Bhatt2012}.
+
+We can even mix different levels of representations to study the relationship between the properties and dynamics among the different levels \cite{Dorat2007}.
+
+Of course, we are not limited to the study of graphs and some studies take into account the content of the conversations, the cross-posting activity of users, and so forth \cite{Whittaker1998}
 
@@
 
-\section{Structural neighbourhoods in discussion trees}
-Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, in this paper we define the \textit{structural neighbourhood} as follows:
+Others: \cite{Gaumont2016}
+
+\section{Structural neighbourhoods}
+Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, we start by the following definition of \textit{structural neighbourhood}:
 
@@
 % pros and cons
-This definition has two drawbacks when used in the context of conversation trees.
-%%% not all dynamics are captured
-First, the dynamics of the conversation (time or order in which posts are attached to the tree) are not entirely captured in the structure of a tree representation.
-We know, for instance, that the time in which a node was attached to the tree is always posterior to that of its parent.
-But it impossible to say, by just looking at the structure of the tree, the order in which a set of sibling posts replied to their common parent post.
-Thus, a single \textit{structural neighbourhood} may sometimes correspond to very different dynamics.
-%TODO: ejemplo artificial para demostrar que la definicion de vecindario estructural puede corresponder a dinamicas muy diferentes.
-%% unbounded size
-Second, a structural neighbourhood at a given radius $r$ has an unbounded number of posts, and therefore the space of possible neighbourhoods is infinite.
-This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent (see Figure \ref{fig:large_neighbourhood}).
+In the context of discussion threads, this definition has two limitations. First,
 
-\begin{figure}
+ the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time they were written.
+In conversations, time plays an important role, therefore this is an important limitation.
+Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite.
+This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent.
+
+\section{Structuro-temporal neighbourhoods}
+As we have discussed, structural neighbourhood falls short in the analysis of conversational structures.
+Instead, we propose two new definitions of neighbourhood that take into account order and time in which posts are written.
+
+\begin{figure*}
 \centering
-	\includegraphics[width=0.24\textwidth]{large_neighbourhood}
-	\includegraphics[width=0.24\textwidth]{small_neighbourhood}
-	\caption{The size of the neighbourhood with radius $r$ is unbounded.
-These two graph represent frequent neighbourhoods of a post (red) that replied to the root (white).
-However, the tree in the left corresponds to a very successful root while the three in the right has not brought the attention of too many users.}
-	\label{fig:large_neighbourhood}
-\end{figure}
-
-\section{Structuro-temporal neighbourhoods in discussion trees}
-In this section, we propose two new time-based definition of neighbourhoods.
-
+\includegraphics[width=0.45\textwidth]{order_neighbourhood}
+\includegraphics[width=0.45\textwidth]{breakpoints}
+\caption{Illustration of order-based (left) and time-based (right) neighbourhoods.
+Grey nodes represent the ego post.
+Nodes with circles are those in the neighbourhood.
+Numbers indicate the order of the post.
+The order-based neighbourhood has parameters $r=3$, $n=6$.
+In the time-based neighbourhood, horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.}
+\label{fig:cutpoints}
+\end{figure*}
 \subsection{Order-based}
@@
 First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are structurally and temporally close.
-Second, the size of the neighbourhood has an upper bound of $max(\mathcal{N}_i(r)), o)$.
+Second, the size of the neighbourhood has an upper bound of $max(\mathcal{N}_i(r)), n)$ thus making the space of possible neighbourhood structures finite.
 
 \subsection{Time-based}
-Our second definition uses the real timestamps of posts in order to decide where the boundaries of a neighbourhood are.
+Our second definition uses the real timestamps of posts in order to automatically decide where the boundaries of a given neighbourhood are.
 We might naively set fixed time-based boundaries for the neighbourhood by only including in the neighbourhood those posts whose timestamp $p_i$ is at distance less than $\tau$ from the ego post $|t_i-t_{ego}|<\tau$.
-However, the pace at which posts are added to the conversation may be very different between conversations, and even between different parts of the same conversation tree.
+However, the pace at which posts are added to the conversation may be very different between conversations, and even between different parts of the same conversation tree, making it difficult to set a fixed temporal radius $\tau$.
 Instead, we propose a definition based on the concept of $\textit{local dynamic}$.
 
-\begin{definition}
-The temporal neighbourhood $\mathcal{N}_{i}^T(r)$ is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the posts belong to the same local dynamic than the ego post.
-\end{definition}
-Our definition of \textit{local dynamic} is based on the detection of changepoints in the time stamps of the posts.
+Our definition of \textit{local dynamic} is based on the detection of changepoints in the time stamps of the posts (ref).
 Given a sequence of consecutive posts in the same branch $b$ of a tree, denoted as $p_s,...,p_e$ we say that they belong to the same vertical dynamic if there is no (vertical) changepoint $p_i$ in $b$ such that $p_1 \prec p_i \preceq p_n$.
 Similarly, given a sequence of chronologically sorted  siblings $s$, denoted as $p_s,...,p_e$, we say that they belong to the same horizontal dynamic if there is no (horizontal) changepoint $p_i$ in $s$ such that $p_1 \prec p_i \preceq p_n$.
+
+\begin{definition}
+Given an ordered tree $T$, the temporal neighbourhood $\mathcal{N}_{i}^T(r)$ is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the posts belong to the same local dynamic than the ego post.
+\end{definition}
 Now we can outline the complete algorithm for time-based neighbourhood (Algorithm \ref{alg:temporal_neighbourhood}).
@@
 
-\begin{figure}
+\subsection{Colored neighbourhoods}
+Once we extract a neighbourhood, we assign colors to their nodes to represent some relevant characteristics.
+In particular, we assign the red color to the ego post, the orange colour to all other posts written by the author of the ego (to identify re-entries) a white color to the root (if included in the neighbourhood), the yellow color to all posts written by the same author who wrote the parent of ego (to identify, for instances, debates between the ego author and the other author), the grey color to all posts written before the ego and the black color to all posts written after the ego.
+Since some posts might correspond to several colors, we perform a sequential assignment of colors given by $\texttt{grey, black, orange, ego, yellow, white}$.
+
+\begin{figure*}
 \centering
-\includegraphics[width=0.45\textwidth]{breakpoints}
-\caption{Time-based neighbourhood.
-Horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.
-Post 6 is the ego, and posts with circular borders are those belonging to the time-based neighbourhood.}
+\includegraphics[width=0.2\textwidth]{neighbourhood_time_342}
+\includegraphics[width=0.19\textwidth]{neighbourhood_time_456}
+\includegraphics[width=0.18\textwidth]{neighbourhood_time_501}
+\includegraphics[width=0.18\textwidth]{neighbourhood_time_524}
+\caption{Illustration of different real structures using different colors for the root (white), the ego (red) and other posts written by the same author (orange) the ego parent and other posts written by the same author (yellow), posts previous to the ego (grey) and posts posterior to the ego.}
 \label{fig:cutpoints}
-\end{figure}
-
-\subsection{Colored neighbourhoods}
-In conversation trees, the root vertex plays an important role since it represents the post that opened the conversation and thus set the topic of the thread.
-Therefore, a reply to the root post is clearly different from a reply to any other post in the tree.
-We distinguish root posts from non-root posts by assigning root posts a unique color (white).
-We also assign a unique color to the ego post (red).
+\end{figure*}
 
 \subsection{Neighbourhood pruning}\label{sec:pruning}
+Although the two proposed definitions limit the neighbour space considerably, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes.
+Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and no children.
 
@@
 	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
-	%\includegraphics[width=0.8\textwidth]{census_orderbased_2}
-	%\includegraphics[width=0.8\textwidth]{census_orderbased_3}
+	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
+	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
 	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{neighbourhoods_time_1}
-	\includegraphics[width=0.8\textwidth]{neighbourhoods_time_2}
+	\includegraphics[width=0.8\textwidth]{census_timebased_1}
+	\includegraphics[width=0.8\textwidth]{census_timebased_2}
 	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
-	%\caption{Time-based neighbourhoods (most frequent)}
+	\caption{Most frequent time-based neighbourhoods}
 	\label{fig:neighbourhoods_time}
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.9\textwidth]{PCA_cluster_orderbased}
+	\includegraphics[width=0.9\textwidth]{PCA_cluster_timebased}
+	\caption{PCA projections of the order-based (above) and time-based (below) neighbourhood features}
+	\label{fig:neighbourhoods_time}
+\end{figure*}
+
 %Interpretation
+
+% Maybe mention that we can take for instance only the posts before the ego and use it to understand when when can recommend a post based on the structure.
 
@@
 % http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
+\bibliographystyle{IEEEtran}
 % argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
+\bibliography{bibliography}
 %
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+%\begin{thebibliography}{1}
 

### 2016-04-15 doc/IEEEtran/paper_IEEEformat.tex
paper ++

@@
 % affiliations
-%\author{\IEEEauthorblockN{Alberto Lumbreras \\ and Bertrand Jouve}
-%\IEEEauthorblockA{Technicolor\\
-%Georgia Institute of Technology\\
-%Atlanta, Georgia 30332--0250\\
-%Email: alberto.lumbreras@irit.fr}
-%\and
-%\IEEEauthorblockN{Bertrand Jouve}
-%\IEEEauthorblockA{Twentieth Century Fox\\
-%Springfield, USA\\
-%Email: jouve@univ-tlse2.fr}
-%\and
-%\IEEEauthorblockN{Marie Guégan}
-%\IEEEauthorblockA{Technicolor Academy\\
-%San Francisco, California 96678--2391\\
-%Telephone: (800) 555--1212\\
-%Fax: (888) 555--1212}
-%\and
-%\IEEEauthorblockN{Marie Guégan}
-%\IEEEauthorblockA{Technicolor Academy\\
-%San Francisco, California 96678--2391\\
-%Telephone: (800) 555--1212\\
-%Fax: (888) 555--1212}}
+\author{\IEEEauthorblockN{Alberto Lumbreras \\ Julien Velcin}
+\IEEEauthorblockA{Laboratoire ERIC\\ Université de Lyon, France\\
+alberto.lumbreras@univ-lyon2.fr\\
+julien.velcin@univ-lyon2.fr}
+\and
+\IEEEauthorblockN{Bertrand Jouve}
+\IEEEauthorblockA{FRAMESPA/IMT\\Université de Toulouse 2, France\\
+jouve@univ-tlse2.fr}
+\and
+\IEEEauthorblockN{Marie Guégan}
+\IEEEauthorblockA{Technicolor, France\\
+marie.guegan@technicolor.com}}
 
@@
 % 
-\author{\IEEEauthorblockN{Alberto Lumbreras\IEEEauthorrefmark{1},
-Bertrand Jouve\IEEEauthorrefmark{2},
-Marie Guégan\IEEEauthorrefmark{3}, 
-Julien Velcin\IEEEauthorrefmark{4}}
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Technicolor, France\\Email: alberto.lumbreras@gmail.com}
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Université de Toulouse; UT2; FRAMESPA/IMT,\\ 5 allée Antonio Machado, 31058 Toulouse, cedex 9\\
-Email: jouve@univ-tlse2.fr}
-\IEEEauthorblockA{\IEEEauthorrefmark{3}Technicolor\\975 Avenue des Champs Blancs\\35576 Cesson-Sevigné,\\France\\Email: marie.guegan@technicolor.com}
-\IEEEauthorblockA{\IEEEauthorrefmark{4}Laboratoire ERIC, Université de Lyon,\\ 5 avenue Pierre Mendès France, 69676, Bron\\France\\Email: julien.velcin@univ-lyon2.fr}}
+%\author{\IEEEauthorblockN{Alberto Lumbreras\IEEEauthorrefmark{1},
+%Bertrand Jouve\IEEEauthorrefmark{2},
+%Marie Guégan\IEEEauthorrefmark{3}, 
+%Julien Velcin\IEEEauthorrefmark{4}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}Technicolor, France\\Email: alberto.lumbreras@gmail.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Université de Toulouse; UT2; FRAMESPA/IMT,\\ 5 allée Antonio Machado, 31058 Toulouse, cedex 9\\
+%Email: jouve@univ-tlse2.fr}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}Technicolor\\975 Avenue des Champs Blancs\\35576 Cesson-Sevigné,\\France\\Email: marie.guegan@technicolor.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Laboratoire ERIC, Université de Lyon,\\ 5 avenue Pierre Mendès France, 69676, Bron\\France\\Email: julien.velcin@univ-lyon2.fr}}
 
 % use for special paper notices
-\IEEEspecialpapernotice{(Draft version 13/04/2016)}
+\IEEEspecialpapernotice{(Draft version: \today)}
 
@@
 \section{Introduction}
-The popularization of online forums has brought a growing interest on their underlying dynamics.
-As any other complex system, the dynamic of online forums can be studied at different levels, from the most macro to the most micro.
-Macro dynamics are, for instance, the evolution of some global properties of the social graph such as its diameter, or its distribution degree.
-Micro dynamics are, for instance, the triadic motifs that represent local phenomena such as transitivity (friends of my friends are also my friends) and that may explain some macro phenomena.
+The interactions between users in online forums are often modelled as complex networks.
+As such, they can be studied from many levels, each of which is represented by a graph where vertices and edges may represent different things.
+The most studied graph is a graph where vertices represent users and edges represent interactions between users (a post from one user replying to a post from other user).
+From the point of view of the community, some typically analysed properties are the degree distribution, the clustering coefficient, density, or diameter of the graph.
+From the point of view of the individual users, analyses are typically focused either on the centrality of users, on the community structure or in blockmodeling (groups of users that tend to interact with the same other groups) \cite{McCallum2007a}.
+Another common graph is a tree graph where vertices represent posts (or e-mails) and and edges from one vertex to another indicates that the first is a reply to the second.
+Given the different discussion trees of a forum (a forest) the global analyses include the depth distribution, branching factors, or even the time between two posts \cite{Bhatt2012}.
+representation include depth distribution of discussions, branching factors and so.
+Studying discussion trees from the point of view of the users means, since trees are discussions, studying the user at a conversational level.
+Unfortunately, there is a lack of structural tools to perform this analysis.
+In this paper, we propose to fill this gap through the concept of \textit{structural-temporal neighbourhood}.
+%El esquema debe ser;
+%- SNA: 
+%*** comunidad: degree distribution, density, clustering coefficient etc
+%*** usuario: centrality
+%- Tree/forest:
+%*** global properties: degree distribution, depth distribution, inter-message times,...
+%** Individual properties: ...
 
-The study of user behaviours has also brought the attention of researchers. 
-An interesting question in online communities is that concerning roles.
-In sociology, roles are generally seen as the set of expected behaviours that are attached  to a position in the community.
-Extrapolating the notion of role, some researchers have looked for roles in online forums.
-Some others have tried to detect the roles and the users who hold that roles.
+Our goal in this paper is two-fold: on the one hand, to illustrate how structural-temporal neighbourhoods can play the role of triads for conversation trees.
+On the other hand, to show how structural-temporal neighbourhoods can be used for the detection of different types of conversationalists in online forums or any other type of online discussion that is representable by a tree structure (e.g.: e-mails).
 
-Roles can also be studied from the macro or the micro perspective.
-If studied from the macro, we can analyse the number of users, the percentage of replied post, its centrality in the network, and  so forth.
-
-In this paper we present two types of structural patterns adapted to discussion trees. These trees can represent e-mails, forums, etc. Here we will use forum data to illustrate out methodology.
-
-In this paper, we focus on the analyse of roles at a micro level, and more specifically at the discussion level.
-We would like to answer the following question: are there different types of users in terms of the kind of conversation they participate in?
-
-Our intuition is that some users like participating in some kind of discussion rather than other.
+%TODO: llevar esto al final 
+%Our intuition is that some users like participating in some kind of discussion rather than other.
 Certainly, an analysis of the textual content will tell us much about a discussion.
@@
 We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
-Then we introduce two richer definitions that take into account the order and the time.
-To illustrate the kind of neighbourhoods we obtain in a real forum, we show the neighbourhood census in a Reddit forum.
+Then we introduce our two definitions of structural-temporal neighbourhoods.
+To illustrate the kind of neighbourhoods that we obtain in a real data, we analyse the neighbourhood census of a Reddit forum.
 Finally, we apply both neighbourhood definitions to detect clusters of users that tend to appear in the same neighbourhoods.
@@
 \section{Discussion trees}
-The interactions like those found in an online forum can be modelled, at least, by two sort of graphs.
+% Introduce el objeto y la nomenclatura
+We represent a discussion thread by a tree graph $G=(V,E)$ where $V$ is a set of $n$ vertices representing the posts (also called messages or comments) and $E$ is a set of $n-1$ directed edges that say which posts replied to which post.
+Vertices have three attributes namely order, time, and author.
+If for two given vertices $v_i,v_j$ there exists an edge  $e=(v_i, v_j) \in E$ then $v_j$ is called the \textit{parent} of $v_i$, denoted as $p(v_i)=v_j$.
+The \textit{root} of the tree is the only vertex with no parent, and corresponds to the post that starts the discussion.
+We say that two vertices $v_i,v_j$ are \textit{siblings} if and only if $p(v_i)=p(v_j)$.
+A vertex $v_i$ is an ancestor of a vertex $v_j$ and $v_j$ is descendant of $v_i$ if and only if $v_i$ is in the path from $v_j$ to the root.
+A \textit{leaf} is a post with no replies.
+A \textit{branch} is the path between a leaf and the root.
+Two vertices $v_i$ and $v_i$ are said to be \textit{neighbours} if either $p(v_i)=v_j$ or $p(v_j)=v_i$ or, in other words, if their distance in the undirected version of $G$ is one.
+
+%The interactions like those found in an online forum can be modelled, at least, by two sort of graphs.
 On the one hand, we can create a graph where vertices represent individuals and an edges represent an interaction between two individuals.
@@
 
-We can even mix different levels of representations to study the relationship between the properties and dynamics among the different levels \cite{Dorat2007}.
+%We can even mix different levels of representations to study the relationship between the properties and dynamics among the different levels \cite{Dorat2007}.
 
-Of course, we are not limited to the study of graphs and some studies take into account the content of the conversations, the cross-posting activity of users, and so forth \cite{Whittaker1998}
+%Of course, we are not limited to the study of graphs and some studies take into account the content of the conversations, the cross-posting activity of users, and so forth \cite{Whittaker1998}
 
-The \textit{root} post of the tree is the post that starts the discussion. It is the only post that has no parent. A \textit{leaf} is a post with no replies.
-A \textit{branch} of the tree is the shortest path between the root and some of the leafs. Two branches may share their first posts.
-
-Others: \cite{Gaumont2016}
+%Others: \cite{Gaumont2016}
 
@@
 \begin{definition}
-Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $d$ from post $i$.
+Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $r$ from post $i$.
 \end{definition}
 % pros and cons
-In the context of discussion threads, this definition has two limitations. First,
-
- the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time they were written.
+In the context of discussion threads, this definition has two limitations.
+First, the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time when they were written.
 In conversations, time plays an important role, therefore this is an important limitation.
@@
 
-\section{Structuro-temporal neighbourhoods}
+\section{Structural-temporal neighbourhoods}
 As we have discussed, structural neighbourhood falls short in the analysis of conversational structures.
-Instead, we propose two new definitions of neighbourhood that take into account order and time in which posts are written.
+Instead, we propose two new definitions of neighbourhood that take into account the order and time in which posts are written.
 
@@
 \begin{definition}
-Given a tree graph $G$, the \textit{structural-temporal neighbourhood} of distance $d$ of post $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closer to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
+Given a ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
 \end{definition}
-
+An example of order-based neighbourhood is given in Figure~\ref{fig:cutpoints} (left).
 This definition has two advantages over the \textit{structural neighbourhood}.
-First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are structurally and temporally close.
-Second, the size of the neighbourhood has an upper bound of $max(\mathcal{N}_i(r)), n)$ thus making the space of possible neighbourhood structures finite.
+First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are not only near to $i$ in the structure but also in time.
+Second, the size of the neighbourhood has an upper bound of $min(|\mathcal{N}_i(r))|, n)$ thus making the space of possible neighbourhood structures finite.
+The main limitation of this definition is that the obtained neighbourhoods might be very different with different choices of $r$ and $n$, and we have no \textit{a priori} criteria to chose the proper parameters other than making them small so that the neighbourhoods capture the local dynamic of the conversation around the post $i$.
+We will tackle this issue in the next definition.
 
 \subsection{Time-based}
-Our second definition uses the real timestamps of posts in order to automatically decide where the boundaries of a given neighbourhood are.
-We might naively set fixed time-based boundaries for the neighbourhood by only including in the neighbourhood those posts whose timestamp $p_i$ is at distance less than $\tau$ from the ego post $|t_i-t_{ego}|<\tau$.
-However, the pace at which posts are added to the conversation may be very different between conversations, and even between different parts of the same conversation tree, making it difficult to set a fixed temporal radius $\tau$.
-Instead, we propose a definition based on the concept of $\textit{local dynamic}$.
-
-Our definition of \textit{local dynamic} is based on the detection of changepoints in the time stamps of the posts (ref).
-Given a sequence of consecutive posts in the same branch $b$ of a tree, denoted as $p_s,...,p_e$ we say that they belong to the same vertical dynamic if there is no (vertical) changepoint $p_i$ in $b$ such that $p_1 \prec p_i \preceq p_n$.
-Similarly, given a sequence of chronologically sorted  siblings $s$, denoted as $p_s,...,p_e$, we say that they belong to the same horizontal dynamic if there is no (horizontal) changepoint $p_i$ in $s$ such that $p_1 \prec p_i \preceq p_n$.
-
+In a first attempt to take time into account, one might set fixed time-based boundaries for the neighbourhood and include only those posts whose timestamp $t_j$ is at distance less than $\tau$ from the ego post $i$, $|t_j-t_i|<\tau$.
+However, the pace at which posts are added to the conversation may be very different between conversation threads (and also within a thread) and we have no \textit{a priori} criteria for a proper choice of $\tau$.
+Rather than looking for a fixed time radius, we may look at changes in the pace how new posts are added to the thread.
+In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values.
+If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$.
+Now we can introduce our second definition of neighbourhood.
 \begin{definition}
-Given an ordered tree $T$, the temporal neighbourhood $\mathcal{N}_{i}^T(r)$ is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the posts belong to the same local dynamic than the ego post.
+Given a tree graph $G$, the \textit{time-based neighbourhood} of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
 \end{definition}
-Now we can outline the complete algorithm for time-based neighbourhood (Algorithm \ref{alg:temporal_neighbourhood}).
-Figure \ref{fig:cutpoints} illustrate a tree with a set of breakpoints and the temporal neighbourhood for a given node.
+An example of time-based neighbourhood is given in Figure~\ref{fig:cutpoints} (right).
+Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$.
+Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods according to the given definition.
 \begin{algorithm}[H]
@@
 
-\subsection{Colored neighbourhoods}
-Once we extract a neighbourhood, we assign colors to their nodes to represent some relevant characteristics.
+\subsection{Neighbourhood colouring}\label{sec:colouring}
+Even if the structure contains some important information about the type of conversation, there is still some ambiguity left, and two similar neighbourhoods can represent very different types of conversation.
+We can easily reduce this ambiguity by assigning colors, or labels, to vertices, identifying some relevant property of the post.
+
 In particular, we assign the red color to the ego post, the orange colour to all other posts written by the author of the ego (to identify re-entries) a white color to the root (if included in the neighbourhood), the yellow color to all posts written by the same author who wrote the parent of ego (to identify, for instances, debates between the ego author and the other author), the grey color to all posts written before the ego and the black color to all posts written after the ego.
@@
 \caption{Illustration of different real structures using different colors for the root (white), the ego (red) and other posts written by the same author (orange) the ego parent and other posts written by the same author (yellow), posts previous to the ego (grey) and posts posterior to the ego.}
-\label{fig:cutpoints}
+\label{fig:colors}
 \end{figure*}
@@
 \subsection{Neighbourhood pruning}\label{sec:pruning}
-Although the two proposed definitions limit the neighbour space considerably, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes.
-Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and no children.
+Although the two previous definitions reduce the neighbour space considerably, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes.
+Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and where neither of them has any children.
+
+Figure~\ref{fig:colors} shows some real examples, from our dataset, of time-based neighbourhoods after colouring and pruning.
 
 \section{Application to Reddit forums}
-In this section, we use our definition of structural-temporal neighbourhood to detect different types of users according to the neighbourhoods in which their posts are embedded.
-
-Our dataset consists of all posts from March 2014 to May 2015 of the Podemos forum in Reddit\footnote{https://www.reddit.com/r/podemos}\footnote{I have also the data for \textit{gameofthrones, france, datascience, machinelearning complexsystems, philosophy, twoxchromosomes, trees, sex}}.
+We applied our two definitions of structural-temporal neighbourhood (order-based and time-based) to a forum dataset.
+Our dataset consists of all posts from March 2014 to May 2015 of the Podemos forum in Reddit\footnote{https://www.reddit.com/r/podemos}.
 The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014.
 Nowadays, its members use it mainly to share and discuss about political news.
-The forum contains 83,6119 posts spread over 47,803 threads and written by 26,193 users. (see Figure~\ref{fig:podemos_distributions})
+The forum contains 83,6119 posts spread over 47,803 threads and written by 26,193 users.
 
@@
 \subsection{Neighbourhood census}
-In this section we analyse the neighbourhood census of our dataset.
-A counter for every neighbourhood structure is created the first time we detect them.
-For every post, we extract its neighbourhood (order-based or time-based) and prune it as explained above (Section~\ref{sec:pruning}).
-Then check whether it is isomorphic to some of the already seen neighbourhoods.
+Given a set of local structures and a graph under study, a census is a counting of how many times does every structure appear in the graph.
+In the context of online discussion forums, triad census in the graph of users interactions are often used (in these graphs, an edge exists from user $u$ to user $v$ if $u$ wrote a post reply to some of the posts written by $v$ in some thread.) \cite{Adamic2008, Lumbreras2013}.
+
+We analysed the order-based and time-based neighbourhood census of our dataset.
+A counter for every neighbourhood structure is created the first time it us detected.
+For every post, we extract its neighbourhood (order-based or time-based) colour and prune it as explained above (Section~\ref{sec:colouring} and \ref{sec:pruning}).
+Then we check whether it is isomorphic to some of the already seen neighbourhoods.
 If this is the case, then we increment the counter for the neighbourhood previously seen.
 Otherwise we create a new entry for the current neighbourhood.
-
 \begin{figure*}
@@
 \end{figure*}
-
-The census obtained with both type of neighbourhood are similar for the most frequent structures? That means that both kind of neighbourhoods are isomorphic for the most popular discussion structures.
-
-%In this section we will build a matrix $M$ for one of the forums of Reddit and then we will cluster users to find groups that have similar preferences over the different types of conversations represented by the neighbourhood structures.
-
 \begin{figure*}
@@
 
+Figure \ref{fig:census_orderbased} and \ref{fig:neighbourhoods_time} show the most frequent order-based and time-based neighbourhoods.
+
+% TODO: Explicar un poco los resultados y conclusiones del censo. Mostrar grafica de frecuencia de cada motif o cumsum, para ilustrar como de distribuido esta el censo.
+
 \subsection{Conversation-based clustering of users}
-In this section we categorize users based on the neighbourhoods of their posts.
+In this section, we cluster users based on the neighbourhoods of their posts.
 We limit our study to a set of 100 \textit{active users} who wrote more than 100 posts.
@@
 We did the clustering over a feature matrix with order-based neighbourhoods and another feature matrix with time-based neighbourhoods.
-Figures~\ref{fig:PCA} and shows the PCA projections of the users coloured by their assigned cluster.
+Figure~\ref{fig:PCA} and shows the PCA projections of the users coloured by their assigned cluster.
+
+% TODO: Paragraph to explain similarity between clusters assignments
+
+% TODO: Paragraph to explain which motifs are prefered for every cluster
+
+% TODO: paragraph discussing a visual/manual analysis of some users
 
@@
 	\caption{PCA projections of the order-based (above) and time-based (below) neighbourhood features}
-	\label{fig:neighbourhoods_time}
+	\label{fig:PCA}
 \end{figure*}

### 2016-04-17 doc/IEEEtran/paper_IEEEformat.tex
loaded gameofthroaded

@@
 
-Our goal in this paper is two-fold: on the one hand, to illustrate how structural-temporal neighbourhoods can play the role of triads for conversation trees.
+Our goal in this paper is two-fold: on the one hand, to illustrate how structural-temporal neighbourhoods can play the role of triads for conversation trees, in the sense that they show us the local structures or dynamics from which the bigger graph emerges.
 On the other hand, to show how structural-temporal neighbourhoods can be used for the detection of different types of conversationalists in online forums or any other type of online discussion that is representable by a tree structure (e.g.: e-mails).
-
-%TODO: llevar esto al final 
-%Our intuition is that some users like participating in some kind of discussion rather than other.
-Certainly, an analysis of the textual content will tell us much about a discussion.
-However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humour, irony, or changing contexts, we turn our attention towards the structure of the discussions.
-More precisely, we analyse the local graph structure in which a user post in embedded, in the hope that this structure will be meaningful since it also reflects the kind of conversation in that part of the thread.
-Formally, these local graphs are known as neighbourhoods.
 
@@
 Nowadays, its members use it mainly to share and discuss about political news.
-The forum contains 83,6119 posts spread over 47,803 threads and written by 26,193 users.
+Our dataset contains contains 75,000 posts spread over 4,262 threads and written by 9,160 users between the April 25th and October 20th 2014.
 
@@
 
- %\noindent\textcolor{red}{\rule{16cm}{1mm}}
 \subsection{Neighbourhood census}
@@
 
-Figure \ref{fig:census_orderbased} and \ref{fig:neighbourhoods_time} show the most frequent order-based and time-based neighbourhoods.
+Figure \ref{fig:census_orderbased} and \ref{fig:neighbourhoods_time} show the most frequent order-based and time-based neighbourhoods in the Podemos dataset.
+In Figure X we compare the census of the most frequent neighbourhoods in the Podemos and the Game of Thrones forums.
 
@@
 In this section, we cluster users based on the neighbourhoods of their posts.
+Our intuition is that some users like participating in some kind of discussion rather than other.
+Certainly, most part of the information necessary to understand the nature of a discussion in in its textual content.
+However, due to the huge diversity of topics, vocabulary, and the difficulty of current algorithms to capture the language subtleties such as humour, irony, or context, we turn our attention towards the structure of the discussions, which can also contain some information.
+We work under the  hypothesis that the structural-neighbourhoods in which a user post in embedded reflect the kind of conversation in that part of the thread.
+
 We limit our study to a set of 100 \textit{active users} who wrote more than 100 posts.
@@
 %Interpretation
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth]{whiskers}
+	\caption{Whiskers plots in the most relevant features (time-based)}
+	\label{fig:PCA}
+\end{figure}
 

### 2016-04-17 R/load_participations.r

@@ -1,17 +1,20 @@
 library(RSQLite)
 library(data.table)
 source('R/extract_from_db.r')
-
+# SELECT p.postid, p.parent, p.user, p.date FROM posts p, threads t WHERE p.thread LIKE t.threadid AND t.forum LIKE 'podemos'
 load_posts <- function(database='reddit', forum='podemos'){
   con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
   
   # Threads in forum
-  thread.ids <- dbGetQuery(con,  paste0("select threadid from threads where forum='", forum, "'"))
+  thread.ids <- dbGetQuery(con,  paste0("SELECT threadid FROM threads WHERE forum='", forum, "'"))
+  
+  #threads <-  dbGetQuery(con,  paste0("SELECT t.threadid, p.date FROM threads t, posts p WHERE p.postid LIKE t.threadid  AND t.forum LIKE '", forum, "'"))
+  #threads$date <- as.numeric(threads$date)
+  #threads.ids <- arrange(threads, date)$threadid
   
   # Dataframe of users and threads where they participated
   df.posts <- data.frame()
   for(i in 1:nrow(thread.ids)){ 
-    #thread.posts <- dbGetQuery(con,  paste0("select user,date from posts where thread='", thread.ids[i,], "'"))
     thread.posts <- dbGetQuery(con,  paste0("SELECT postid, parent, user, date FROM posts WHERE thread='", thread.ids[i,], "'"))
     thread.posts$thread <- thread.ids[i,]
     

### 2016-04-17 pipeline_clustering.r

@@ -34,7 +34,14 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 ###################################################
 #df.posts <- load_posts(database='reddit', forum='podemos')
 #save(df.posts,file="dfposts.Rda")
-load('dfposts.Rda') # 836119 posts, 47803 threads
+#load('dfposts.Rda') # 836119 posts, 47803 threads
+
+df.posts <- load_posts(database='reddit', forum='gameofthrones')
+#save(df.posts,file="dfposts.Rda")
+
+
+# TODO: get N first threads by date
+
 df.posts <- data.frame(df.posts) %>% arrange(date)
 #df.posts <- df.posts[1:300000,]
 #df.posts <- df.posts[1:100000,]
@@ -48,7 +55,10 @@ df.posts <- data.frame(df.posts) %>% arrange(date)
 #df.posts <- df.posts[80000:85000,] # ok
 #df.posts <- df.posts[1:5000,] # # 15 mins
 #df.posts <- df.posts[1:25000,] # in progress # 15 mins #fail!
-df.posts <- df.posts[1:75000,] 
+
+df.posts <- df.posts[1:75000,] # Paper
+
+#df.posts <- df.posts[1:5000,] # Debug
 
 
 df.threads <- plyr::count(df.posts, "thread")
@@ -96,15 +106,13 @@ length(chunks)
 ##############################"
 # sequential
 #par(mfrow=c(1,1))
-res.seq <- count_motifs_by_post(as.vector(unlist(chunks[1])), 
+res.seq <- count_motifs_by_post(as.vector(unlist(chunks)), 
                                 database='reddit',
                                 neighbourhood='order')
-#"t3_2a5f81" problematic
-#t3_2aga0t
 
-res.seq.dyn <- count_motifs_by_post('t3_29yrey', 
-                                    database='reddit',
-                                    neighbourhood='time')
+
+
+
 
 library('lineprof')
 l <- lineprof(count_motifs_by_post(unlist(chunks)[1:20], 
@@ -115,7 +123,7 @@ res <- res.seq.dyn
 #plot.motif.counts(res.seq)
 
 # parallel
-ncores <- detectCores() - 9
+ncores <- detectCores() - 2
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
 pck <- c('RSQLite', 'data.table', 'changepoint')
@@ -145,13 +153,14 @@ if(FALSE){
[…]

### 2016-04-17 pipeline_compare_census.r

@@ -1,0 +1,17 @@
+# Compare the nieghbourhood census of two or more forums
+# author: Alberto Lumbreras
+
+df.posts <- load_posts(database='reddit', forum='gameofthrones')
+save(df.posts,file="dfposts_gameofthrones.Rda")
+
+df.posts <- load_posts(database='reddit', forum='sex')
+save(df.posts,file="dfposts_sex.Rda")
+
+df.posts <- load_posts(database='reddit', forum='4chan')
+save(df.posts,file="dfposts_4chan.Rda")
+
+df.posts <- load_posts(database='reddit', forum='complexsystems')
+save(df.posts,file="dfposts_complexsystems.Rda")
+
+df.posts <- load_posts(database='reddit', forum='datascience')
+save(df.posts,file="dfposts_datascience.Rda")

### 2016-04-18 pipeline_clustering.r
gameoftrhrones

@@ -38,28 +38,11 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 
 df.posts <- load_posts(database='reddit', forum='gameofthrones')
 #save(df.posts,file="dfposts.Rda")
-
-
-# TODO: get N first threads by date
-
+load('./R_objects/dfposts_gameofthrones.Rda')
+df.posts$date <- as.numeric(df.posts$date)
 df.posts <- data.frame(df.posts) %>% arrange(date)
-#df.posts <- df.posts[1:300000,]
-#df.posts <- df.posts[1:100000,]
-#df.posts <- df.posts[30000:60000,] # test debug
-#df.posts <- df.posts[50000:60000,] # test debug
-#df.posts <- df.posts[60000:100000,] # test debug error aqui
-#df.posts <- df.posts[70000:80000,] # test debug error aqui
-#df.posts <- df.posts[82500:85000,] # ok
-#df.posts <- df.posts[80000:82500,] # 
-
-#df.posts <- df.posts[80000:85000,] # ok
-#df.posts <- df.posts[1:5000,] # # 15 mins
-#df.posts <- df.posts[1:25000,] # in progress # 15 mins #fail!
-
 df.posts <- df.posts[1:75000,] # Paper
-
 #df.posts <- df.posts[1:5000,] # Debug
-
 
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
@@ -101,7 +84,7 @@ title('Users posts (cumulative)')
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/300))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/295))
 length(chunks)
 ##############################"
 # sequential
@@ -123,7 +106,7 @@ res <- res.seq.dyn
 #plot.motif.counts(res.seq)
 
 # parallel
-ncores <- detectCores() - 2
+ncores <- detectCores() - 5
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
 pck <- c('RSQLite', 'data.table', 'changepoint')
@@ -135,24 +118,13 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   #                                  120, onTimeout='warning')
   count_motifs_by_post(chunks[[i]], 
                        database='reddit',
[…]

### 2016-04-18 R/count_motifs.r
FIXED: re-indexing of motifs by frequency (minor bug)

@@ -208,8 +208,6 @@ merge.motif.counts.pair <- function(res1, res2){
 
   df.posts.motifs.1 <- res1$posts.motifs
   df.posts.motifs.2 <- res2$posts.motifs
-  #plot.motif.counts(res1)
-  #plot.motif.counts(res2)
   motifs.1 <- res1$motifs
   motifs.2 <- res2$motifs
   duplicated_motifs <- c()
@@ -239,8 +237,8 @@ merge.motif.counts.pair <- function(res1, res2){
   
   motifs.merged <- c(motifs.1, motifs.1.new)
   df.posts.motifs.2$motif <- mapping[df.posts.motifs.2$motif]  
-
   df.posts.motifs.merged <- rbind(df.posts.motifs.1, df.posts.motifs.2)
+  
   res <- list(posts.motifs = df.posts.motifs.merged,
               motifs = motifs.merged)
   
@@ -262,6 +260,14 @@ merge.motif.counts <- function(df.list){
   posts.motifs$motif <- match(posts.motifs$motif, idx)
   motifs <- motifs[idx]
   
+  # check this is sorted
+  n <- table(posts.motifs$motif[idx])
+  n <- table(posts.motifs$motif)
+  if(! all(n == cummin(n))){
+    cat('Something wrong in sorting motifs!')
+  }
+  
+  
   return(list(posts.motifs = posts.motifs,
               motifs = motifs))
 }
@@ -272,7 +278,6 @@ plot.motif.counts <- function(res){
   ####################################################
   # Plot found neighborhoods
   ####################################################
-  #mypalette <- c("black", "red", "white", 'orange')
   mypalette <- c("grey", "black", "yellow", "orange", "red", "white")
   par(mfrow=c(3,5))
   for(i in 1:length(motifs)){
@@ -283,7 +288,6 @@ plot.motif.counts <- function(res){
          vertex.color=mypalette[V(motifs[[i]])$color],
          vertex.label = "",
          edge.arrow.size=0.6)
-    #title(1, sub=sum(df.post.motif$motif==i))
     cat(sum(df.post.motif$motif==i))
     title(paste(i),sub=sum(df.post.motif$motif==i))  
   }

### 2016-04-18 R/load_participations.r

@@ -7,11 +7,7 @@ load_posts <- function(database='reddit', forum='podemos'){
   
   # Threads in forum
   thread.ids <- dbGetQuery(con,  paste0("SELECT threadid FROM threads WHERE forum='", forum, "'"))
-  
-  #threads <-  dbGetQuery(con,  paste0("SELECT t.threadid, p.date FROM threads t, posts p WHERE p.postid LIKE t.threadid  AND t.forum LIKE '", forum, "'"))
-  #threads$date <- as.numeric(threads$date)
-  #threads.ids <- arrange(threads, date)$threadid
-  
+
   # Dataframe of users and threads where they participated
   df.posts <- data.frame()
   for(i in 1:nrow(thread.ids)){ 

### 2016-04-18 pipeline_clustering.r

@@ -36,7 +36,7 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 #save(df.posts,file="dfposts.Rda")
 #load('dfposts.Rda') # 836119 posts, 47803 threads
 
-df.posts <- load_posts(database='reddit', forum='gameofthrones')
+#df.posts <- load_posts(database='reddit', forum='gameofthrones')
 #save(df.posts,file="dfposts.Rda")
 load('./R_objects/dfposts_gameofthrones.Rda')
 df.posts$date <- as.numeric(df.posts$date)
@@ -124,7 +124,7 @@ stopCluster(cl)
 
 res <- merge.motif.counts(res.parallel)
 
-save(res,file="res_time_75000_gameofthrones.Rda")
+#save(res,file="res_time_75000_gameofthrones.Rda")
 #load("res_time_75000.Rda")
 
 #save(res, file='res_2_4_order_75000.Rda') 
@@ -139,13 +139,32 @@ dev.off()
 
 
 #########################################
+#df.posts.bck <- df.posts
 
+df.posts <-df.posts.bck
 df.post.motif  <- res$posts.motifs
 motifs <- res$motifs
 
-# Add motif info to posts dataframe
-df.posts <- merge(df.posts, df.post.motif, all.x=FALSE, sort=FALSE)
+# Check they are sorted by frequency
+n <- as.numeric(table(df.post.motif$motif))
+all(n == cummin(n))
 
+# Add motif info to posts dataframe
+df.posts <- merge(df.posts, df.post.motif, all=FALSE, sort=FALSE)
+
+# Because the counting is based on threads, there might be some posts that were 
+# recovered in df.post.motif but that were cut in df.posts[1:N]
+# so re-sort again motifs by frequency
+# Sort by frequency (and relabel: 1 for the most frequent and so forth)
+# make NA appear in table so that they are convetred to 0
+idx <- order(tabulate(df.posts$motif), decreasing = TRUE)
+df.posts$motif <- match(df.posts$motif, idx)
+motifs <- motifs[idx]
+
+
+# Check they are STILL sorted by frequency (they should be)
+n <- as.numeric(table(df.posts$motif))
+all(n == cummin(n))
 
 ###############################################
 # Census
@@ -153,7 +172,8 @@ df.posts <- merge(df.posts, df.post.motif, all.x=FALSE, sort=FALSE)
[…]

### 2016-04-19 pipeline_clustering.r
processed gameofthrones order-based

@@ -118,7 +118,7 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   #                                  120, onTimeout='warning')
   count_motifs_by_post(chunks[[i]], 
                        database='reddit',
-                       neighbourhood='time')
+                       neighbourhood='order')
 }
 stopCluster(cl)
 
@@ -126,6 +126,9 @@ res <- merge.motif.counts(res.parallel)
 
 #save(res,file="res_time_75000_gameofthrones.Rda")
 #load("res_time_75000.Rda")
+
+#save(res,file="res_order_2_4_75000_gameofthrones.Rda")
+load("res_order_2_4_75000_gameofthrones.Rda")
 
 #save(res, file='res_2_4_order_75000.Rda') 
 #load('res_2_4_order_75000.Rda') 

### 2016-04-21 R/count_motifs.r
...

@@ -54,7 +54,7 @@ prune <- function(motif){
 
 count_motifs_by_post <- function(threads, 
                                  database='reddit', 
-                                 neighbourhood=c('order', 'time'),
+                                 neighbourhood=c('struct', 'order', 'time'),
                                  rad=2,
                                  max.neighbors=4){
   
@@ -126,6 +126,9 @@ count_motifs_by_post <- function(threads,
         }
         if(neighbourhood=='time'){
           eg <- neighborhood.temporal(gp, j, 2, breakpoints.v, breakpoints.h)
+        }
+        if(neighbourhood=='struct'){
+          eg <- make_ego_graph(gp, rad, nodes=j)[[1]]
         }
         
         # See if it matches any seen motif

### 2016-04-21 pipeline_clustering.r

@@ -79,31 +79,25 @@ plot(cumsum(table(df.users$posts)), pch=19, cex=0.5,
      ylab='Posts (cum)', xlab='User')
 title('Users posts (cumulative)')
 
+#########################################################
 # Compute neighborhood around every post
-
+#########################################################
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
 chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/300))
 length(chunks)
-##############################"
+
+# Profiling
+#library('lineprof')
+#l <- lineprof(count_motifs_by_post(unlist(chunks)[1:20], 
+#                                   database='reddit',
+#                                   neighbourhood='time'))
+
 # sequential
-#par(mfrow=c(1,1))
 res.seq <- count_motifs_by_post(as.vector(unlist(chunks)), 
                                 database='reddit',
                                 neighbourhood='order')
-
-
-
-
-
-library('lineprof')
-l <- lineprof(count_motifs_by_post(unlist(chunks)[1:20], 
-                                   database='reddit',
-                                   neighbourhood='time'))
-res <- res.seq.dyn
-#res <- res.seq
-#plot.motif.counts(res.seq)
 
 # parallel
 ncores <- detectCores() - 10
@@ -118,11 +112,12 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   #                                  120, onTimeout='warning')
   count_motifs_by_post(chunks[[i]], 
                        database='reddit',
-                       neighbourhood='time')
+                       neighbourhood='struct')
 }
 stopCluster(cl)
-
 res <- merge.motif.counts(res.parallel)
+save(res, file='./R_objects/res_struct_75000_podemos.Rda') 
+#save(res, file='./R_objects/res_order_2_4_75000_4chan.Rda') 
 
 #save(res,file="res_time_75000_4chan.Rda")
 #load("res_time_75000.Rda")
@@ -130,12 +125,11 @@ res <- merge.motif.counts(res.parallel)
 #save(res,file="res_order_2_4_75000_gameofthrones.Rda")
 #load("res_order_2_4_75000_gameofthrones.Rda")
 
-save(res, file='res_order_2_4_75000_podemos.Rda') 
 #load('res_2_4_order_75000.Rda') 
 
 # Plot found motifs and their frequency
 plot.motif.counts(res)
[…]

### 2016-04-21 doc/IEEEtran/paper_IEEEformat.tex
...

@@
 
+\newcommand\alberto[1]{\textcolor{black}{#1}}
+
 % *** Do not adjust lengths that control margins, column widths, etc. ***
@@
 Second, the size of the neighbourhood has an upper bound of $min(|\mathcal{N}_i(r))|, n)$ thus making the space of possible neighbourhood structures finite.
-The main limitation of this definition is that the obtained neighbourhoods might be very different with different choices of $r$ and $n$, and we have no \textit{a priori} criteria to chose the proper parameters other than making them small so that the neighbourhoods capture the local dynamic of the conversation around the post $i$.
+The main limitation of this definition is that the obtained neighbourhoods might be very different with different choices of $r$ and $n$, and we have no \textit{a priori} criteria to choose the proper parameters other than making them small so that the neighbourhoods capture the local dynamic of the conversation around the post $i$.
 We will tackle this issue in the next definition.
@@
 
-\subsection{Neighbourhood colouring}\label{sec:colouring}
+\section{Neighbourhood colouring and pruning}\label{sec:colouring_pruning}
 Even if the structure contains some important information about the type of conversation, there is still some ambiguity left, and two similar neighbourhoods can represent very different types of conversation.
@@
 
-\subsection{Neighbourhood pruning}\label{sec:pruning}
-Although the two previous definitions reduce the neighbour space considerably, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes.
+At this point, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes.
 Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and where neither of them has any children.
@@
 Our dataset contains contains 75,000 posts spread over 4,262 threads and written by 9,160 users between the April 25th and October 20th 2014.
+\subsection{Neighbourhood extraction}
+As a previous step for all further analysis we detect the neighbourhood around every post in our datasets.
+Each neighbourhood structure (including colours) is given a unique integer number.
+Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of of these trees would not be seen in the data.
+It appears then a much more practical approach to enumerate only those trees that are seen in the data.
+The consequence of this is that every dataset will have its own dictionary of trees.
+
+We detect the order-based and time-based neighbourhood of every post as follows.
+A counter for every neighbourhood structure is created the first time it us detected.
+For every post, we extract its neighbourhood (order-based or time-based) colour and prune it as explained above (Section~\ref{sec:colouring_pruning}).
+Then we check whether it is isomorphic to some of the already seen neighbourhoods.
+If this is the case, then we increment the counter for the neighbourhood previously seen.
+Otherwise we create a new entry for the current neighbourhood.
+The computational cost of this operation has a upper bound of $\mathcal O(n^2)$, $n$ being the number of posts, for the case where each neighbourhood occurs only once.
+\footnote{TODO: It's actually much less, probably logarithmic, since the size dictionary grows incrementally}
+
+\subsection{Agreement between neighbourhood methods}
+The detection of the time-based and order-based neighbourhood in the Game of Thrones dataset gave us a list of 2000 different (non-isomorphic) time-based neighbourhoods and 180 different (non-isomorphic) \footnote{TODO: Review numbers} order-based neighbourhoods, which means that the time-based is able to capture a much wider set of structures.
+Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times.
+Indeed, Figure \ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to account for the 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage.
+Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type.
+The reason why the frequency of the order-based neighbourhoods decay faster might be that the most frequent neighbourhood is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.
+
+The divergent number of neighbourhoods between order-based and times-based brings us to the following question:  how differ the order-based and time-based neighbourhoods given to a same post?
+We re-labelled the order-based neighbourhoods and assigned them the label of their isomorphic time-based counterparts or the first unused number in the time-based index.
+This left us a set of 2183 trees.
+Figure~\ref{fig:confusion} shows, for every post, its time-based neighbourhood and its order-based neighbourhood.
+Those in the diagonal (34\%) see no difference between the time-based and order-based.
+Out of the diagonal, there are two rows indicating a high number of posts being assign the same order-based neighbourhood but very different time-based neighbourhoods.
+These order-based neighbourhoods are the number 1 and the number 50.
+These collapse of many time-based neighbourhoods into one is significant because it happens a lot among very frequent time-based neighbourhood.
+In the right plot of Figure ~\ref{fig:confusion} we can see other order-based neighbourhoods that collapse many of the most frequent time-based neighbourhoods, such as the 22, 17, 26, 29, 49, 56, 47 and 101.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{users_posts}%
-\includegraphics[width=0.5\textwidth]{users_posts_cum}
-\caption{Log-log and cumulative distributions and of posts per user.}
-\label{fig:podemos_distributions}
+\includegraphics[width=0.5\textwidth]{confusion_time_order_gameofthrones}%
+\includegraphics[width=0.5\textwidth]{confusion_time_order_gameofthrones_zoom}
+\caption{Neighbourhood of posts from the point of view of order-based and time-based.
+Neighbourhoods with the same label in the order-based and the time-based axes are isomorphic.
+The points in the diagonal correspond to posts to whom the two methods assign exactly the same neighbourhood.
+Horizontal rows of points correspond to posts that are assigned many different time-based neighbourhoods but the same single order-based neighbourhood.}
+\label{fig:confusion}
 \end{figure*}
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth]{whiskers}
+	\caption{Whiskers plots in the most relevant features (time-based)}
+	\label{fig:PCA}
+\end{figure}
+
 \subsection{Neighbourhood census}
-Given a set of local structures and a graph under study, a census is a counting of how many times does every structure appear in the graph.
-In the context of online discussion forums, triad census in the graph of users interactions are often used (in these graphs, an edge exists from user $u$ to user $v$ if $u$ wrote a post reply to some of the posts written by $v$ in some thread.) \cite{Adamic2008, Lumbreras2013}.
+Given a set of local structures and a graph under study, a census is a counting of how many times does each structure appear in the graph.
+In the context of online discussion forums, triad census has been sometimes used in the graph of users interactions are often used (in these graphs, an edge exists from user $u$ to user $v$ if $u$ wrote a post reply to some of the posts written by $v$ in some thread.) \cite{Adamic2008, Lumbreras2013}.
+However, triads are poor objects when studying trees since there are only three possible structures with, at most, three nodes: a single vertex, a directed dyad (reply), star (two replies to the same post) and a chain (reply and reply to the reply).
+On the contrary, our neighbourhoods are potentially richer structures that capture the real structure of conversations.
 
-We analysed the order-based and time-based neighbourhood census of our dataset.
-A counter for every neighbourhood structure is created the first time it us detected.
-For every post, we extract its neighbourhood (order-based or time-based) colour and prune it as explained above (Section~\ref{sec:colouring} and \ref{sec:pruning}).
-Then we check whether it is isomorphic to some of the already seen neighbourhoods.
-If this is the case, then we increment the counter for the neighbourhood previously seen.
-Otherwise we create a new entry for the current neighbourhood.
 \begin{figure*}
 \centering
-	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
-	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
-	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
-	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$}
-	\label{fig:census_orderbased}
+\includegraphics[width=0.5\textwidth]{order_vs_time_census_distribution}%
+\includegraphics[width=0.5\textwidth]{order_vs_time_census_distribution_cum}
+\caption{Comparison of census distributions with order-based and time-based neighbourhoods.}
+\label{fig:census_distributions}
 \end{figure*}
+
+In Figure \ref{fig:census_compare} we compare the neighbourhood census in two different forums, Podemos (politics, news and discussion among party supporters) and Game of Thrones (discussion around the TV series).
+Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones.
+These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25).
+The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ...
+(27)
+
 \begin{figure*}
 \centering
-	\includegraphics[width=0.8\textwidth]{census_timebased_1}
-	\includegraphics[width=0.8\textwidth]{census_timebased_2}
-	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
-	\caption{Most frequent time-based neighbourhoods}
-	\label{fig:neighbourhoods_time}
+\includegraphics[width=0.5\textwidth]{time_neighbourhood_census_compare_forums}%
+\includegraphics[width=0.5\textwidth]{order_neighbourhood_census_compare_forums}
+\caption{Comparison of census for two different forums.}
+\label{fig:census_compare}
 \end{figure*}
 
-Figure \ref{fig:census_orderbased} and \ref{fig:neighbourhoods_time} show the most frequent order-based and time-based neighbourhoods in the Podemos dataset.
-In Figure X we compare the census of the most frequent neighbourhoods in the Podemos and the Game of Thrones forums.
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{census_size_vs_freq_gameofthrones}
+\caption{Size of neighbourhoods and their frequencies.}
+\label{fig:census_size_vs_freq}
+\end{figure}
 
-% TODO: Explicar un poco los resultados y conclusiones del censo. Mostrar grafica de frecuencia de cada motif o cumsum, para ilustrar como de distribuido esta el censo.
+The analysis followed in this section lead us to think that the time-based neighbourhood is more convenient than the order-based.
+The main reason, we think, is that order seems to be a too dogmatic way to decide where a neighbourhood ends.
 
@@
 %Interpretation
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth]{whiskers}
-	\caption{Whiskers plots in the most relevant features (time-based)}
-	\label{fig:PCA}
-\end{figure}
 
@@
 
-% references section
+\bibliographystyle{IEEEtran}
+\bibliography{bibliography}
+\newpage
+\section*{Appendix}
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
+	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
+	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
+	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$ (Podemos)}
+	\label{fig:census_orderbased}
+\end{figure*}
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.8\textwidth]{census_timebased_1}
+	\includegraphics[width=0.8\textwidth]{census_timebased_2}
+	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
+	\caption{Most frequent time-based neighbourhoods (Podemos)}
+	\label{fig:neighbourhoods_time}
+\end{figure*}
 
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{bibliography}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
+\begin{figure*}
+\centering
+\includegraphics[width=1\textwidth]{confusion_time_order_motifs_gameofthrones_1}
+\caption{Most popular time-based neighbourhoods (up) and how they are mostly detected in order-based neighbourhood (down).}
+\label{fig:census_size_vs_freq}
+\end{figure*}
 

### 2016-04-21 pipeline_compare_census.r

@@ -1,17 +1,406 @@
 # Compare the nieghbourhood census of two or more forums
 # author: Alberto Lumbreras
 
-df.posts <- load_posts(database='reddit', forum='gameofthrones')
-save(df.posts,file="dfposts_gameofthrones.Rda")
+library(dplyr)
+library(igraph)
+if(FALSE){
+  df.posts <- load_posts(database='reddit', forum='podemos')
+  save(df.posts,file="dfposts_podemos.Rda")
+  
+  df.posts <- load_posts(database='reddit', forum='gameofthrones')
+  save(df.posts,file="dfposts_gameofthrones.Rda")
+  
+  df.posts <- load_posts(database='reddit', forum='sex')
+  save(df.posts,file="dfposts_sex.Rda")
+  
+  df.posts <- load_posts(database='reddit', forum='4chan')
+  save(df.posts,file="dfposts_4chan.Rda")
+  
+  df.posts <- load_posts(database='reddit', forum='complexsystems')
+  save(df.posts,file="dfposts_complexsystems.Rda")
+  
+  df.posts <- load_posts(database='reddit', forum='datascience')
+  save(df.posts,file="dfposts_datascience.Rda")
+}
 
-df.posts <- load_posts(database='reddit', forum='sex')
-save(df.posts,file="dfposts_sex.Rda")
+# Load posts dataframe (df.posts)
+load('./R_objects/dfposts_gameofthrones.Rda')
+#load('./R_objects/dfposts_podemos.Rda')
+df.posts$date <- as.numeric(df.posts$date)
+df.posts <- data.frame(df.posts) %>% arrange(date)
+df.posts <- df.posts[1:75000,] # Paper
 
-df.posts <- load_posts(database='reddit', forum='4chan')
-save(df.posts,file="dfposts_4chan.Rda")
+df.threads <- plyr::count(df.posts, "thread")
+df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
+names(df.threads)[2] <- "length"
+names(df.users)[2] <- "posts"
 
-df.posts <- load_posts(database='reddit', forum='complexsystems')
-save(df.posts,file="dfposts_complexsystems.Rda")
 
-df.posts <- load_posts(database='reddit', forum='datascience')
-save(df.posts,file="dfposts_datascience.Rda")
+# Load the results of the neighbourhoud detection (res)
[…]

### 2016-04-21 sandbox/sorting.r

@@ -1,0 +1,23 @@
+motifs <- c('alberto', 'elisa', 'david', 'bertrand', 'julien')
+
+counts <- c('alberto', 'alberto', 'alberto', 'alberto',
+             'elisa',
+             'david', 'david', 'david',
+             'bertrand', 'bertrand', 'bertrand','bertrand',
+             'julien', 'julien')
+
+idx <- order(table(counts)[motifs], decreasing=TRUE)
+
+motifs <- motifs[idx]
+
+match(motifs, idx)
+
+#############"
+
+set.seed(3)
+x <- sample(100,1000, replace=TRUE)
+idx <- order(table(x), decreasing=TRUE)
+xx <- match(x, idx)
+n <- table(xx)
+n
+all(n == cummin(n))

### 2016-04-23 R/count_motifs.r
faster counting avoiding ineficient vertex searches in igraph. Also less colours in neighbourhood

@@ -56,7 +56,8 @@ count_motifs_by_post <- function(threads,
                                  database='reddit', 
                                  neighbourhood=c('struct', 'order', 'time'),
                                  rad=2,
-                                 max.neighbors=4){
+                                 max.neighbors=4,
+                                 chunk.id=0){
   
   con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
 
@@ -79,15 +80,15 @@ count_motifs_by_post <- function(threads,
     tryCatch({
       
       write.table(threads[i], file="tried_threads.csv", row.names=FALSE, col.names=FALSE, append=TRUE)
-      cat(threads[i], file=filelog, sep=",", append=TRUE)
+      #cat(threads[i], file=filelog, sep=",", append=TRUE)
       
       #Extract the egos of every post
       g <- database.to.graph(threads[i], con, database)
       gp <- g$gp
-      cat('graph', file=filelog, sep=",", append=TRUE)
+      #cat('graph', file=filelog, sep=",", append=TRUE)
       
       size <- vcount(gp)
-      cat('\n', i, '/', nthreads, '/', threads[i], ' size: ', size)
+      cat('\n', i, '/', nthreads, '/', threads[i], ' size: ', size, ' chunk.id: ', chunk.id )
       if(size>1000){
         cat("\n WARNING: skipping thread of size ", size)
         next 
@@ -99,7 +100,7 @@ count_motifs_by_post <- function(threads,
       V(gp)$date <- dates
       
       root.post <- V(gp)[which(degree(gp, mode='out')==0)]$name
-      cat('root', file=filelog, sep=",", append=TRUE)
+      #cat('root', file=filelog, sep=",", append=TRUE)
       
       if(neighbourhood=='time'){
         breakpoints <- changepoints.cp(gp, vertical=T, horizontal=T)
@@ -113,12 +114,22 @@ count_motifs_by_post <- function(threads,
         V(gp)[breakpoints.v]$bp.v <- TRUE
         V(gp)[breakpoints.h]$bp.h <- TRUE
         V(gp)[breakpoints.vh]$bp.vh <- TRUE
-        cat('bp', file=filelog, sep=",", append=TRUE)
+        #cat('bp', file=filelog, sep=",", append=TRUE)
       }
       
       for(j in 1:vcount(gp)){
-        user.name <- V(gp)[j]$user
-        post.id <- V(gp)[j]$name
+        if(j %% 25>1){
[…]

### 2016-04-23 pipeline_clustering.r

@@ -41,8 +41,8 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 load('./R_objects/dfposts_podemos.Rda')
 df.posts$date <- as.numeric(df.posts$date)
 df.posts <- data.frame(df.posts) %>% arrange(date)
-df.posts <- df.posts[1:75000,] # Paper
-#df.posts <- df.posts[1:5000,] # Debug
+#df.posts <- df.posts[1:75000,] # Paper
+df.posts <- df.posts[17100:17109,] # Debug
 
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
@@ -85,22 +85,24 @@ title('Users posts (cumulative)')
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/300))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/70))
 length(chunks)
 
 # Profiling
-#library('lineprof')
-#l <- lineprof(count_motifs_by_post(unlist(chunks)[1:20], 
-#                                   database='reddit',
-#                                   neighbourhood='time'))
+library(profvis)
+prof <- profvis({
+           res.seq <- count_motifs_by_post(as.vector(unlist(chunks))[1], 
+                                           database='reddit',
+                                           neighbourhood='struct')
+})
 
 # sequential
 res.seq <- count_motifs_by_post(as.vector(unlist(chunks)), 
                                 database='reddit',
-                                neighbourhood='order')
+                                neighbourhood='struct')
 
 # parallel
-ncores <- detectCores() - 10
+ncores <- detectCores() - 5
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
 pck <- c('RSQLite', 'data.table', 'changepoint')
@@ -112,11 +114,11 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   #                                  120, onTimeout='warning')
   count_motifs_by_post(chunks[[i]], 
                        database='reddit',
-                       neighbourhood='struct')
+                       neighbourhood='struct', chunk.id=i)
 }
 stopCluster(cl)
[…]

### 2016-04-23 pipeline_compare_census.r

@@ -380,7 +380,7 @@ df.confusions <- data.frame(motif.time = most.frequent.time,
 ####################################################
 # Plot selection of neighborhoods
 ####################################################
-mypalette <- c("grey", "black", "yellow", "orange", "red", "white")
+mypalette <- c("black", "yellow", "orange", "red", "white")
 par(mfcol=c(2,5))
 selected.motifs <- top.atractors
 for(i in 1:nrow(df.confusions)){

### 2016-04-23 doc/IEEEtran/paper_IEEEformat.tex
paper ++

@@
 \end{definition}
+For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by its own authors is available in the R library \texttt{changepoint}.}.
 An example of time-based neighbourhood is given in Figure~\ref{fig:cutpoints} (right).
@@
 \section{Application to Reddit forums}
-We applied our two definitions of structural-temporal neighbourhood (order-based and time-based) to a forum dataset.
-Our dataset consists of all posts from March 2014 to May 2015 of the Podemos forum in Reddit\footnote{https://www.reddit.com/r/podemos}.
+Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}.
 The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014.
 Nowadays, its members use it mainly to share and discuss about political news.
-Our dataset contains contains 75,000 posts spread over 4,262 threads and written by 9,160 users between the April 25th and October 20th 2014.
+The Game of Thrones is a casual discussion forum about the Game of Thrones TV series.
+The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between the April 25th and October 20th 2014.
+The Game of Thrones dataset contains 75,000 posts corresponding to XXX threads written by XXXXX users.
+
 \subsection{Neighbourhood extraction}
@@
 Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times.
-Indeed, Figure \ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to account for the 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage.
+Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to cover 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage.
 Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.5\textwidth]{whiskers}
-	\caption{Whiskers plots in the most relevant features (time-based)}
-	\label{fig:PCA}
-\end{figure}
-
 \subsection{Neighbourhood census}
@@
 
-In Figure \ref{fig:census_compare} we compare the neighbourhood census in two different forums, Podemos (politics, news and discussion among party supporters) and Game of Thrones (discussion around the TV series).
+In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums.
 Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones.
@@
 
-The analysis followed in this section lead us to think that the time-based neighbourhood is more convenient than the order-based.
+This results lead us to think that the time-based neighbourhood is more convenient than the order-based.
 The main reason, we think, is that order seems to be a too dogmatic way to decide where a neighbourhood ends.
@@
 Figure~\ref{fig:PCA} and shows the PCA projections of the users coloured by their assigned cluster.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.5\textwidth]{whiskers}
+	\caption{Whiskers plots in the most relevant features (time-based)}
+	\label{fig:PCA}
+\end{figure}
 

### 2016-04-23 R/count_motifs.r
fixed

@@ -118,7 +118,7 @@ count_motifs_by_post <- function(threads,
       }
       
       for(j in 1:vcount(gp)){
-        if(j %% 25>1){
+        if(j %% 25==0){
           cat('\n', i, '/', nthreads, '/', threads[i], ' size: ', size, ' chunk.id: ', chunk.id, " post: ", j )
           
         }
@@ -145,16 +145,17 @@ count_motifs_by_post <- function(threads,
         u <- V(eg)[post.id] # identify the ego vertex. Faster than V(eg)[V(eg)$name==post.id] 
         uu <- V(eg)[V(eg)$user==user.name] # posts writen by ego author
         
+        V(eg)$color <- 1
         # note the <= in case two siblings have the same date
-        V(eg)$color[V(eg)$date<=V(eg)[u]$date] <- 1 
-        V(eg)$color[V(eg)$date>V(eg)[u]$date] <- 2 
+        #V(eg)$color[V(eg)$date<=V(eg)[u]$date] <- 1 
+        #V(eg)$color[V(eg)$date>V(eg)[u]$date] <- 2 
         if(degree(eg, v=u, mode='out')==1){
           p.user.name <- neighbors(eg, u, mode='out')$user
-          V(eg)[V(eg)$user==p.user.name]$color <- 3 # posts writen by parent of ego
+          V(eg)[V(eg)$user==p.user.name]$color <- 2 # posts writen by parent of ego
         }
-        V(eg)[uu]$color <- 4
-        V(eg)[u]$color <- 5
-        tryCatch({V(eg)[root.post]$color <- 6}, error = function(e){}) # exception if root not in there
+        V(eg)[uu]$color <- 3
+        V(eg)[u]$color <- 4
+        tryCatch({V(eg)[root.post]$color <- 5}, error = function(e){}) # exception if root not in there
         
         # Prune neighbourhood to reduce number of possible neighbourhoods
         eg <- prune(eg)

### 2016-04-24 R/count_motifs.r
if neighbourhood after prunning is too large, skip

@@ -1,5 +1,7 @@
 library(RSQLite)
 library(data.table)
+library(R.utils)
+
 source('R/extract_from_db.r')
 source('R/tree_changepoints.r')
 
@@ -132,7 +134,6 @@ count_motifs_by_post <- function(threads,
         #post.id <- V(gp)[j]$name
         
         # Detect the neighborhood with the prefered method
-        cat('\n...get eg...')
         if(neighbourhood=='order'){
           eg <- neighborhood.temporal.order(gp, j, rad, max.neighbors)
         }
@@ -142,9 +143,7 @@ count_motifs_by_post <- function(threads,
         if(neighbourhood=='struct'){
           eg <- make_ego_graph(gp, rad, nodes=j)[[1]]
         }
-        cat('\n...eg size: ', vcount(eg) )
-        
-        cat('\n...colouring...')
+
         u <- V(eg)[post.id] # identify the ego vertex. Faster than V(eg)[V(eg)$name==post.id] 
         uu <- V(eg)[V(eg)$user==user.name] # posts writen by ego author
         
@@ -160,15 +159,17 @@ count_motifs_by_post <- function(threads,
         V(eg)[u]$color <- 4
         tryCatch({V(eg)[root.post]$color <- 5}, error = function(e){}) # exception if root not in there
         
-        cat('\n pruning...')
         # Prune neighbourhood to reduce number of possible neighbourhoods
         eg <- prune(eg)
-        cat("\nSize after pruning: ", vcount(eg))
-        
+
+        # If neighbourhood is too big, sometimes isomorphism check takes too long
+        if(vcount(eg)>25){
+          #cat("\nSkipping neighbourhood isomorphism in thread ", threads[i])
+          next
+        }
+
         # See if it matches any seen motif
         ###################################
-        cat('\nCheck if already seen')
-        cat('\n motifs indexed: ', length(motifs))
         is.new <- TRUE
         if(length(motifs)>0){
           for(motif.id in 1:length(motifs)){   
@@ -178,6 +179,14 @@ count_motifs_by_post <- function(threads,
             if(vcount(eg) != vcount(gmotif)){
               next
             }          
+
+            if(length(unique(V(eg)$color)) != length(unique(V(gmotif)$color))){
+              next
+            } 
+            
+            if(all(table(V(eg)$color) != table(V(gmotif)$color))){
+              next
+            }
             
[…]

### 2016-04-24 R/plotting.r

@@ -1,0 +1,20 @@
+#
+# Plotting functions
+#
+
+plot.tree <- function(gtree){
+  # Plots a tree graph
+  # Arguments:
+  #   gtree: a igraph object graph with no cycles (tree)
+
+  mypalette <- c("black", "yellow", "orange", "red", "white")
+  par(mfrow=c(1,1))
+  gtree.un <- as.undirected(gtree)
+  la = layout_as_tree(gtree.un, mode='out', root=which.min(V(gtree.un)$date))
+  plot(gtree.un,
+       layout = la,
+       vertex.color=mypalette[V(gtree.un)$color],
+       vertex.label = "",
+       vertex.size=1,
+       edge.arrow.size=0.6)
+}

### 2016-04-24 pipeline_clustering.r

@@ -1,30 +1,13 @@
-# Extract list of triads where a user participates.
-#
-# author: Alberto Lumbreras
-#
-#https://www.reddit.com/r/europe/CasualConversation
-#https://www.reddit.com/r/europe/
-#https://www.reddit.com/r/datascience
-#https://www.reddit.com/r/science/
-#https://www.reddit.com/r/france
-#https://www.reddit.com/r/catalunya
-#https://www.reddit.com/r/es
-library(parallel)
-library(doParallel)
-library(foreach)
+
 library(ggplot2)
 library(ggbiplot)
 library(gplots)
-
 library(dplyr)
 library(reshape2)
 library(igraph)
 library(ggbiplot)
-#library(RSQLite)
-#library(GGally)
 
 source('R/load_participations.r')
-source('R/count_motifs.r')
 source('R/normalize_counts.r')
 source('R/clustering.r')
 
@@ -32,12 +15,6 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 ###################################################
 # Load data
 ###################################################
-#df.posts <- load_posts(database='reddit', forum='podemos')
-#save(df.posts,file="dfposts.Rda")
-#load('dfposts.Rda') # 836119 posts, 47803 threads
-
-#df.posts <- load_posts(database='reddit', forum='gameofthrones')
-#save(df.posts,file="dfposts.Rda")
 load('./R_objects/dfposts_podemos.Rda')
 df.posts$date <- as.numeric(df.posts$date)
 df.posts <- data.frame(df.posts) %>% arrange(date)
@@ -79,69 +56,14 @@ plot(cumsum(table(df.users$posts)), pch=19, cex=0.5,
      ylab='Posts (cum)', xlab='User')
 title('Users posts (cumulative)')
 
-#########################################################
-# Compute neighborhood around every post
-#########################################################
-# Only long threads
-#df.threads <- filter(df.threads, length>10)
-
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/70))
-length(chunks)
-
-chunks <- "t3_2bmb4v"
-# Profiling
-library(profvis)
-prof <- profvis({
-  res.seq <- count_motifs_by_post(as.vector(unlist(chunks))[1], 
-                                  database='reddit',
-                                  neighbourhood='struct')
-})
-
-# sequential
-res.seq <- count_motifs_by_post(as.vector(unlist(chunks)), 
[…]

### 2016-04-24 pipeline_extract_neighbourhoods.r

@@ -1,0 +1,119 @@
+# Extract list of triads where a user participates.
+#
+# author: Alberto Lumbreras
+#
+#https://www.reddit.com/r/europe/CasualConversation
+#https://www.reddit.com/r/europe/
+#https://www.reddit.com/r/datascience
+#https://www.reddit.com/r/science/
+#https://www.reddit.com/r/france
+#https://www.reddit.com/r/catalunya
+#https://www.reddit.com/r/es
+library(parallel)
+library(doParallel)
+library(foreach)
+library(ggplot2)
+library(ggbiplot)
+library(gplots)
+
+library(dplyr)
+library(reshape2)
+library(igraph)
+library(ggbiplot)
+#library(RSQLite)
+#library(GGally)
+
+source('R/load_participations.r')
+source('R/count_motifs.r')
+source('R/normalize_counts.r')
+source('R/clustering.r')
+source('R/plotting.r')
+
+##########################################################
+# Load Data
+##########################################################
+load('./R_objects/dfposts_podemos.Rda')
+df.posts$date <- as.numeric(df.posts$date)
+df.posts <- data.frame(df.posts) %>% arrange(date)
+df.posts <- df.posts[1:75000,] # Paper
+#df.posts <- df.posts[17100:17109,] # Debug
+
+df.threads <- plyr::count(df.posts, "thread")
+df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
+names(df.threads)[2] <- "length"
+names(df.users)[2] <- "posts"
+
+# Print dates range
+start.date <- as.POSIXct(min(as.numeric(df.posts$date)), origin = "1970-01-01") 
+end.date <- as.POSIXct(max(as.numeric(df.posts$date)), origin = "1970-01-01")
+print(paste("Start date:", start.date))
+print(paste("End date:", end.date))
+
+cat("Nunmber of posts:", nrow(df.posts))
+cat('Number of threads: ', nrow(df.threads))
+cat('Number of users: ', nrow(df.users))
+cat('Number of active users', nrow(filter(df.users, posts>MIN_POSTS)))
+
+#########################################################
+# Compute neighborhood around every post
+#########################################################
+# Only long threads
+#df.threads <- filter(df.threads, length>10)
+
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/70))
+length(chunks)
+
[…]

### 2016-04-24 sandbox/timeout.r

@@ -1,0 +1,27 @@
+library(igraph)
+library(R.utils)
+
+
+
+foo1 <- function() {
+  print("Tic");
+  for (kk in 1:100) {
+    print(kk);
+    Sys.sleep(0.3);
+  }
+  print("Tac");
+}
+
+withTimeout(foo1(), timeout = 1, onTimeout = "warning")
+
+foo2 <- function() {
+  print("Tic");
+  N <- 10000
+  g <- erdos.renyi.game(N, 0.75)
+  V(g)$color <- sample(5,N, replace=TRUE)
+  is_isomorphic_to(g, g, method='vf2')
+  print("Tac");
+}
+
+withTimeout(foo2(), timeout = 1, onTimeout = "warning")
+

### 2016-04-26 pipeline_extract_neighbourhoods.r
more neighbourhoods extracted

@@ -60,7 +60,7 @@ cat('Number of active users', nrow(filter(df.users, posts>MIN_POSTS)))
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/70))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/235))
 length(chunks)
 
 if(FALSE){
@@ -82,7 +82,7 @@ res.seq <- count_motifs_by_post(as.vector(unlist(chunks)),
 }
 
 # parallel
-ncores <- detectCores() - 5
+ncores <- detectCores() - 6
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
 pck <- c('RSQLite', 'data.table', 'changepoint')
@@ -98,7 +98,13 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
 }
 stopCluster(cl)
 res <- merge.motif.counts(res.parallel)
-#save(res, file='./R_objects/res_struct_35000_podemos.Rda') 
+# save(res, file='./R_objects/res_time_75000_podemos.Rda') 
+# save(res, file='./R_objects/res_order_75000_podemos.Rda') 
+save(res, file='./R_objects/res_struct_75000_podemos.Rda') 
+# save(res, file='./R_objects/res_time_75000_gameofthrones.Rda') 
+# save(res, file='./R_objects/res_order_75000_gameofthrones.Rda') 
+# save(res, file='./R_objects/res_time_75000_4chan.Rda')
+# save(res, file='./R_objects/res_order_75000_4chan.Rda')
 
 #load("res_time_75000.Rda")
 

### 2016-04-27 pipeline_compare_census.r
paper experiments on census

@@ -2,6 +2,7 @@
 # author: Alberto Lumbreras
 
 library(dplyr)
+library(reshape2)
 library(igraph)
 if(FALSE){
   df.posts <- load_posts(database='reddit', forum='podemos')
@@ -23,94 +24,111 @@ if(FALSE){
   save(df.posts,file="dfposts_datascience.Rda")
 }
 
-# Load posts dataframe (df.posts)
+########################################################
+# Load posts
+########################################################
 load('./R_objects/dfposts_gameofthrones.Rda')
-#load('./R_objects/dfposts_podemos.Rda')
-df.posts$date <- as.numeric(df.posts$date)
-df.posts <- data.frame(df.posts) %>% arrange(date)
-df.posts <- df.posts[1:75000,] # Paper
+df.posts.got <- df.posts
+df.posts.got$date <- as.numeric(df.posts.got$date)
+df.posts.got <- data.frame(df.posts.got) %>% arrange(date)
+df.posts.got <- df.posts.got[1:75000,] # Paper
+df.threads.got <- plyr::count(df.posts.got, "thread")
+df.users.got <- plyr::count(df.posts.got, 'user')                                                                                                                                   
+names(df.threads.got)[2] <- "length"
+names(df.users.got)[2] <- "posts"
 
-df.threads <- plyr::count(df.posts, "thread")
-df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
-names(df.threads)[2] <- "length"
-names(df.users)[2] <- "posts"
+load('./R_objects/dfposts_podemos.Rda')
+df.posts.pod <- df.posts
+df.posts.pod$date <- as.numeric(df.posts.pod$date)
+df.posts.pod <- data.frame(df.posts.pod) %>% arrange(date)
+df.posts.pod <- df.posts.pod[1:75000,] # Paper
+df.threads.pod <- plyr::count(df.posts.pod, "thread")
+df.users.pod <- plyr::count(df.posts.pod, 'user')                                                                                                                                   
+names(df.threads.pod)[2] <- "length"
+names(df.users.pod)[2] <- "posts"
+
+load('./R_objects/dfposts_4chan.Rda')
+df.posts.4ch <- df.posts
[…]

### 2016-04-27 sandbox/cost.r

@@ -1,0 +1,30 @@
+test <- function(n){
+  cost <- c()
+  for (i in 1:n){
+    
+  }
+}
+
+
+1
+2
+3
+4
+5
+6
+7
+8
+9
+10
+11
+12
+13
+14
+15
+16
+17
+18
+19
+20
+
+probability 

### 2016-04-27 pipeline_extract_neighbourhoods.r
extracted structural neighbourhoods

@@ -32,7 +32,7 @@ source('R/plotting.r')
 ##########################################################
 # Load Data
 ##########################################################
-load('./R_objects/dfposts_podemos.Rda')
+load('./R_objects/dfposts_4chan.Rda')
 df.posts$date <- as.numeric(df.posts$date)
 df.posts <- data.frame(df.posts) %>% arrange(date)
 df.posts <- df.posts[1:75000,] # Paper
@@ -60,7 +60,7 @@ cat('Number of active users', nrow(filter(df.users, posts>MIN_POSTS)))
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/235))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/135))
 length(chunks)
 
 if(FALSE){
@@ -100,12 +100,13 @@ stopCluster(cl)
 res <- merge.motif.counts(res.parallel)
 # save(res, file='./R_objects/res_time_75000_podemos.Rda') 
 # save(res, file='./R_objects/res_order_75000_podemos.Rda') 
-save(res, file='./R_objects/res_struct_75000_podemos.Rda') 
+#save(res, file='./R_objects/res_struct_75000_podemos.Rda') 
 # save(res, file='./R_objects/res_time_75000_gameofthrones.Rda') 
 # save(res, file='./R_objects/res_order_75000_gameofthrones.Rda') 
+#save(res, file='./R_objects/res_struct_75000_gameofthrones.Rda') 
 # save(res, file='./R_objects/res_time_75000_4chan.Rda')
 # save(res, file='./R_objects/res_order_75000_4chan.Rda')
-
+save(res, file='./R_objects/res_struct_75000_4chan.Rda')
 #load("res_time_75000.Rda")
 
 #save(res, file='./R_objects/res_order_2_4_75000_4chan.Rda') 

### 2016-04-27 doc/IEEEtran/paper_IEEEformat.tex

@@
 We close the paper with some suggestions of future research.
-\begin{figure}
-	\centering
+%\begin{figure}
+%	\centering
 	%\includegraphics[width=0.3\textwidth]{tree1}
-	\includegraphics[width=0.3\textwidth]{tree2}
+%	\includegraphics[width=0.3\textwidth]{tree2}
 	%\includegraphics[width=0.3\textwidth]{tree3}
-	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
-	\label{fig:trees}
-\end{figure}
+%	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
+%	\label{fig:trees}
+%\end{figure}
 
@@
 We represent a discussion thread by a tree graph $G=(V,E)$ where $V$ is a set of $n$ vertices representing the posts (also called messages or comments) and $E$ is a set of $n-1$ directed edges that say which posts replied to which post.
-Vertices have three attributes namely order, time, and author.
+Vertices have two attributes namely time and author.
 If for two given vertices $v_i,v_j$ there exists an edge  $e=(v_i, v_j) \in E$ then $v_j$ is called the \textit{parent} of $v_i$, denoted as $p(v_i)=v_j$.
@@
 The main limitation of this definition is that the obtained neighbourhoods might be very different with different choices of $r$ and $n$, and we have no \textit{a priori} criteria to choose the proper parameters other than making them small so that the neighbourhoods capture the local dynamic of the conversation around the post $i$.
-We will tackle this issue in the next definition.
 
@@
 However, the pace at which posts are added to the conversation may be very different between conversation threads (and also within a thread) and we have no \textit{a priori} criteria for a proper choice of $\tau$.
+
 Rather than looking for a fixed time radius, we may look at changes in the pace how new posts are added to the thread.
 In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values.
+If sequences are timestamps, they are monotonic increasing, therefore the changepoints will correspond to sudden pauses in the conversation.
 If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$.
+For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by its own authors is available in the R library \texttt{changepoint}.}.
 Now we can introduce our second definition of neighbourhood.
@@
 \end{definition}
-For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by its own authors is available in the R library \texttt{changepoint}.}.
+Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$.
+
+% Explain algorithm
+Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the given definition.
+Since the breakpoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical breakpoints in the tree.
+In the case of multiple branches with some common posts, we consider that a common post is a vertical breakpoint if it is a vertical breakpoint in any of the branches.
+Once we have the breakpoints we can proceed with the algorithm.
+First, we extract the structural-neighbourhood.
+The time-based neighbourhood will be a subset of the later.
+Then, we look for horizontal and vertical breakpoints, which mark the frontiers of the time-based neighbourhood.
+There are four possible cases:
+
+\begin{itemize}
+\item A \textit{vertical breakpoint in the ancestors}:  If the breakpoint is in the path between the post and the root, then the breakpoint started the new local dynamic to which the ego post belongs.
+Thus, we remove the ancestors of the breakpoint, but not the breakpoint.
+\item A  \textit{vertical breakpoint in the descendants}: If the breakpoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the breakpoint and the breakpoint itself.
+\item An \textit{horizontal breakpoint either in the older siblings or in the ancestors}: If the breakpoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs.
+Similarly, if the breakpoint is among the ancestors, then every older sibling of the breakpoint belongs to a previous local dynamic.
+In both cases, we remove the older siblings of the breakpoint but not the breakpoint.
+\item An \textit{horizontal breakpoint elsewhere}: In any other case, the horizontal breakpoint starts a different local dynamic and therefore we remove its younger siblings and the breakpoint. 
+\end{itemize}
+Any time a vertex is remove we also remove its descendants.
+The final neighbourhood will be connected induced subgraph from the structural neighbourhood.
+If there are no changepoints in the structural neighbourhood, the time-base neighbourhood will be exactly the structural neighbourhood.
+
+%Example
 An example of time-based neighbourhood is given in Figure~\ref{fig:cutpoints} (right).
-Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$.
-Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods according to the given definition.
+
 \begin{algorithm}[H]
@@
 \ENDFOR
-\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts) from $\mathcal{N}_i(r)$
+\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts $\cup$ descendants(dump\_posts)) from $\mathcal{N}_i(r)$
 \end{algorithmic}
@@
 \section{Neighbourhood colouring and pruning}\label{sec:colouring_pruning}
+\subsection{Colouring}
 Even if the structure contains some important information about the type of conversation, there is still some ambiguity left, and two similar neighbourhoods can represent very different types of conversation.
@@
 
-In particular, we assign the red color to the ego post, the orange colour to all other posts written by the author of the ego (to identify re-entries) a white color to the root (if included in the neighbourhood), the yellow color to all posts written by the same author who wrote the parent of ego (to identify, for instances, debates between the ego author and the other author), the grey color to all posts written before the ego and the black color to all posts written after the ego.
-Since some posts might correspond to several colors, we perform a sequential assignment of colors given by $\texttt{grey, black, orange, ego, yellow, white}$.
+In particular, we assign the following colours:
+
+\begin{itemize}
+\item \textit{Red}: ego post.
+\item \textit{Orange}: other posts written by the author of ego. It allows to identify re-entries (the same author participating more times in the discussion \cite{Backstrom2013})
+\item \textit{Yellow}: parent of ego post and other posts written by the same author.
+It allows to identify, for instance, debates between the ego author (red and orange) and the other author (yellow).
+We have observed this phenomena in our data, and it is often the cause of long chains.
+\item \textit{White}: root post.
+Differentiating the root post from the rest has been proven by previous research on online discussions.
+For instance, some types of users seem to get more replies than others when they initiate a thread (\cite{Himelboim2009,  Lumbreras2013}).
+Also, in terms of preferential attachment, root posts use to get more replies than the non-roots \cite{Gomez2010, Gomez2012}.
+\item \textit{Black}: none of the above.
+\end{itemize}
+Since some posts might correspond to several colors, we perform a sequential assignment of colors given by $\texttt{black, orange, red, yellow, white}$.
+
+We do not claim this choice of labels to be of universal.
+Indeed, other labellings might also give interesting results since they would look at conversations from new points of view (e.g.: a colour for leaf vertices, colours according to the length of the post, or colours to represent the sentiment of the post).
 
@@
 
+\subsection{Pruning}
 At this point, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes.
 Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and where neither of them has any children.
+The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences.
+In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.
 
@@
 The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between the April 25th and October 20th 2014.
-The Game of Thrones dataset contains 75,000 posts corresponding to XXX threads written by XXXXX users.
+The Game of Thrones dataset contains 75,000 posts corresponding to 5,862 threads written by 18,907 users.
 

### 2016-04-28 R/count_motifs.r
processed time-based with infinite radius in GoT

@@ -138,7 +138,7 @@ count_motifs_by_post <- function(threads,
           eg <- neighborhood.temporal.order(gp, j, rad, max.neighbors)
         }
         if(neighbourhood=='time'){
-          eg <- neighborhood.temporal(gp, j, 2, breakpoints.v, breakpoints.h)
+          eg <- neighborhood.temporal(gp, j, rad, breakpoints.v, breakpoints.h)
         }
         if(neighbourhood=='struct'){
           eg <- make_ego_graph(gp, rad, nodes=j)[[1]]
@@ -164,7 +164,6 @@ count_motifs_by_post <- function(threads,
 
         # If neighbourhood is too big, sometimes isomorphism check takes too long
         if(vcount(eg)>10){
-          #cat("\nSkipping neighbourhood isomorphism in thread ", threads[i])
           next
         }
 

### 2016-04-28 pipeline_extract_neighbourhoods.r

@@ -32,7 +32,7 @@ source('R/plotting.r')
 ##########################################################
 # Load Data
 ##########################################################
-load('./R_objects/dfposts_4chan.Rda')
+load('./R_objects/dfposts_gameofthrones.Rda')
 df.posts$date <- as.numeric(df.posts$date)
 df.posts <- data.frame(df.posts) %>% arrange(date)
 df.posts <- df.posts[1:75000,] # Paper
@@ -60,7 +60,7 @@ cat('Number of active users', nrow(filter(df.users, posts>MIN_POSTS)))
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/135))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/300))
 length(chunks)
 
 if(FALSE){
@@ -82,7 +82,7 @@ res.seq <- count_motifs_by_post(as.vector(unlist(chunks)),
 }
 
 # parallel
-ncores <- detectCores() - 6
+ncores <- detectCores() - 5
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
 pck <- c('RSQLite', 'data.table', 'changepoint')
@@ -94,7 +94,7 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   #                                  120, onTimeout='warning')
   count_motifs_by_post(chunks[[i]], 
                        database='reddit',
-                       neighbourhood='struct', chunk.id=i)
+                       neighbourhood='time', rad=100, chunk.id=i)
 }
 stopCluster(cl)
 res <- merge.motif.counts(res.parallel)
@@ -104,9 +104,11 @@ res <- merge.motif.counts(res.parallel)
 # save(res, file='./R_objects/res_time_75000_gameofthrones.Rda') 
 # save(res, file='./R_objects/res_order_75000_gameofthrones.Rda') 
 #save(res, file='./R_objects/res_struct_75000_gameofthrones.Rda') 
+save(res, file='./R_objects/res_time_noradius_75000_gameofthrones.Rda') 
+
 # save(res, file='./R_objects/res_time_75000_4chan.Rda')
 # save(res, file='./R_objects/res_order_75000_4chan.Rda')
-save(res, file='./R_objects/res_struct_75000_4chan.Rda')
[…]

### 2016-04-28 doc/IEEEtran/paper_IEEEformat.tex
backup

@@
 \begin{definition}
-Given a tree graph $G$, the \textit{structural neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $r$ from post $i$.
+Given a tree graph $G$, the \textit{structure-based neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $r$ from post $i$.
 \end{definition}
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
@@
 \label{fig:cutpoints}
-\end{figure*}
+\end{figure}
 \subsection{Order-based}
@@
 In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.
-
 Figure~\ref{fig:colors} shows some real examples, from our dataset, of time-based neighbourhoods after colouring and pruning.
@@
 \footnote{TODO: It's actually much less, probably logarithmic, since the size dictionary grows incrementally}
+Some of the anlysis in the next section requires that, if two type of neighbourhoods are isomorphics, they have the same label.
+Thus, we merge the dictionaries to obtain a global dictionary of neighbours.
+Figure XXX shows some frequent neighbourhoods that will be discussed later.
 
-\subsection{Agreement between neighbourhood methods}
+\subsection{Comparing neighbourhood methods}
+In this section we compare different aspects of our three classes of neighbourhood, namely (a) the size and frequency distribution of the neighbours for each class, (b) the discrepancies between methods when they extract the neighbourhood of a same post, (c) the neighbourhood census obtained for each class in a same forum.
+
+\subsubsection{Size and frequency distribution}
 The detection of the time-based and order-based neighbourhood in the Game of Thrones dataset gave us a list of 2000 different (non-isomorphic) time-based neighbourhoods and 180 different (non-isomorphic) \footnote{TODO: Review numbers} order-based neighbourhoods, which means that the time-based is able to capture a much wider set of structures.
@@
 Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type.
-The reason why the frequency of the order-based neighbourhoods decay faster might be that the most frequent neighbourhood is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.
+The main reason why the frequency of the order-based neighbourhoods decay faster is that the most frequent neighbourhood (id XX) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.
 
-The divergent number of neighbourhoods between order-based and times-based brings us to the following question:  how differ the order-based and time-based neighbourhoods given to a same post?
-We re-labelled the order-based neighbourhoods and assigned them the label of their isomorphic time-based counterparts or the first unused number in the time-based index.
-This left us a set of 2183 trees.
+\subsubsection{Discrepancies}
+The divergent number of neighbourhoods between order-based and times-based suggests that posts with different time-based neighbourhoods have similar order-based neighbourhoods.
 Figure~\ref{fig:confusion} shows, for every post, its time-based neighbourhood and its order-based neighbourhood.
@@
 These collapse of many time-based neighbourhoods into one is significant because it happens a lot among very frequent time-based neighbourhood.
-In the right plot of Figure ~\ref{fig:confusion} we can see other order-based neighbourhoods that collapse many of the most frequent time-based neighbourhoods, such as the 22, 17, 26, 29, 49, 56, 47 and 101.
+Other order-based neighbourhoods that collapse many of the most frequent time-based neighbourhoods are the 22, 17, 26, 29, 49, 56, 47 and 101.
+The right figure \\TODO: expliquer... c'est time vs structure
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{confusion_time_order_gameofthrones}%
-\includegraphics[width=0.5\textwidth]{confusion_time_order_gameofthrones_zoom}
+\includegraphics[width=0.5\textwidth]{confusion_time_order}%
+\includegraphics[width=0.5\textwidth]{confusion_struct_time}
+\includegraphics[width=0.5\textwidth]{confusion_time_order_zoom}%
+\includegraphics[width=0.5\textwidth]{confusion_struct_time_zoom}
 \caption{Neighbourhood of posts from the point of view of order-based and time-based.
@@
 
-\subsection{Neighbourhood census}
+\subsubsection{Neighbourhood census}
 Given a set of local structures and a graph under study, a census is a counting of how many times does each structure appear in the graph.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{order_vs_time_census_distribution}%
-\includegraphics[width=0.5\textwidth]{order_vs_time_census_distribution_cum}
+\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones}%
+\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones_cum}
 \caption{Comparison of census distributions with order-based and time-based neighbourhoods.}
@@
 
-In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums.
+In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums for the 20 most frequent neighbourhood of each type.
+Because some of the most frequent neighbourhoods in one method are also among the most frequent in the other methods, the total set contains 30 different neighbourhoods.
+
 Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones.
@@
 The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ...
-(27)
+(27), suggesting that this method fails to capture some relevant structures.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{census_size_vs_freq_gameofthrones}
-\caption{Size of neighbourhoods and their frequencies.}
-\label{fig:census_size_vs_freq}
-\end{figure}
-
-This results lead us to think that the time-based neighbourhood is more convenient than the order-based.
-The main reason, we think, is that order seems to be a too dogmatic way to decide where a neighbourhood ends.
+We conclude that, even if at the cost of a sparser set of structures, the time-based neighbourhood does a better job in capturing structures that have both a real meaning and a level of frequency that makes them relevant in the analysis.
 
 \subsection{Conversation-based clustering of users}
-In this section, we cluster users based on the neighbourhoods of their posts.
+In this section, we cluster users in the Podemos dataset based on the neighbourhoods of their posts.
 Our intuition is that some users like participating in some kind of discussion rather than other.
@@
 
-We limit our study to a set of 100 \textit{active users} who wrote more than 100 posts.
+Our dataset contains a set of 100 \textit{active users} who wrote more than 100 posts.
 For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the census, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$.
@@
 We did the clustering over a feature matrix with order-based neighbourhoods and another feature matrix with time-based neighbourhoods.
-Figure~\ref{fig:PCA} and shows the PCA projections of the users coloured by their assigned cluster.
+
+Since the set of feature dimension is relatively large, we show in Figure \ref{fig:whiskers} how the users in a cluster are distributed for each feature.
+The features shown are the ones that have an bigger absolute value (in mean) for each cluster.
+
+% Explicacion
+\begin{itemize}
+\item The green cluster (attraction for X, X, repulsion XXX) correspond to users that...
+\item The red cluster...
+\item The blue cluster...
+\end{itemize}
 
@@
 	\caption{Whiskers plots in the most relevant features (time-based)}
-	\label{fig:PCA}
+	\label{fig:whiskers}
 \end{figure}
@@
 % TODO: paragraph discussing a visual/manual analysis of some users
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=0.9\textwidth]{PCA_cluster_orderbased}
-	\includegraphics[width=0.9\textwidth]{PCA_cluster_timebased}
-	\caption{PCA projections of the order-based (above) and time-based (below) neighbourhood features}
-	\label{fig:PCA}
-\end{figure*}
 
@@
 
+\section{Future work}
+In this paper we focused on the comparative analysis between three kind of neighbourhoods in order to demonstrate that the time-based neighbourhoods are able to capture relevant structures that other definitions of neighbourhood cannot.
+We also showed that it can be used to characterise users according to the kind of conversation in which they participate.
+
+We can see many applications of this kind of neighbourhood.
+
+\begin{itemize}
+\item  Caracterise reactions to user posts: instead of capturing the neighbours all around a post, we can limit we neighbourhood to the descendants.
+This would allow to cluster users from a slightly (but maybe important) point of view.
+\item The boundaries of our time-based neighbourhood are decided by the changepoint algorithm PELT. How sensible is time-based neighbourhood to the choice of the algorithm?  
+\item Merge more neighbourhoods with sociological criteria.
+\item Use different coulours to represent different phenomena.
+\end{itemize}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.8\textwidth]{dictionary1}
+	\includegraphics[width=0.8\textwidth]{dictionary2}
+	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
+	\caption{Neighbourhoods reference list}
+	\label{fig:census_orderbased}
+\end{figure*}
+
 % use section* for acknowledgment
@@
 \newpage
-\section*{Appendix}
+\section*{Appendix 2}
+
 \begin{figure*}
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.9\textwidth]{PCA_cluster_orderbased}
+	\includegraphics[width=0.9\textwidth]{PCA_cluster_timebased}
+	\caption{PCA projections of the order-based (above) and time-based (below) neighbourhood features}
+	\label{fig:PCA}
+\end{figure*}
+
 % that's all folks

### 2016-04-28 R/plotting.r

@@ -18,3 +18,24 @@ plot.tree <- function(gtree){
        vertex.size=1,
        edge.arrow.size=0.6)
 }
+
+plot.trees <- function(trees, labels){
+  # Plots a set of trees in a grid
+  # Arguments:
+  #   trees: a list of igraph tree objects
+  #   labels: a label for each tree.
+  mypalette <- c("black", "yellow", "orange", "red", "white")
+  par(mfrow=c(3,5))
+  
+  for(i in 1:length(trees)){
+    gmotif <- as.undirected(trees[[i]])
+    la = layout_as_tree(gmotif, mode='out', root=which.min(V(gmotif)$date))
+    plot(gmotif,
+         layout = la,
+         vertex.color=mypalette[V(gmotif)$color],
+         vertex.label = "",
+         edge.arrow.size=0.6)
+    title(labels[i])
+  }
+  
+}

### 2016-04-28 pipeline_compare_census.r

@@ -4,6 +4,7 @@
 library(dplyr)
 library(reshape2)
 library(igraph)
+source('R/merge_indices.r')
 if(FALSE){
   df.posts <- load_posts(database='reddit', forum='podemos')
   save(df.posts,file="dfposts_podemos.Rda")
@@ -25,7 +26,7 @@ if(FALSE){
 }
 
 ########################################################
-# Load posts
+# Load posts from each forum
 ########################################################
 load('./R_objects/dfposts_gameofthrones.Rda')
 df.posts.got <- df.posts
@@ -36,6 +37,7 @@ df.threads.got <- plyr::count(df.posts.got, "thread")
 df.users.got <- plyr::count(df.posts.got, 'user')                                                                                                                                   
 names(df.threads.got)[2] <- "length"
 names(df.users.got)[2] <- "posts"
+df.posts.got$forum <- 'gameofthrones'
 
 load('./R_objects/dfposts_podemos.Rda')
 df.posts.pod <- df.posts
@@ -46,6 +48,7 @@ df.threads.pod <- plyr::count(df.posts.pod, "thread")
 df.users.pod <- plyr::count(df.posts.pod, 'user')                                                                                                                                   
 names(df.threads.pod)[2] <- "length"
 names(df.users.pod)[2] <- "posts"
+df.posts.pod$forum <- 'podemos'
 
 load('./R_objects/dfposts_4chan.Rda')
 df.posts.4ch <- df.posts
@@ -56,13 +59,14 @@ df.threads.4ch <- plyr::count(df.posts.4ch, "thread")
 df.users.4ch <- plyr::count(df.posts.4ch, 'user')                                                                                                                                   
 names(df.threads.4ch)[2] <- "length"
 names(df.users.4ch)[2] <- "posts"
+df.posts.4ch$forum <- '4chan'
 
+#df.posts <- rbind(df.posts.got, df.posts.pod, df.posts.4ch)
 
 ########################################################
 # Load neighbourhoods
 ########################################################
-#load("./R_objects/res_struct_75000_gameofthrones.Rda")
-load("./R_objects/res_struct_75000_podemos.Rda")
[…]

### 2016-04-28 doc/IEEEtran/paper_IEEEformat.tex
backup

@@
 %
-%\usepackage{amsmath}
+\usepackage{amsmath}
 % A popular package from the American Mathematical Society that provides
@@
 % Do not put math or special symbols in the title.
-\title{You are the way you structurally talk: structural-temporal neighbourhoods of posts to characterize users in online forums}
+\title{Structural-temporal neighbourhoods of posts to characterize conversations in online forums}
 
@@
 As such, they can be studied from many levels, each of which is represented by a graph where vertices and edges may represent different things.
-The most studied graph is a graph where vertices represent users and edges represent interactions between users (a post from one user replying to a post from other user).
-From the point of view of the community, some typically analysed properties are the degree distribution, the clustering coefficient, density, or diameter of the graph.
+The most studied graph is a graph where vertices represent users and edges represent interactions between users (a post from one user replying to a post from another user).
+From the point of view of the community, some typically analysed properties are the degree distribution, the clustering coefficient, density, or the diameter of the graph.
 From the point of view of the individual users, analyses are typically focused either on the centrality of users, on the community structure or in blockmodeling (groups of users that tend to interact with the same other groups) \cite{McCallum2007a}.
-Another common graph is a tree graph where vertices represent posts (or e-mails) and and edges from one vertex to another indicates that the first is a reply to the second.
-Given the different discussion trees of a forum (a forest) the global analyses include the depth distribution, branching factors, or even the time between two posts \cite{Bhatt2012}.
-representation include depth distribution of discussions, branching factors and so.
+Another common graph is a tree graph where vertices represent posts (or e-mails) and edges from one vertex to another indicate that the first is a reply to the second.
+Given the different discussion trees of a forum (a forest) the global analyses include depth distribution, branching factors, or even the time between two posts \cite{Bhatt2012}.
 Studying discussion trees from the point of view of the users means, since trees are discussions, studying the user at a conversational level.
@@
 
-\section{Structural neighbourhoods}
+\section{Structural neighbourhoods}\label{sec:structural}
 Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, we start by the following definition of \textit{structural neighbourhood}:
@@
 \begin{definition}
-Given a tree graph $G$, the \textit{structure-based neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed of all the vertices that are at distance equal or less than $r$ from post $i$.
+Given a tree graph $G$, the \textit{structure-based neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed by all the vertices that are at distance equal or less than $r$ from post $i$.
 \end{definition}
@@
 Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite.
-This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically equivalent.
+This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically similar.
 
-\section{Structural-temporal neighbourhoods}
-As we have discussed, structural neighbourhood falls short in the analysis of conversational structures.
-Instead, we propose two new definitions of neighbourhood that take into account the order and time in which posts are written.
+\section{Structural-temporal neighbourhoods}\label{sec:temporal}
+We propose two refinements of the structure-based neighbourhood to take into account the order and time in which posts are written.
+As we will see, the criteria to choose which posts are neighbours of a post $i$  is not only based on the structural distance but also on the temporal distance to $i$.
+In the order-based definition, a new radius is set to include at most the $n$ closest posts in time.
+In the time-based definition, the new radius is set to include all posts (in the structural-neighbourhood) until the first point where the conversation slows down.
 
@@
 \includegraphics[width=0.45\textwidth]{breakpoints}
-\caption{Illustration of order-based (left) and time-based (right) neighbourhoods.
+\caption{Illustration of order-based (top) and time-based (bottom) neighbourhoods.
 Grey nodes represent the ego post.
@@
 The order-based neighbourhood has parameters $r=3$, $n=6$.
+The time-based neighbourhood has parameter $r=3$.
 In the time-based neighbourhood, horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.}
@@
 \subsection{Order-based}
-Our first definition is based on the order in which posts are attached to the tree. 
 \begin{definition}
-Given a ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
+Given a ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ and order $n$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed by the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
 \end{definition}
-An example of order-based neighbourhood is given in Figure~\ref{fig:cutpoints} (left).
+An example of order-based neighbourhood is given in Figure~\ref{fig:cutpoints}.
 This definition has two advantages over the \textit{structural neighbourhood}.
@@
 Second, the size of the neighbourhood has an upper bound of $min(|\mathcal{N}_i(r))|, n)$ thus making the space of possible neighbourhood structures finite.
-The main limitation of this definition is that the obtained neighbourhoods might be very different with different choices of $r$ and $n$, and we have no \textit{a priori} criteria to choose the proper parameters other than making them small so that the neighbourhoods capture the local dynamic of the conversation around the post $i$.
+The main limitation of this definition is that the we have no \textit{a priori} criteria to choose the proper parameter $n$ other than making it small to capture only the local dynamic of the conversation around the post $i$.
 
 \subsection{Time-based}
-In a first attempt to take time into account, one might set fixed time-based boundaries for the neighbourhood and include only those posts whose timestamp $t_j$ is at distance less than $\tau$ from the ego post $i$, $|t_j-t_i|<\tau$.
+In order to take time explicitly into account, one might set fixed time-based boundaries for the neighbourhood and include only those posts whose timestamp $t_j$ is at distance less than $\tau$ from the ego post $i$, $|t_j-t_i|<\tau$.
 However, the pace at which posts are added to the conversation may be very different between conversation threads (and also within a thread) and we have no \textit{a priori} criteria for a proper choice of $\tau$.
 
-Rather than looking for a fixed time radius, we may look at changes in the pace how new posts are added to the thread.
+Rather than looking for a fixed time radius, we may instead decide it by looking at changes in the pace how new posts are added to the thread.
 In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values.
@@
 \begin{definition}
-Given a tree graph $G$, the \textit{time-based neighbourhood} of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
+Given a tree graph $G$, the \textit{time-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
 \end{definition}
@@
 \subsection{Neighbourhood extraction}
-As a previous step for all further analysis we detect the neighbourhood around every post in our datasets.
-Each neighbourhood structure (including colours) is given a unique integer number.
+As a previous step for all further analysis we extract the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums.
+We label every structure with a unique integer number, where a structure is a coloured graph.
+Two neighbours have the same structure, and therefore they have the same integer number, if their coloured graphs are isomorphic.
+
 Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of of these trees would not be seen in the data.
 It appears then a much more practical approach to enumerate only those trees that are seen in the data.
-The consequence of this is that every dataset will have its own dictionary of trees.
+In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings) we re-label the structures after the extraction for every post is finished.
+We will call this final labelling a \textit{dictionary} of neighbourhoods.
 
-We detect the order-based and time-based neighbourhood of every post as follows.
+Given a type of neighbourhood (struture-based, order-based or time-based) we extract the neighbourhood around every post of a forum as follows.
 A counter for every neighbourhood structure is created the first time it us detected.
-For every post, we extract its neighbourhood (order-based or time-based) colour and prune it as explained above (Section~\ref{sec:colouring_pruning}).
-Then we check whether it is isomorphic to some of the already seen neighbourhoods.
+For every post, we extract its neighbourhood, colour and prune it (Sections~\ref{sec:structural} and \ref{sec:temporal}).
+Then we check whether it is isomorphic to some of the neighbourhoods that we have already seen.
 If this is the case, then we increment the counter for the neighbourhood previously seen.
 Otherwise we create a new entry for the current neighbourhood.
-The computational cost of this operation has a upper bound of $\mathcal O(n^2)$, $n$ being the number of posts, for the case where each neighbourhood occurs only once.
-\footnote{TODO: It's actually much less, probably logarithmic, since the size dictionary grows incrementally}
-Some of the anlysis in the next section requires that, if two type of neighbourhoods are isomorphics, they have the same label.
-Thus, we merge the dictionaries to obtain a global dictionary of neighbours.
-Figure XXX shows some frequent neighbourhoods that will be discussed later.
+
+Once the six extractions are finished we merge there dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth).
+Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.
+
+% Computational cost
+In the worst case, where every post has a unique neighbourhood, every extracted neighbourhood should be compared against every neighbourhood in already in the dictionary before concluding that it needs a new entry.
+In this case, the computational cost of this operation is $\mathcal{O}(n^2)$ where $n$ is the number of posts.
+In practice, however, the bigger the number of posts processed, the less likely that we go through all the dictionary without finding and isomorphism.
+Moreover, if we keep the dictionary sorted by frequency, a large number of neighbourhoods will find and isomorphism among the first neighbours of the dictionary.
 
@@
 \subsubsection{Size and frequency distribution}
-The detection of the time-based and order-based neighbourhood in the Game of Thrones dataset gave us a list of 2000 different (non-isomorphic) time-based neighbourhoods and 180 different (non-isomorphic) \footnote{TODO: Review numbers} order-based neighbourhoods, which means that the time-based is able to capture a much wider set of structures.
+We detected, in the the whole dataset, 3791 different structure-based neighbourhoods, 174 different order-based and 1488 different time-based.
+which means that the structure-based captures a much wider set of structures.
 Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times.
-Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods occur only once, and that we only need 384 time-based to cover 95\% of the census, while for the order-based, 53 structures are enough to cover the same percentage.
-Moreover, there is no much difference, in terms of frequency, between the 50 most frequent neighbours of each type.
-The main reason why the frequency of the order-based neighbourhoods decay faster is that the most frequent neighbourhood (id XX) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.
+Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods (41\%) and the structure-based neighbourhoods (60\%) occur only once, while the number of neighbourhoods that appear more than 100 times (among 150,000 posts) is similar for the three  types of neighbourhood.
+To cover 95\% of the posts we need 623 structure-based neighbourhoods, 40 order-based and 155 time-based.
+
+% Why order-based is not good enough
+The main reason why the frequency of the order-based neighbourhoods decay faster is that the most frequent neighbourhood (id 1) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.
+This is the main handicap of the order-based neighbourhood since, as we will see later, it avoids finding richer structures.
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones}%
+\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones_cum}
+\caption{Comparison of census distributions with order-based and time-based neighbourhoods.}
+\label{fig:census_distributions}
+\end{figure*}
 
 \subsubsection{Discrepancies}
-The divergent number of neighbourhoods between order-based and times-based suggests that posts with different time-based neighbourhoods have similar order-based neighbourhoods.
-Figure~\ref{fig:confusion} shows, for every post, its time-based neighbourhood and its order-based neighbourhood.
-Those in the diagonal (34\%) see no difference between the time-based and order-based.
-Out of the diagonal, there are two rows indicating a high number of posts being assign the same order-based neighbourhood but very different time-based neighbourhoods.
+The divergent number of neighbourhoods between structure-based, order-based and times-based suggests that some neighbourhoods, for instance, in time-based, might be systematically mapped to the same neighbourhoods in order-based.
+Or that some structure-base neighbourhoods are systematically mapped into the same time-based neighbourhoods.
+Figure~\ref{fig:confusion} gives a general overview of these mappings.
+The left figure shows, for every post, its time-based neighbourhood and its order-based neighbourhood, while the right figure shows the time-based and the structure-based.
+The points in the diagonal correspond to posts that have been assigned the same neighbourhood by the two compared methods.
+
+The diagonal of the time-based / order-based plot contains 36\% of posts.
+The diagonal of the structure-based / time-based plot contains 40\% of the posts.
+For the order-based / struct-based (not shown) there are 41\% of posts in the diagonal.
+In general, the structures with bigger agreement are among the most frequent.
+This is comprehensible since the most frequent tend to be also the most simple structures with less room for discrepancy.
+There is a 30.75\% of posts that are assigned the same neighbourhood structure by the three methods.
+
+Rather than looking at where the methods agree, it is even more interesting to see where they disagree, since this will reveal whether one of the methods is incapable of detecting some interesting conversational structure.
+
+%por aqui
+
+Out of the diagonal, there are two rows indicating a high number of posts being assigned to the same order-based neighbourhood but to very different time-based neighbourhoods.
 These order-based neighbourhoods are the number 1 and the number 50.
@@
 \includegraphics[width=0.5\textwidth]{confusion_struct_time}
-\includegraphics[width=0.5\textwidth]{confusion_time_order_zoom}%
-\includegraphics[width=0.5\textwidth]{confusion_struct_time_zoom}
 \caption{Neighbourhood of posts from the point of view of order-based and time-based.
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones}%
-\includegraphics[width=0.5\textwidth]{compare_census_distribution_gameofthones_cum}
-\caption{Comparison of census distributions with order-based and time-based neighbourhoods.}
-\label{fig:census_distributions}
-\end{figure*}
-
 In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums for the 20 most frequent neighbourhood of each type.
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-\includegraphics[width=0.5\textwidth]{time_neighbourhood_census_compare_forums}%
-\includegraphics[width=0.5\textwidth]{order_neighbourhood_census_compare_forums}
-\caption{Comparison of census for two different forums.}
+\includegraphics[width=0.5\textwidth]{structure_based_census}
+\includegraphics[width=0.5\textwidth]{order_based_census}
+\includegraphics[width=0.5\textwidth]{time_based_census}
+\caption{Comparison of neighbourhood census obtained by each type of neighbourhood in two different forums.
+Order-based is more similar across forums since its most frequent neighbourhood dominates the census.}
 \label{fig:census_compare}
-\end{figure*}
+\end{figure}
 
@@
 	\includegraphics[width=0.8\textwidth]{dictionary2}
-	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
 	\caption{Neighbourhoods reference list}
-	\label{fig:census_orderbased}
+	\label{fig:dictionary}
 \end{figure*}
@@
 \newpage
-\section*{Appendix 2}
+\section*{Appendix  (WILL ME REMOVED)}
 
@@
 \centering
-	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
-	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
+%\includegraphics[width=0.5\textwidth]{time_neighbourhood_census_compare_forums}%
+%\includegraphics[width=0.5\textwidth]{order_neighbourhood_census_compare_forums}
+\includegraphics[width=0.5\textwidth]{neighbourhood_census_compare_gameofthrones}%
+\includegraphics[width=0.5\textwidth]{neighbourhood_census_compare_podemos}
+\caption{Comparison of census for two different forums. POUR BERTRAND. WE CANT SEE ANYTHING HERE}
+\label{fig:census_compare}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+%	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
+%	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
 	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
@@
 	\centering
-	\includegraphics[width=0.8\textwidth]{census_timebased_1}
-	\includegraphics[width=0.8\textwidth]{census_timebased_2}
+%	\includegraphics[width=0.8\textwidth]{census_timebased_1}
+%	\includegraphics[width=0.8\textwidth]{census_timebased_2}
 	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
@@
 \centering
-\includegraphics[width=1\textwidth]{confusion_time_order_motifs_gameofthrones_1}
+%\includegraphics[width=1\textwidth]{confusion_time_order_motifs_gameofthrones_1}
 \caption{Most popular time-based neighbourhoods (up) and how they are mostly detected in order-based neighbourhood (down).}

### 2016-04-28 pipeline_compare_census.r

@@ -50,18 +50,16 @@ names(df.threads.pod)[2] <- "length"
 names(df.users.pod)[2] <- "posts"
 df.posts.pod$forum <- 'podemos'
 
-load('./R_objects/dfposts_4chan.Rda')
-df.posts.4ch <- df.posts
-df.posts.4ch$date <- as.numeric(df.posts.4ch$date)
-df.posts.4ch <- data.frame(df.posts.4ch) %>% arrange(date)
-df.posts.4ch <- df.posts.4ch[1:75000,] # Paper
-df.threads.4ch <- plyr::count(df.posts.4ch, "thread")
-df.users.4ch <- plyr::count(df.posts.4ch, 'user')                                                                                                                                   
-names(df.threads.4ch)[2] <- "length"
-names(df.users.4ch)[2] <- "posts"
-df.posts.4ch$forum <- '4chan'
-
-#df.posts <- rbind(df.posts.got, df.posts.pod, df.posts.4ch)
+# load('./R_objects/dfposts_4chan.Rda')
+# df.posts.4ch <- df.posts
+# df.posts.4ch$date <- as.numeric(df.posts.4ch$date)
+# df.posts.4ch <- data.frame(df.posts.4ch) %>% arrange(date)
+# df.posts.4ch <- df.posts.4ch[1:75000,] # Paper
+# df.threads.4ch <- plyr::count(df.posts.4ch, "thread")
+# df.users.4ch <- plyr::count(df.posts.4ch, 'user')                                                                                                                                   
+# names(df.threads.4ch)[2] <- "length"
+# names(df.users.4ch)[2] <- "posts"
+# df.posts.4ch$forum <- '4chan'
 
 ########################################################
 # Load neighbourhoods
@@ -112,7 +110,7 @@ colnames(df.posts.got)[which(names(df.posts.got) == "motif")] <- "motif.time"
 n <- as.numeric(table(df.posts.got$motif.time))
 all(n == cummin(n))
 
-#################################################
+#####################
 load("./R_objects/res_struct_75000_podemos.Rda")
 res.pod.struct <- res
 
@@ -155,56 +153,56 @@ colnames(df.posts.pod)[which(names(df.posts.pod) == "motif")] <- "motif.time"
 n <- as.numeric(table(df.posts.pod$motif.time))
 all(n == cummin(n))
 
-#########
-load("./R_objects/res_struct_75000_4chan.Rda")
-res.4ch.struct <- res
-
[…]

### 2016-04-29 doc/IEEEtran/paper_IEEEformat.tex
backup

@@
 The divergent number of neighbourhoods between structure-based, order-based and times-based suggests that some neighbourhoods, for instance, in time-based, might be systematically mapped to the same neighbourhoods in order-based.
-Or that some structure-base neighbourhoods are systematically mapped into the same time-based neighbourhoods.
+Or similarly, that some structure-base neighbourhoods are systematically mapped into the same time-based neighbourhoods.
 Figure~\ref{fig:confusion} gives a general overview of these mappings.
@@
 Rather than looking at where the methods agree, it is even more interesting to see where they disagree, since this will reveal whether one of the methods is incapable of detecting some interesting conversational structure.
+In Figure~\ref{fig:confusion} some horizontals lines can be spotted specially in the lower part of the plots.
+These correspond to posts with different time-based neighbourhoods that are assigned the same order-based neighbourhood (left) and different time-based neighbourhoods that are assigned the same order-based neighbourhood (right).
+Looking for the order-based neighbourhoods that have a correspondence with a larger set of time-based neighbourhoods, we detect that these correspond to neighbourhoods 1 (77 different time-based) and 50 (57 different time-based), 17 and 18.
+Looking for the time-based neighbourhoods that have a correspondence with a larger set of structure-base neighbourhoods, we detect that these are neighbourhoods 10, 4, 8, 19, 14.
 
-%por aqui
-
-Out of the diagonal, there are two rows indicating a high number of posts being assigned to the same order-based neighbourhood but to very different time-based neighbourhoods.
-These order-based neighbourhoods are the number 1 and the number 50.
-These collapse of many time-based neighbourhoods into one is significant because it happens a lot among very frequent time-based neighbourhood.
-Other order-based neighbourhoods that collapse many of the most frequent time-based neighbourhoods are the 22, 17, 26, 29, 49, 56, 47 and 101.
-The right figure \\TODO: expliquer... c'est time vs structure
+% Wrap-up
+In retrospective, these results are not surprising.
+What we see is simply that structures that are often captured by one method collapse into another single structure when other method with a smaller space of structures is used.
+Yet, the three methods mostly agree when they find a frequent neighbourhood.
+But this might be due to the fact that simpler motifs tend to be in shorter threads, where there is no room for two neighbourhood methods to detect a different structure.
 
@@
 However, triads are poor objects when studying trees since there are only three possible structures with, at most, three nodes: a single vertex, a directed dyad (reply), star (two replies to the same post) and a chain (reply and reply to the reply).
-On the contrary, our neighbourhoods are potentially richer structures that capture the real structure of conversations.
+On the contrary, our neighbourhoods are potentially richer structures that respect the real structure of conversations.
 

### 2016-04-29 pipeline_compare_census.r

@@ -416,6 +416,13 @@ df.agree <- filter(df.posts.global, motif.order==motif.struct)
 nrow(df.agree)/nrow(df.posts.global)
 # 41.48%
 
+# Which order-based neighgbourhoods have a bigger diversity of time-based ounterparts?
+# use the confusion matrices, it will be easier
+n.sources <- apply(confusion.order_time, 1, function(x) length(unique(x)))
+sort(n.sources[1:100], decreasing = TRUE)
+
+n.sources <- apply(confusion.time_struct, 1, function(x) length(unique(x)))
+sort(n.sources[1:100], decreasing = TRUE)
 #######################################################################
 # Census of the 3 types of neighbourhood in the same forum
 #######################################################################

### 2016-04-29 doc/IEEEtran/paper_IEEEformat.tex
census finished

@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-\includegraphics[width=0.2\textwidth]{neighbourhood_time_342}
+%\includegraphics[width=0.2\textwidth]{neighbourhood_time_342}
 \includegraphics[width=0.19\textwidth]{neighbourhood_time_456}
 \includegraphics[width=0.18\textwidth]{neighbourhood_time_501}
-\includegraphics[width=0.18\textwidth]{neighbourhood_time_524}
+%\includegraphics[width=0.18\textwidth]{neighbourhood_time_524}
 \caption{Illustration of different real structures using different colors for the root (white), the ego (red) and other posts written by the same author (orange) the ego parent and other posts written by the same author (yellow), posts previous to the ego (grey) and posts posterior to the ego.}
 \label{fig:colors}
-\end{figure*}
+\end{figure}
 
@@
 
-In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums for the 20 most frequent neighbourhood of each type.
-Because some of the most frequent neighbourhoods in one method are also among the most frequent in the other methods, the total set contains 30 different neighbourhoods.
+In Figure \ref{fig:census_compare} we show, for every neighbourhood method, the resulting census in the two forums.
+For each method, the neighbourhoods shown are those among the 30 most frequent in either of the forums.
+Out of this selection, we checked that no neighbourhood has a frequency higher than 1000 (near a 1.5\% of the posts in each forum), which guarantees that we are not missing important information in the plots.
 
-Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones.
-These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25).
-The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ...
-(27), suggesting that this method fails to capture some relevant structures.
+According to the structure-based method, the Podemos census has a bigger presence of 9 and 20, though 9 is also dominant in Game of Thrones.
+The Game of Thrones has a bigger presence of 10, 11 and 27.
+The 20 is interesting since it represents chain discussions between two users, which in the context of a political forum suggests a debate.
+The 11 represents the ego user opening a new discussion but getting only one reply, which means an unsuccessful thread.
+This structure is much less frequent in Podemos meaning that, for some reason, in Podemos discussions have easier success.
+The 27 represents a user replying, among others, to the same post which is, in its turn, a reply to another regular post.
+
+According to the order-based method, the Podemos census has a bigger presence of 1, 3, 2, 13, 4 and 18, most of them representing chains.
+The Game of Thrones has a bigger presence of 1, 12, 9, 3, 11 and 5.
+Although it captures some of the chain structure in the Podemos forum, the dominance of the motifs 1 clearly deteriorates the census.
+
+Lastly, the time-based method gives the Podemos census a bigger presence of  2,3,4 and 8.
+The Game of Thrones has a bigger presence of 5, 9, 11 and 25.
+As seen in the structure-based, 9 and 11 suggest threads that did not attract the attention of the users.
 
@@
 
-We conclude that, even if at the cost of a sparser set of structures, the time-based neighbourhood does a better job in capturing structures that have both a real meaning and a level of frequency that makes them relevant in the analysis.
+It is clear from the last sections that the order-based does not do a good job on capturing all the richness of the conversational structures.
+Nevertheless, it is harder to decide between the time-based or the structure-based in terms of the richness of the captured conversations.
+There are two other reasons to choose the time-neighbourhood over the structure based.
+First, the time-base has less spurious structures, covering 95\% of posts with 155 neighbourhoods, against the 623 neighbours necessary to cover that amount with the structure-based.
+Second, the time-based neighbourhood has more grounded criteria to decide where the neighbourhood ends, while the structure-based has no criteria at all.
 

### 2016-04-29 pipeline_compare_census.r

@@ -434,87 +434,110 @@ df.posts <- df.posts.global[df.posts.global$forum=='podemos',]
 df.posts <- filter(df.posts.global, forum=='gameofthrones')
 #df.posts <- filter(df.posts.global, forum=='podemos')
 
-ntop <- 50
-top.struct <- order(tabulate(df.posts$motif.struct), decreasing = TRUE)[1:ntop]
-top.order <- order(tabulate(df.posts$motif.order), decreasing = TRUE)[1:ntop]
-top.time <- order(tabulate(df.posts$motif.time), decreasing = TRUE)[1:ntop]
-top <- union(union(top.order, top.time), top.struct) %>% sort
-#top <- 1:max(top) # fill the gaps
+# We don't see anything clear in this plots
+if(FALSE){
+  ntop <- 50
+  top.struct <- order(tabulate(df.posts$motif.struct), decreasing = TRUE)[1:ntop]
+  top.order <- order(tabulate(df.posts$motif.order), decreasing = TRUE)[1:ntop]
+  top.time <- order(tabulate(df.posts$motif.time), decreasing = TRUE)[1:ntop]
+  top <- union(union(top.order, top.time), top.struct) %>% sort
+  #top <- 1:max(top) # fill the gaps
+  
+  counts.struct <- df.posts$motif.struct[df.posts$motif.struct %in% top]
+  counts.order <- df.posts$motif.order[df.posts$motif.order %in% top]
+  counts.time <- df.posts$motif.time[df.posts$motif.time %in% top]
+  census.struct <- sapply(top, function(x) sum(counts.struct==x))
+  census.order <- sapply(top, function(x) sum(counts.order==x))
+  census.time <- sapply(top, function(x) sum(counts.time==x))
+  
+  par(mfrow=c(1,1))
+  plot(1:length(top), census.order, pch=19, col='red',
+       type='l',xlab='Neighborhood', ylab='Frequency', axes=FALSE)
+  lines(1:length(top), census.time, pch=19, col='black')
+  lines(1:length(top), census.struct, pch=19, col='blue')
+  axis(1, at=1:length(top), labels=top, cex=0.2) 
+  axis(2) 
+  legend(length(top)-15,12000, c('order-based', 'time_based', 'struct-based'), col=c('red', 'black', 'blue'), pch=c(18,20))
+  title('Neighbourhood census in Game of Thrones')
+  
+  plot.trees(motifs.global[top], top)
+  
+}
 
[…]

### 2016-04-29 doc/IEEEtran/paper_IEEEformat.tex
paper almost final

--- 2016-04-29 doc/IEEEtran/paper_IEEEformat.tex
@@
 \usepackage{algorithm}
+\usepackage{multirow}
 \newtheorem{definition}{Definition}
@@
 If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$.
-For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by its own authors is available in the R library \texttt{changepoint}.}.
+For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by the authors themselves is available in the R library \texttt{changepoint}.}.
 Now we can introduce our second definition of neighbourhood.
@@
 % Explain algorithm
-Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the given definition.
+Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the above definition.
 Since the breakpoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical breakpoints in the tree.
 In the case of multiple branches with some common posts, we consider that a common post is a vertical breakpoint if it is a vertical breakpoint in any of the branches.
-Once we have the breakpoints we can proceed with the algorithm.
+Once we have the breakpoints, we can proceed with the algorithm.
 First, we extract the structural-neighbourhood.
@@
 \item A \textit{vertical breakpoint in the ancestors}: If the breakpoint is in the path between the post and the root, then the breakpoint started the new local dynamic to which the ego post belongs.
-Thus, we remove the ancestors of the breakpoint, but not the breakpoint.
+Thus, we remove the ancestors of the breakpoint, but not the breakpoint itself.
 \item A \textit{vertical breakpoint in the descendants}: If the breakpoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the breakpoint and the breakpoint itself.
-\item An \textit{horizontal breakpoint either in the older siblings or in the ancestors}: If the breakpoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs.
+\item A \textit{horizontal breakpoint either in the older siblings or in the ancestors}: If the breakpoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs.
 Similarly, if the breakpoint is among the ancestors, then every older sibling of the breakpoint belongs to a previous local dynamic.
-In both cases, we remove the older siblings of the breakpoint but not the breakpoint.
-\item An \textit{horizontal breakpoint elsewhere}: In any other case, the horizontal breakpoint starts a different local dynamic and therefore we remove its younger siblings and the breakpoint. 
+In both cases, we remove the older siblings of the breakpoint but not the breakpoint itself.
+\item A \textit{horizontal breakpoint elsewhere}: In any other case, the horizontal breakpoint starts a different local dynamic and therefore we remove its younger siblings and the breakpoint. 
 \end{itemize}
-Any time a vertex is remove we also remove its descendants.
+Any time a vertex is removed we also remove its descendants.
 The final neighbourhood will be connected induced subgraph from the structural neighbourhood.
-If there are no changepoints in the structural neighbourhood, the time-base neighbourhood will be exactly the structural neighbourhood.
+If there are no changepoints in the structural neighbourhood, the time-based neighbourhood will be exactly the structural neighbourhood.
 
@@
 Even if the structure contains some important information about the type of conversation, there is still some ambiguity left, and two similar neighbourhoods can represent very different types of conversation.
-We can easily reduce this ambiguity by assigning colors, or labels, to vertices, identifying some relevant property of the post.
+We can easily reduce this ambiguity by assigning colours to vertices, which allows us to identify some relevant property of the post.
 
@@
 \item \textit{Red}: ego post.
-\item \textit{Orange}: other posts written by the author of ego. It allows to identify re-entries (the same author participating more times in the discussion \cite{Backstrom2013})
+\item \textit{Orange}: other posts written by the author of ego. It allows to identify re-entries (when the same author participates several times in the discussion \cite{Backstrom2013})
 \item \textit{Yellow}: parent of ego post and other posts written by the same author.
@@
 For instance, some types of users seem to get more replies than others when they initiate a thread (\cite{Himelboim2009,  Lumbreras2013}).
-Also, in terms of preferential attachment, root posts use to get more replies than the non-roots \cite{Gomez2010, Gomez2012}.
+Also, in terms of preferential attachment, root posts usually get more replies than the non-roots (\cite{Gomez2010, Gomez2012}).
 \item \textit{Black}: none of the above.
 \end{itemize}
-Since some posts might correspond to several colors, we perform a sequential assignment of colors given by $\texttt{black, orange, red, yellow, white}$.
+Since some posts might correspond to several colours, we perform a sequential assignment of colours given by $\texttt{black, orange, red, yellow, white}$.
 
 We do not claim this choice of labels to be of universal.
-Indeed, other labellings might also give interesting results since they would look at conversations from new points of view (e.g.: a colour for leaf vertices, colours according to the length of the post, or colours to represent the sentiment of the post).
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.2\textwidth]{neighbourhood_time_342}
-\includegraphics[width=0.19\textwidth]{neighbourhood_time_456}
-\includegraphics[width=0.18\textwidth]{neighbourhood_time_501}
-\includegraphics[width=0.18\textwidth]{neighbourhood_time_524}
-\caption{Illustration of different real structures using different colors for the root (white), the ego (red) and other posts written by the same author (orange) the ego parent and other posts written by the same author (yellow), posts previous to the ego (grey) and posts posterior to the ego.}
-\label{fig:colors}
-\end{figure*}
+Indeed, other labellings might also give interesting results since they would look at conversations from new points of view (e.g.: a colour for leaf vertices, colours according to the post length, or colours to represent the sentiment of the post).
 
 \subsection{Pruning}
-At this point, we can still find structures whose only difference is that one of them has more leafs hanging from some of the nodes.
-Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same color and where neither of them has any children.
+At this point, we can still find structures the only difference of which is that one of them has more leafs hanging from some of the nodes.
+Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same colour and where neither of them has any children.
 The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences.
 In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.
-Figure~\ref{fig:colors} shows some real examples, from our dataset, of time-based neighbourhoods after colouring and pruning.
 
@@
 Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}.
-The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid the October 18th and 19th 2014.
+The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid on October 18th and 19th, 2014.
 Nowadays, its members use it mainly to share and discuss about political news.
-The Game of Thrones is a casual discussion forum about the Game of Thrones TV series.
-The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between the April 25th and October 20th 2014.
+The Game of Thrones forum is a casual discussion forum about the Game of Thrones TV series.
+The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between April 25th and October 20th, 2014.
 The Game of Thrones dataset contains 75,000 posts corresponding to 5,862 threads written by 18,907 users.
@@
 \subsection{Neighbourhood extraction}
-As a previous step for all further analysis we extract the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums.
+As a previous step for all further analysis, we extract the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums.
 We label every structure with a unique integer number, where a structure is a coloured graph.
-Two neighbours have the same structure, and therefore they have the same integer number, if their coloured graphs are isomorphic.
+Two neighbours have the same structure, and therefore the same integer number, if their coloured graphs are isomorphic.
 
-Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of of these trees would not be seen in the data.
+Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of these trees would not be seen in the data.
 It appears then a much more practical approach to enumerate only those trees that are seen in the data.
-In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings) we re-label the structures after the extraction for every post is finished.
+In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings), we re-label the structures after the extraction for every post is finished.
 We will call this final labelling a \textit{dictionary} of neighbourhoods.
 
-Given a type of neighbourhood (struture-based, order-based or time-based) we extract the neighbourhood around every post of a forum as follows.
-A counter for every neighbourhood structure is created the first time it us detected.
+Given a type of neighbourhood (structure-based, order-based or time-based) we extract the neighbourhood around every post of a forum as follows.
+A new label is assigned to every neighbourhood structure the first time it is detected.
 For every post, we extract its neighbourhood, colour and prune it (Sections~\ref{sec:structural} and \ref{sec:temporal}).
-Then we check whether it is isomorphic to some of the neighbourhoods that we have already seen.
-If this is the case, then we increment the counter for the neighbourhood previously seen.
-Otherwise we create a new entry for the current neighbourhood.
+Then we check whether it is isomorphic to some of the $n$ neighbourhoods that we have already seen.
+If this is the case, then the current neighbourhood is given the same label than the neighbourhood that we already saw.
+Otherwise we create a new label for the current neighbourhood.
 
-Once the six extractions are finished we merge there dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth).
+Once the six extractions are finished, we merge the dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth).
 Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.
@@
 % Computational cost
-In the worst case, where every post has a unique neighbourhood, every extracted neighbourhood should be compared against every neighbourhood in already in the dictionary before concluding that it needs a new entry.
+In the worst case, where every post has a unique neighbourhood, every extracted neighbourhood should be compared against every neighbourhood that already appears in the dictionary before concluding that it needs a new entry.
 In this case, the computational cost of this operation is $\mathcal{O}(n^2)$ where $n$ is the number of posts.
-In practice, however, the bigger the number of posts processed, the less likely that we go through all the dictionary without finding and isomorphism.
-Moreover, if we keep the dictionary sorted by frequency, a large number of neighbourhoods will find and isomorphism among the first neighbours of the dictionary.
+In practice, however, the bigger the number of posts processed, the less likely we go through all the dictionary without finding an isomorphism.
+Moreover, if we keep the dictionary sorted by frequency, a large number of neighbourhoods will find an isomorphism among the first neighbours of the dictionary.
 
 \subsection{Comparing neighbourhood methods}
-In this section we compare different aspects of our three classes of neighbourhood, namely (a) the size and frequency distribution of the neighbours for each class, (b) the discrepancies between methods when they extract the neighbourhood of a same post, (c) the neighbourhood census obtained for each class in a same forum.
+In this section we compare different aspects of our three classes of neighbourhoods, namely (a) the size and frequency distribution of the neighbours for each class, (b) the discrepancies between methods when they extract the neighbourhood of a same post, and (c) the neighbourhood census obtained for each class in the same forum.
 
 \subsubsection{Size and frequency distribution}
-We detected, in the the whole dataset, 3791 different structure-based neighbourhoods, 174 different order-based and 1488 different time-based.
-which means that the structure-based captures a much wider set of structures.
-Of course, this is not necessarily and advantage since these extra structures may be just spurious and only occur a very small number of times.
+We detected in the the whole dataset 3,791 different structure-based neighbourhoods, 174 different order-based and 1,488 different time-based, which means that the structure-based captures a much wider set of structures.
+Of course, this is not necessarily an advantage since these extra structures may be just spurious and only occur a very small number of times.
 Indeed, Figure~\ref{fig:census_distributions} shows that around half of the time-based neighbourhoods (41\%) and the structure-based neighbourhoods (60\%) occur only once, while the number of neighbourhoods that appear more than 100 times (among 150,000 posts) is similar for the three types of neighbourhood.
@@
 % Why order-based is not good enough
-The main reason why the frequency of the order-based neighbourhoods decay faster is that the most frequent neighbourhood (id 1) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.
+The main reason why the frequency of the order-based neighbourhoods decays faster is that the most frequent neighbourhood (id 1) is much more frequent (upper-left red point in the left plot of figure \ref{fig:census_distributions}) than the most frequent time-based neighbourhood.
 This is the main handicap of the order-based neighbourhood since, as we will see later, it avoids finding richer structures.
@@
 
+\begin{table}
+ \begin{center}
+   \tabcolsep = 0.8\tabcolsep
+   \begin{tabular}{|cc|cc|cc|}
+      \hline
+      \multicolumn{2}{|c|}{struct-based} & \multicolumn{2}{|c|}{order-based} &  \multicolumn{2}{|c|}{time-based} \\
+      %\hline
+     	 POD & GOT 						& POD & GOT 		& POD & GOT\\
+   \hline \hline
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 1 (12.5) &  1 (12.5) & 3 (12.5)\\    
+     1 (12.5)& 3(12.5)   &  2 (12.5) & 2 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\  
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
+     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\          
+   \hline
+   \end{tabular}
+\label{tab:table}
+ \end{center}
+ \caption{Frequence}
+\end{table}
+
 \subsubsection{Discrepancies}
-The divergent number of neighbourhoods between structure-based, order-based and times-based suggests that some neighbourhoods, for instance, in time-based, might be systematically mapped to the same neighbourhoods in order-based.
-Or similarly, that some structure-base neighbourhoods are systematically mapped into the same time-based neighbourhoods.
+The divergent number of neighbourhoods between structure-based, order-based and times-based suggests that some neighbourhoods, for instance, time-based, might be systematically mapped to the same neighbourhoods in order-based.
+Or similarly, that some structure-based neighbourhoods are systematically mapped into the same time-based neighbourhoods.
 Figure~\ref{fig:confusion} gives a general overview of these mappings.
@@
 
-The diagonal of the time-based / order-based plot contains 36\% of posts.
+The diagonal of the time-based / order-based plot contains 36\% of the posts.
 The diagonal of the structure-based / time-based plot contains 40\% of the posts.
-For the order-based / struct-based (not shown) there are 41\% of posts in the diagonal.
-In general, the structures with bigger agreement are among the most frequent.
+For the order-based / struct-based (not shown) there are 41\% of the posts in the diagonal.
+In general, the structures with a bigger agreement are among the most frequent.
 This is comprehensible since the most frequent tend to be also the most simple structures with less room for discrepancy.
-There is a 30.75\% of posts that are assigned the same neighbourhood structure by the three methods.
+There is a 30.75\% of the posts that are assigned the same neighbourhood structure by the three methods.
 
@@
 In Figure~\ref{fig:confusion} some horizontals lines can be spotted specially in the lower part of the plots.
-These correspond to posts with different time-based neighbourhoods that are assigned the same order-based neighbourhood (left) and different time-based neighbourhoods that are assigned the same order-based neighbourhood (right).
+These correspond to posts with different time-based neighbourhoods that are assigned the same order-based neighbourhood (left), and different time-based neighbourhoods that are assigned the same order-based neighbourhood (right).
 Looking for the order-based neighbourhoods that have a correspondence with a larger set of time-based neighbourhoods, we detect that these correspond to neighbourhoods 1 (77 different time-based) and 50 (57 different time-based), 17 and 18.
-Looking for the time-based neighbourhoods that have a correspondence with a larger set of structure-base neighbourhoods, we detect that these are neighbourhoods 10, 4, 8, 19, 14.
+Looking for the time-based neighbourhoods that have a correspondence with a larger set of structure-based neighbourhoods, we detect that these are neighbourhoods 10, 4, 8, 19, 14.
 
@@
 In retrospective, these results are not surprising.
-What we see is simply that structures that are often captured by one method collapse into another single structure when other method with a smaller space of structures is used.
+What we see is simply that structures that are often captured by one method collapse into another single structure when another method with a smaller space of structures is used.
 Yet, the three methods mostly agree when they find a frequent neighbourhood.
@@
 
-\subsubsection{Neighbourhood census}
-Given a set of local structures and a graph under study, a census is a counting of how many times does each structure appear in the graph.
-In the context of online discussion forums, triad census has been sometimes used in the graph of users interactions are often used (in these graphs, an edge exists from user $u$ to user $v$ if $u$ wrote a post reply to some of the posts written by $v$ in some thread.) \cite{Adamic2008, Lumbreras2013}.
-However, triads are poor objects when studying trees since there are only three possible structures with, at most, three nodes: a single vertex, a directed dyad (reply), star (two replies to the same post) and a chain (reply and reply to the reply).
-On the contrary, our neighbourhoods are potentially richer structures that respect the real structure of conversations.
-
-In Figure \ref{fig:census_compare} we compare the neighbourhood census in the Podemos and the Game of Thrones forums for the 20 most frequent neighbourhood of each type.
-Because some of the most frequent neighbourhoods in one method are also among the most frequent in the other methods, the total set contains 30 different neighbourhoods.
-
-Interestingly, two time-based neighbourhoods are  more than three times likelier in the Game of Thrones.
-These neighbourhoods correspond to a discussion between two users (21) and to XXXXXX (25).
-The same comparison for the order-based triad is harder to interpret, since the outstanding neighbourhoods in the Game of Thrones correspond to XXXX (25) and ...
-(27), suggesting that this method fails to capture some relevant structures.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{structure_based_census}
-\includegraphics[width=0.5\textwidth]{order_based_census}
-\includegraphics[width=0.5\textwidth]{time_based_census}
-\caption{Comparison of neighbourhood census obtained by each type of neighbourhood in two different forums.
-Order-based is more similar across forums since its most frequent neighbourhood dominates the census.}
-\label{fig:census_compare}
-\end{figure}
-
-We conclude that, even if at the cost of a sparser set of structures, the time-based neighbourhood does a better job in capturing structures that have both a real meaning and a level of frequency that makes them relevant in the analysis.
+It is clear from the last sections that the order-based does not do a good job on capturing all the richness of the conversational structures.
+There are two other reasons to choose the time-neighbourhood over the structure based.
+First, the time-based has less spurious structures, covering 95\% of posts with 155 neighbourhoods, against the 623 neighbours necessary to cover that amount with the structure-based.
+Second, the time-based neighbourhood has more grounded criteria to decide where the neighbourhood ends, while the structure-based has no criteria at all.
 
@@
 Our dataset contains a set of 100 \textit{active users} who wrote more than 100 posts.
-For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the census, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$.
+For those users, we create an initial feature matrix $U\times N$ where $U$ is the number of users and $N$ is the number of neighbourhoods in the dictionary, and where the the position $(u,n)$ is a counter of the number of times that a post written by user $u$ has a neighbourhood isomorphic to $n$.
 We drop those feature columns that are zero for every user (these are neighbourhoods seen only around posts of users with low-level activity).
@@
 
-% TODO: paragraph discussing a visual/manual analysis of some users
-
 %Interpretation
 
-% Maybe mention that we can take for instance only the posts before the ego and use it to understand when when can recommend a post based on the structure.
+\section{Conclusions}
+The goal of is paper was to shed some light on what would be a proper method to characterise the discussion in which a post is embedded.
+For that, we proposed three different kinds of neighbourhood applicable to tree structures by which conversation threads are represented.
+We named this three types of neighbourhood structure-based, order-based and time-based.
+While the first one is a mere extension of the classical concept of neighbourhood in a graph, the aim of the other two is to explicitly consider time when deciding which are the neighbours of a given post.
 
-\section{Conclusions}
-We presented a method to characterise conversations of users in online threads.
+We showed the limitations of the order-based to capture the structure of conversations.
+We think this finding is relevant because, in some contexts, order is more relevant than real time, and assuming that posts (or whatever vertices represent) arrive in homogeneous intervals gives excellent results \cite{Gomez2012}.
+
+We also showed that the main contribution of the time-based neighbourhood with respect to the most basic structure-based is that, thanks to the use of changepoints to set limits of the neighbourhoods, the space of neighbourhoods is considerably reduced and with less occasional neighbourhoods that occur only once.
+Besides, it prevents neighbourhoods from being too big, even though some post-pruning is still necessary.
+
+We analysed the time .......
+
+ the  it is common in the literature to use the order rather than the real time. The
+
+We introduced three different  a method to characterise conversations of users in online threads.
 Due to the tree nature of online threads, traditional patters such as triads are not able to capture much of relevant dynamics of a conversation.
@@
 \bibliography{bibliography}
-\newpage
-\section*{Appendix  (WILL ME REMOVED)}
-
-\begin{figure*}
-\centering
-%\includegraphics[width=0.5\textwidth]{time_neighbourhood_census_compare_forums}%
-%\includegraphics[width=0.5\textwidth]{order_neighbourhood_census_compare_forums}
-\includegraphics[width=0.5\textwidth]{neighbourhood_census_compare_gameofthrones}%
-\includegraphics[width=0.5\textwidth]{neighbourhood_census_compare_podemos}
-\caption{Comparison of census for two different forums. POUR BERTRAND. WE CANT SEE ANYTHING HERE}
-\label{fig:census_compare}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-%	\includegraphics[width=0.8\textwidth]{census_orderbased_1}
-%	\includegraphics[width=0.8\textwidth]{census_orderbased_2}
-	%\includegraphics[width=0.75\textwidth]{census_orderbased_3}
-	\caption{Most frequent order-based neighbourhoods with $r=2$ and $n=4$ (Podemos)}
-	\label{fig:census_orderbased}
-\end{figure*}
-\begin{figure*}
-	\centering
-%	\includegraphics[width=0.8\textwidth]{census_timebased_1}
-%	\includegraphics[width=0.8\textwidth]{census_timebased_2}
-	%\includegraphics[width=0.8\textwidth]{neighbourhoods_time_3}
-	\caption{Most frequent time-based neighbourhoods (Podemos)}
-	\label{fig:neighbourhoods_time}
-\end{figure*}
-
-\begin{figure*}
-\centering
-%\includegraphics[width=1\textwidth]{confusion_time_order_motifs_gameofthrones_1}
-\caption{Most popular time-based neighbourhoods (up) and how they are mostly detected in order-based neighbourhood (down).}
-\label{fig:census_size_vs_freq}
-\end{figure*}
-
-\begin{figure*}
-	\centering
-	\includegraphics[width=0.9\textwidth]{PCA_cluster_orderbased}
-	\includegraphics[width=0.9\textwidth]{PCA_cluster_timebased}
-	\caption{PCA projections of the order-based (above) and time-based (below) neighbourhood features}
-	\label{fig:PCA}
-\end{figure*}
 

### 2016-04-29 pipeline_clustering.r

@@ -16,15 +16,15 @@ MIN_POSTS <- 100 # number of post to consider a user as active
 # Load data
 ###################################################
 load('./R_objects/dfposts_podemos.Rda')
+df.posts <- df.posts
 df.posts$date <- as.numeric(df.posts$date)
 df.posts <- data.frame(df.posts) %>% arrange(date)
-#df.posts <- df.posts[1:75000,] # Paper
-df.posts <- df.posts[17100:17109,] # Debug
-
+df.posts <- df.posts[1:75000,] # Paper
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
 names(df.threads)[2] <- "length"
 names(df.users)[2] <- "posts"
+df.posts$forum <- 'podemos'
 
 # Print dates range
 start.date <- as.POSIXct(min(as.numeric(df.posts$date)), origin = "1970-01-01") 
@@ -36,34 +36,10 @@ cat('Number of threads: ', nrow(df.threads))
 cat('Number of users: ', nrow(df.users))
 cat('Number of active users', nrow(filter(df.users, posts>MIN_POSTS)))
 
-
-# Plot general overview of forum
-n <- hist(df.threads$length, breaks=0:max(df.threads$length))$counts
-plot(1:max(df.threads$length), n, log='xy', 
-     ylab='Number of threads', xlab='Length')
-title(main='Threads length')
-
-plot(cumsum(table(df.threads$length)), pch=19, cex=0.5,
-     ylab='Length (cum)', xlab='Thread')
-title('Threads length (cumulative)')
-
-n <- hist(df.users$posts, breaks=0:max(df.users$posts))$counts
-plot(1:max(df.users$posts), n, log='xy', 
-     ylab='Number of users', xlab='Posts')
-title(main='Users posts')
-
-plot(cumsum(table(df.users$posts)), pch=19, cex=0.5,
-     ylab='Posts (cum)', xlab='User')
-title('Users posts (cumulative)')
-
-
 #########################################
 # Load neighbourhoods previously extracted
 #########################################
-load("res_time_75000.Rda")
-
-df.posts.bck <- df.posts
-#df.posts <-df.posts.bck
+load("./R_objects/res_time_75000_podemos.Rda")
 df.post.motif  <- res$posts.motifs
 motifs <- res$motifs
 
@@ -73,6 +49,12 @@ all(n == cummin(n))
 
 # Add motif info to posts dataframe
[…]

### 2016-04-29 pipeline_compare_census.r

@@ -538,94 +538,11 @@ axis(2)
 legend(length(motifs)-10,5800, c('podemos', 'gameofthrones'), col=c('blue', 'red'), pch=c(18,20))
 title('Time-based')
 
-############################################################################
-############################################################################
-############################################################################
-#############################################################
-# Get first motifs of first dataset as reference and plot 
-# and compare occurences between the two à la Adamic 
-# Podemos vs Game of Thrones
-############################################################
-library(igraph)
-
-load('./R_objects/dfposts_podemos.Rda')
-df.posts$date <- as.numeric(df.posts$date)
-df.posts <- data.frame(df.posts) %>% arrange(date)
-df.posts <- df.posts[1:75000,] # Paper
-
-df.threads <- plyr::count(df.posts, "thread")
-df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
-names(df.threads)[2] <- "length"
-names(df.users)[2] <- "posts"
-
-#load("./R_objects/res_time_75000_podemos.Rda")
-load("./R_objects/res_order_2_4_75000_podemos.Rda")
-
-# Add motif info to posts dataframe and sort by frequency
-df.post.motif  <- res$posts.motifs
-motifs <- res$motifs
-df.posts <- merge(df.posts, df.post.motif, all=FALSE, sort=FALSE)
-idx <- order(tabulate(df.posts$motif), decreasing = TRUE)
-df.posts$motif <- match(df.posts$motif, idx)
-motifs <- motifs[idx]
-
-# unique name
-df.posts.1 <- df.posts
-motifs.1 <- motifs
-
-#################
-load('./R_objects/dfposts_gameofthrones.Rda')
-df.posts$date <- as.numeric(df.posts$date)
-df.posts <- data.frame(df.posts) %>% arrange(date)
-df.posts <- df.posts[1:75000,] # Paper
-
-df.threads <- plyr::count(df.posts, "thread")
-df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
-names(df.threads)[2] <- "length"
-names(df.users)[2] <- "posts"
-
-#load("./R_objects/res_time_75000_gameofthrones.Rda")
-load("./R_objects/res_order_2_4_75000_gameofthrones.Rda")
-
[…]

### 2016-04-30 doc/IEEEtran/paper_IEEEformat.tex
backup

@@
 \IEEEauthorblockN{Bertrand Jouve}
-\IEEEauthorblockA{FRAMESPA/IMT\\Université de Toulouse 2, France\\
+\IEEEauthorblockA{FRAMESPA/IMT\\Université de Toulouse, France\\
 jouve@univ-tlse2.fr}
@@
 We close the paper with some suggestions of future research.
-%\begin{figure}
-%	\centering
-	%\includegraphics[width=0.3\textwidth]{tree1}
-%	\includegraphics[width=0.3\textwidth]{tree2}
-	%\includegraphics[width=0.3\textwidth]{tree3}
-%	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
-%	\label{fig:trees}
-%\end{figure}
 
@@
 The \textit{root} of the tree is the only vertex with no parent, and corresponds to the post that starts the discussion.
-We say that two vertices $v_i,v_j$ are \textit{siblings} if and only if $p(v_i)=p(v_j)$.
-A vertex $v_i$ is an ancestor of a vertex $v_j$ and $v_j$ is descendant of $v_i$ if and only if $v_i$ is in the path from $v_j$ to the root.
+We say that two vertices $v_i,v_k$ are \textit{siblings} if and only if $p(v_i)=p(v_k)$.
+A vertex $v_l$ is an ancestor of a vertex $v_i$ and $v_i$ is descendant of $v_l$ if and only if $v_l$ is in the path from $v_i$ to the root.
 A \textit{leaf} is a post with no replies.
 A \textit{branch} is the path between a leaf and the root.
-Two vertices $v_i$ and $v_i$ are said to be \textit{neighbours} if either $p(v_i)=v_j$ or $p(v_j)=v_i$ or, in other words, if their distance in the undirected version of $G$ is one.
+Two vertices $v_i$ and $v_m$ are said to be \textit{neighbours} if either $p(v_i)=v_m$ or $p(v_m)=v_i$.
+The neighbours of a vertex $i$ are its parent and its children.
 
@@
 \section{Structural neighbourhoods}\label{sec:structural}
-Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, we start by the following definition of \textit{structural neighbourhood}:
+Extending the classic definition of neighbourhood, we recall the definition of \textit{structural neighbourhood}\footnote{We coined the term \textit{structural} to differentiate if from the other neighbourhoods discussed in this paper.}:
 
@@
 In the context of discussion threads, this definition has two limitations.
-First, the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time when they were written.
-In conversations, time plays an important role, therefore this is an important limitation.
+First, the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time when they were written although, in conversations, time plays an important role.
 Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite.
-This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically similar.
+This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically similar (e.g.: we might not want to put in different categories two structures representing, respectively, a post with 50 replies and another with 40).
 
 \section{Structural-temporal neighbourhoods}\label{sec:temporal}
-We propose two refinements of the structure-based neighbourhood to take into account the order and time in which posts are written.
-As we will see, the criteria to choose which posts are neighbours of a post $i$  is not only based on the structural distance but also on the temporal distance to $i$.
+We propose two refinements of the structure-based neighbourhood to take into account the order and time at which posts are written.
 In the order-based definition, a new radius is set to include at most the $n$ closest posts in time.
@@
 \begin{definition}
-Given a ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ and order $n$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed by the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.
+Given an ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ and order $n$ of vertex $i$, denoted as $\mathcal{N}_{i}^O(r,n)$ is the induced subgraph from its structural neighbourhood $\mathcal{N}_{i}(r)$ composed by the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^O(r,n)$.
 \end{definition}
@@
 If sequences are timestamps, they are monotonic increasing, therefore the changepoints will correspond to sudden pauses in the conversation.
-If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$.
-For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by the authors themselves is available in the R library \texttt{changepoint}.}.
+In the following, we will differentiate \textit{horizontal changepoints}, that arise between siblings, from \textit{vertical changepoints}, that arise within a branch.
+We say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$.
+For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by the authors themselves is available in the R library \texttt{changepoint}.
+We use their \texttt{cpt.meanvar} function.}.
 Now we can introduce our second definition of neighbourhood.
 \begin{definition}
-Given a tree graph $G$, the \textit{time-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
+Given a tree graph $G$, the \textit{time-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic as $i$.
 \end{definition}
@@
 Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the above definition.
-Since the breakpoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical breakpoints in the tree.
-In the case of multiple branches with some common posts, we consider that a common post is a vertical breakpoint if it is a vertical breakpoint in any of the branches.
-Once we have the breakpoints, we can proceed with the algorithm.
+Since the changepoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical changepoints in the tree.
+In the case of multiple branches with some common posts, we consider that a common post is a vertical changepoint if it is a vertical changepoint in any of the branches.
+Once we have the changepoints, we can proceed with the algorithm.
 First, we extract the structural-neighbourhood.
 The time-based neighbourhood will be a subset of the later.
-Then, we look for horizontal and vertical breakpoints, which mark the frontiers of the time-based neighbourhood.
+Then, we look for horizontal and vertical changepoints, which mark the frontiers of the time-based neighbourhood.
 There are four possible cases:
@@
 \begin{itemize}
-\item A \textit{vertical breakpoint in the ancestors}: If the breakpoint is in the path between the post and the root, then the breakpoint started the new local dynamic to which the ego post belongs.
-Thus, we remove the ancestors of the breakpoint, but not the breakpoint itself.
-\item A \textit{vertical breakpoint in the descendants}: If the breakpoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the breakpoint and the breakpoint itself.
-\item A \textit{horizontal breakpoint either in the older siblings or in the ancestors}: If the breakpoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs.
-Similarly, if the breakpoint is among the ancestors, then every older sibling of the breakpoint belongs to a previous local dynamic.
-In both cases, we remove the older siblings of the breakpoint but not the breakpoint itself.
-\item A \textit{horizontal breakpoint elsewhere}: In any other case, the horizontal breakpoint starts a different local dynamic and therefore we remove its younger siblings and the breakpoint. 
+\item A \textit{vertical changepoint in the ancestors}: If the changepoint is in the path between the post and the root, then the changepoint started the new local dynamic to which the ego post belongs.
+Thus, we remove the ancestors of the changepoint, but not the changepoint itself.
+\item A \textit{vertical changepoint in the descendants}: If the changepoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the changepoint and the changepoint itself.
+\item A \textit{horizontal changepoint either in the older siblings or in the ancestors}: If the changepoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs.
+Similarly, if the changepoint is among the ancestors, then every older sibling of the changepoint belongs to a previous local dynamic.
+In both cases, we remove the older siblings of the changepoint but not the changepoint itself.
+\item A \textit{horizontal changepoint elsewhere}: In any other case, the horizontal changepoint starts a different local dynamic and therefore we remove its younger siblings and the changepoint. 
 \end{itemize}
 Any time a vertex is removed we also remove its descendants.
-The final neighbourhood will be connected induced subgraph from the structural neighbourhood.
-If there are no changepoints in the structural neighbourhood, the time-based neighbourhood will be exactly the structural neighbourhood.
+$\mathcal{N}_i^T(r)$ is a connected induced subgraph of $\mathcal{N}_i(r)$.
+If there are no changepoints in the structural neighbourhood, then $\mathcal{N}_i^T(r) = \mathcal{N}_i(r)$.
 
@@
 \begin{algorithmic}
-\REQUIRE Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$
-\ENSURE Subgraph of $g$ with all vertices in $V(g)$
+\REQUIRE Posts tree $G$, vertical changepoints, horizontal changepoints, ego post $i$, radius $r$
+\ENSURE $\mathcal{N}_i^{T}(r)$: Time-based neighbourhood of $i$ at radius $r$
 \STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
-\STATE ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$
-\STATE older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$
+\STATE ancestors $\leftarrow$  ancestors($i$) in $\mathcal{N}_i(r)$
+\STATE older\_siblings $\leftarrow$ older\_siblings($i$) in $\mathcal{N}_i(r)$
 \STATE dump $\leftarrow \varnothing$
-\FOR{bp $\in$ vertical breakpoints}
+\FOR{bp $\in$ vertical changepoints}
  \IF{bp $\in$ ancestors}
@@
 \ENDFOR
-\FOR{bp $\in$ horizontal breakpoints}
+\FOR{bp $\in$ horizontal changepoints}
   \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
@@
 \ENDFOR
-\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts $\cup$ descendants(dump\_posts)) from $\mathcal{N}_i(r)$
+\STATE $\mathcal{N}_i^{T}(r) \leftarrow$ delete(dump\_posts $\cup$ descendants(dump\_posts)) from $\mathcal{N}_i(r)$
 \end{algorithmic}
@@
 \item \textit{Yellow}: parent of ego post and other posts written by the same author.
-It allows to identify, for instance, debates between the ego author (red and orange) and the other author (yellow).
+It allows to identify, for instance, debates between the ego author (red and orange) and this other author (yellow).
 We have observed this phenomena in our data, and it is often the cause of long chains.
@@
 In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.
+Moreover, this choice encourage smaller structures, which allows the computation of isomorphisms (Section~\ref{sec:extraction}) even in real time.
 
 \section{Application to Reddit forums}
+
 Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}.
@@
 
-\subsection{Neighbourhood extraction}
-As a previous step for all further analysis, we extract the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums.
-We label every structure with a unique integer number, where a structure is a coloured graph.
-Two neighbours have the same structure, and therefore the same integer number, if their coloured graphs are isomorphic.
-
-Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of these trees would not be seen in the data.
-It appears then a much more practical approach to enumerate only those trees that are seen in the data.
-In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings), we re-label the structures after the extraction for every post is finished.
-We will call this final labelling a \textit{dictionary} of neighbourhoods.
+\subsection{Neighbourhood extraction}\label{sec:extraction}
+As a previous step for all further analysis, we extracted the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums.
 
@@
 
+%We identified every type of neighbourhood with a unique integer (label). Two neighbourhoods have the same label if their coloured graphs are isomorphic.
+
+%In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings), we re-label the structures after the extraction for every post is finished.
+We will call this final labelling a \textit{dictionary} of neighbourhoods.
+
+The set of all neighbourhoods $\mathcal{N}_i(r)$, $\mathcal{N}_i^O(r,n)$ and  $\mathcal{N}_i^T(r)$ for the two forums constitutes our \textit{dictionary}.
+
 Once the six extractions are finished, we merge the dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth).
 Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.
-
 % Computational cost
@@
    \tabcolsep = 0.8\tabcolsep
-   \begin{tabular}{|cc|cc|cc|}
+   \begin{tabular}{|rr|rr|rr|}
       \hline
@@
    \hline \hline
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 1 (12.5) &  1 (12.5) & 3 (12.5)\\    
-     1 (12.5)& 3(12.5)   &  2 (12.5) & 2 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\  
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
-     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\          
+     7 (9.6)	& 7 (7.4)   		&  		1 (25.2) & 1 (25.2) &  4 (9.0) & 4 (6.4)\\    
+     3 (16.5)& 9 (12.6)   	&  	3 (32.1) & 3 (32.1) &  8 (17.2) & 8 (12.2)\\ 
+     2 (23.1)& 3 (17.8)   	&  	2 (38.8) & 2 (38.8) &  2 (25.3) & 2 (17.9)\\ 
+     20 (27.8)& 11 (22.9)   &  	13 (45.3) & 13 (45.3) &  3 (31.2) & 9 (23.2)\\    
+     4 (31.1)& 5 (27.8)   	&  	4 (50.1) & 4 (50.1) &  10 (36.6) & 14(28.4)\\ 
+     13 (34.1)& 2 (32.3)   	&  	18 (54.7) & 18 (54.7) &  14 (41.8) & 11(33.5)\\ 
+     6 (36.9)& 12 (36.1)   	&  	6 (58.9) & 6 (58.9) &  6 (45.8) & 10(38.5)\\    
+     5 (39.4)& 27 (38.8)   	&	  	12 (65.5) & 10 (62.8) &  15 (48.9) & 5 (43.1)\\ 
+     15 (41.9)& 4 (41.4)   	&  	16 (68.5) & 12 (65.5) &  19 (51.5) & 6 (47.4)\\  
+     12 (43.7)& 6 (43.9)   	&  	5 (70.6) & 16 (68.1) &  5 (53.9) & 3 (51.3)\\    
+     21 (45.3)& 21 (45.9)   	&  	17 (72.8) & 5 (70.7) &  7 (55.9) & 25(55.1)\\ 
+     30 (46.9)& 15 (47.7)   	&  	23 (75.5) & 17 (72.8)&  16 (57.6) & 15(57.8)\\ 
+     16 (48.4)& 13 (49.2)   	&  	31 (76.1) & 23 (74.5)&  36 (59.4) & 12(60.2)\\    
+     39 (49.8)& 20 (50.6)   	&  	19 (77.5) & 31 (76.1) &  25 (61.2) & 7 (62.2)\\ 
+     44 (51.2)& 30 (51.9)   	&  	22 (79.0) & 19 (77.5) &  21 (62.6) & 21(63.9)\\ 
+     8 (52.4)& 22 (53.1)   	&	  	24 (80.3) & 22 (78.9) &  34 (63.9) & 19(65.5)\\    
+     22 (53.3)& 43 (54.2)   	&  	26 (81.5) & 24 (80.3) &  38 (65.1) & 17(66.7)\\ 
+     51 (54.2)& 8 (55.2)   	&  	29 (82.7)& 26 (81.5) &  23 (66.4) & 36(68.0)\\          
    \hline
@@
  \end{center}
- \caption{Frequence}
+ \caption{Frequence this is an example}
 \end{table}

### 2016-04-30 pipeline_clustering.r

@@ -55,21 +55,6 @@ res <- merge.dictionaries(motifs.global, motifs)
 df.posts$motif <- res$mapping[df.posts$motif]
 motifs <- res$dict
 
-
-# Because the counting is based on threads, there might be some posts that were 
-# recovered in df.post.motif but that were cut in df.posts[1:N]
-# so re-sort again motifs by frequency
-# Sort by frequency (and relabel: 1 for the most frequent and so forth)
-# make NA appear in table so that they are convetred to 0
-idx <- order(tabulate(df.posts$motif), decreasing = TRUE)
-df.posts$motif <- match(df.posts$motif, idx)
-motifs <- motifs[idx]
-
-
-# Check they are STILL sorted by frequency (they should be)
-n <- as.numeric(table(df.posts$motif))
-all(n == cummin(n))
-
 ######################################################
 # Count motifs in which each user appears
 ######################################################

### 2016-04-30 pipeline_compare_census.r

@@ -283,6 +283,8 @@ sum(cumsum(n2) < th)
 #######################################################################
 # Create a common neighbourhood dictionary index
 #######################################################################
+if(FALSE){
+  
 df.posts.global <- df.posts
 motifs.global <- motifs.got.time
 
@@ -313,7 +315,7 @@ motifs.global <- res$dict
 
 #save(df.posts.global, file="df.posts.global.Rda")
 #save(motifs.global, file="motifs.global.Rda")
-
+}
 load("df.posts.global.Rda")
 load("motifs.global.Rda")
 
@@ -394,6 +396,7 @@ plot(df.posts.global$motif.struct, df.posts.global$motif.time, cex=0.1, pch=19,
 #confusion <- m1%*%t(m2) 
 confusion.order_time <- acast(df.posts.global, motif.order~motif.time, fun.aggregate = length) # faster?
 confusion.time_struct <- acast(df.posts.global, motif.time~motif.struct, fun.aggregate = length) # faster?
+
 #save(confusion.order_time, file="confusion.order_time.Rda")
 #save(confusion.time_struct, file="confusion.time_struct.Rda")
 load("confusion.order_time.Rda")
@@ -493,6 +496,11 @@ axis(2)
 legend(length(motifs)-10,5800, c('podemos', 'gameofthrones'), col=c('blue', 'red'), pch=c(18,20))
 title('Structure-based')
 
+# for the table in the paper
+recall.podemos <- cumsum(sapply(top.motifs.podemos, function(x) sum(motifs.podemos==x)/length(motifs.podemos))*100)
+recall.gameofthrones <- cumsum(sapply(top.motifs.gameofthrones, function(x) sum(motifs.gameofthrones==x)/length(motifs.gameofthrones))*100)
+cbind(top.motifs.podemos, recall.podemos)
+cbind(top.motifs.gameofthrones, recall.gameofthrones)
 # order
 ##########
 ntop <- 30
@@ -515,6 +523,11 @@ axis(2)
 legend(length(motifs)-10,14000, c('podemos', 'gameofthrones'), col=c('blue', 'red'), pch=c(18,20))
 title('Order-based')
 
+# for the table in the paper
+recall.podemos <- cumsum(sapply(top.motifs.podemos, function(x) sum(motifs.podemos==x)/length(motifs.podemos))*100)
[…]

### 2016-05-05 R/count_motifs.r
Pruning improved. motif_counts store hash of conversation

@@ -1,58 +1,12 @@
 library(RSQLite)
 library(data.table)
+library(digest)
 #library(R.utils)
 
 source('R/extract_from_db.r')
 source('R/tree_changepoints.r')
+source('R/pruning.r')
 
-
-prune <- function(motif){
-  # Arguments:
-  #   motif: a tree graph
-  # Returns:
-  #   a pruned motif
-  # 4 colors: 1(black) 2(red) 4(orange)
-  # we can only remove black nodes
-  # there can be only two black nodes together
-  # delete 
-  parents <- which(degree(motif, mode='in')>2)
-  to.delete <- c()
-  if(length(parents)==0){
-    return(motif)
-  }
-  # If more than two siblings with no children, keep only the first two
-  # (the most ancient)
-  for (p in 1:length(parents)){
-    siblings <- neighbors(motif, parents[p], mode='in')
-    siblings <- siblings[order(V(motif)$date[siblings])]
-    siblings.indegree <- degree(motif, v=siblings, mode='in')
-    
-    if(length(siblings)<3){
-      next
-    }
-    x <- siblings$color
-    
-    counts <- 1
-    delete <- rep(FALSE, length(x))
-    for(i in 2:length(x)){
-      if(x[i] != x[i-1]){
-        counts <- 1
-      }
-      else if(siblings.indegree[i]>0){
-        counts <- 1
-      }
-      else{
-        counts <- counts + 1
-      }
-      if(counts>2){
-        delete[i] <- TRUE
-      }
-    }
-    to.delete <- union(to.delete, siblings[delete])
-  }
-  motif <- delete.vertices(motif, to.delete)
-  motif
-}
 
 count_motifs_by_post <- function(threads, 
                                  database='reddit', 
@@ -82,13 +36,11 @@ count_motifs_by_post <- function(threads,
     tryCatch({
       
       write.table(threads[i], file="tried_threads.csv", row.names=FALSE, col.names=FALSE, append=TRUE)
-      #cat(threads[i], file=filelog, sep=",", append=TRUE)
-      
+
       #Extract the egos of every post
       g <- database.to.graph(threads[i], con, database)
       gp <- g$gp
-      #cat('graph', file=filelog, sep=",", append=TRUE)
-      
+
       size <- vcount(gp)
       cat('\n', i, '/', nthreads, '/', threads[i], ' size: ', size, ' chunk.id: ', chunk.id )
       if(size>1000){
@@ -102,8 +54,7 @@ count_motifs_by_post <- function(threads,
[…]

### 2016-05-05 R/count_motifs_by_segment.r

@@ -1,0 +1,175 @@
+library(data.table)
+source('R/extract_from_db.r')
+source('R/tree_changepoints.r')
+
+tree.segments <- function(gp, breakpoints.v, breakpoints.h){
+  # Return the segments (subgraphs) of g according to its changepoints
+  cat("\nCOMPUTE SEGMENTS OF TREE\n")
+  segs <- list()
+  processed <- c()
+  cat("breakpoints:", breakpoints.v)
+  for (i in 1:vcount(gp)){
+    # check if already in some segment
+    if (V(gp)[i] %in% processed){ 
+      cat('\n next')
+      next
+    }
+    cat('\nnode: ', i)
+    # detect neighbourhood
+    gs <- neighborhood.temporal(gp, i, 10, breakpoints.v, breakpoints.h)
+    cat('\nsegment: ', as.numeric(V(gp)[V(gs)$name]))
+    
+    # plot it 
+    V(gp)[breakpoints.v]$color <- 'green'
+    V(gp)[breakpoints.h]$color <- 'blue'
+    V(gp)[V(gs)$name]$color <- "red"
+    plot.tree(gp, labels = 'id')
+    V(gp)[V(gs)$name]$color <- "white"
+    
+    segs[[length(segs)+1]] <- gs  
+    # add them to processed list
+    processed <- union(processed, V(gp)[V(gs)$name])
+    #plot.tree.breakpoints(gp, breakpoints.v, breakpoints.h)
+  }
+  segs
+}
+
+count_motifs_by_usersegment <- function(threads, 
+                                 database='reddit', 
+                                 rad=2,
+                                 max.neighbors=4,
+                                 chunk.id=0){
+  
+  con <- dbConnect(dbDriver("SQLite"), dbname = paste0("./data/", database, ".db"))
+  
+  # For every thread, recontruct its graph. Then count 
+  # if neighborhood at distance 1 and binary tree, max nodes is 4
+  # if neighborhood at distance 1 and binary tree, max nodes is 10
+  # 4   10   22   30   46   62   93  123...
+  motifs <- list()
+  posts.id <- vector()
+  posts.motifs.id <- vector()
+  plotted.trees <- 1
+  
+  id.process <- Sys.getpid()
+  filelog <- paste0("log", id.process, ".csv")
+  
+  
+  nthreads <- length(threads)
+  posts.motifs <- data.frame()
+  for(i in 1:nthreads){
+
+      #Extract the egos of every post
[…]

### 2016-05-05 R/plotting.r

@@ -2,20 +2,52 @@
 # Plotting functions
 #
 
-plot.tree <- function(gtree){
-  # Plots a tree graph
-  # Arguments:
-  #   gtree: a igraph object graph with no cycles (tree)
 
+plot.tree.breakpoints <- function(gtree, breakpoints.v, breakpoints.h, labels=c('name', 'id')){
   mypalette <- c("black", "yellow", "orange", "red", "white")
   par(mfrow=c(1,1))
   gtree.un <- as.undirected(gtree)
+  V(gtree.un)$color <- 'white'
+  V(gtree.un)[breakpoints.v]$color <- 'green'
+  V(gtree.un)[breakpoints.h]$color <- 'blue'
+  
+  V(gtree.un)$size <- 3
+  V(gtree.un)[breakpoints.v]$size <- 10
+  V(gtree.un)[breakpoints.h]$size <- 10  
+  
   la = layout_as_tree(gtree.un, mode='out', root=which.min(V(gtree.un)$date))
+  if (labels=='name'){
+    labels <- V(gtree)$name
+  }
+  else{
+    labels <- as.numeric(V(gtree))
+  }
   plot(gtree.un,
        layout = la,
-       vertex.color=mypalette[V(gtree.un)$color],
-       vertex.label = "",
-       vertex.size=1,
+       vertex.label = labels,
+       edge.arrow.size=0.6)
+}
+
+plot.tree <- function(gtree, labels=c('name', 'id')){
+  # Plots a tree graph
+  # Arguments:
+  #   gtree: a igraph object graph with no cycles (tree)
+  if (missing(labels)){
+    labels <- NA
+  }
+  else{
+  labels <- switch(labels, 
+                   'name' = V(gtree)$name, 
+                   'id' = as.numeric(V(gtree)))
+  }
+  par(mfrow=c(1,1))
+  gtree.un <- as.undirected(gtree)
+  la = layout_as_tree(gtree.un, mode='out', root=which.min(V(gtree.un)$date))
+
+  plot(gtree.un,
+       layout = la,
+       vertex.label = labels,
+       vertex.size=3,
        edge.arrow.size=0.6)
 }
 

### 2016-05-05 pipeline_compare_census.r

@@ -143,7 +143,12 @@ load("./R_objects/res_time_75000_podemos.Rda")
 res.pod.time <- res
 
 # Add motif info to posts dataframe and sort by frequency
-df.posts.pod <- merge(df.posts.pod, res.pod.time$posts.motifs, all=FALSE, sort=FALSE)
+df.posts.pod <- merge(df.posts.pod, res.pod.time$posts.motifs, all=FALSE, sort=FALSE) %>%
+                select(user, conversation, motif)
+df.posts.pod <- df.posts.pod %>% distinct
+# check whether some user A in same conversation has two or more different motifs
+test <- group_by(df.posts.pod, user, conversation) %>% summarise(length(unique(motif))) %>% as.data.frame
+
 idx <- order(tabulate(df.posts.pod$motif), decreasing = TRUE) # get order by frequency
 df.posts.pod$motif <- match(df.posts.pod$motif, idx) # re-arrange pointers to motifs
 motifs.pod.time <- res.pod.time$motifs[idx] # re-sort motifs
@@ -152,6 +157,8 @@ colnames(df.posts.pod)[which(names(df.posts.pod) == "motif")] <- "motif.time"
 # Check they are sorted by frequency (they should be)
 n <- as.numeric(table(df.posts.pod$motif.time))
 all(n == cummin(n))
+
+
 
 ##############
 # load("./R_objects/res_struct_75000_4chan.Rda")

### 2016-05-05 pipeline_extract_neighbourhoods.r

@@ -32,11 +32,11 @@ source('R/plotting.r')
 ##########################################################
 # Load Data
 ##########################################################
-load('./R_objects/dfposts_4chan.Rda')
+load('./R_objects/dfposts_podemos.Rda')
 df.posts$date <- as.numeric(df.posts$date)
 df.posts <- data.frame(df.posts) %>% arrange(date)
 df.posts <- df.posts[1:75000,] # Paper
-#df.posts <- df.posts[17100:17109,] # Debug
+#df.posts <- df.posts[1:5000,] # Debug
 
 df.threads <- plyr::count(df.posts, "thread")
 df.users <- plyr::count(df.posts, 'user')                                                                                                                                   
@@ -60,7 +60,7 @@ cat('Number of active users', nrow(filter(df.users, posts>MIN_POSTS)))
 # Only long threads
 #df.threads <- filter(df.threads, length>10)
 
-chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/135))
+chunks <- split(df.threads$thread, ceiling(seq_along(df.threads$thread)/1000))
 length(chunks)
 
 if(FALSE){
@@ -78,14 +78,14 @@ prof <- profvis({
 # sequential
 res.seq <- count_motifs_by_post(as.vector(unlist(chunks)), 
                                 database='reddit',
-                                neighbourhood='struct')
+                                neighbourhood='time')
 }
 
 # parallel
-ncores <- detectCores() - 6
+ncores <- detectCores() - 2
 cl<-makeCluster(ncores, outfile="", port=11439)
 registerDoParallel(cl)
-pck <- c('RSQLite', 'data.table', 'changepoint')
+pck <- c('RSQLite', 'data.table', 'changepoint', 'digest')
 res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   source('R/extract_from_db.r')
   #withTimeout(  count_motifs_by_post(chunks[[i]], 
@@ -94,11 +94,12 @@ res.parallel <- foreach(i=1:length(chunks), .packages = pck)%dopar%{
   #                                  120, onTimeout='warning')
   count_motifs_by_post(chunks[[i]], 
                        database='reddit',
-                       neighbourhood='struct', chunk.id=i)
[…]

### 2016-05-05 doc/2016-05 Luchon /presentation_SIG_Toulouse_Avril_2016.tex

--- 2016-04-10 doc/2016-04 SIG Toulouse/presentation_SIG_Toulouse_Avril_2016.tex
@@
 }
+\AtBeginSubsection[]
+{
+	\begin{frame}
+		\begin{NoHyper}
+		\tableofcontents[currentsection, currentsubsection]
+		\end{NoHyper}
+	\end{frame}
+}
+\AtBeginSubsubsection[]
+{
+	\begin{frame}
+		\begin{NoHyper}
+		\tableofcontents[currentsection, currentsusubbsection]
+		\end{NoHyper}
+	\end{frame}
+}
 \begin{document}
+
 \addtobeamertemplate{footline}{\insertframenumber/\inserttotalframenumber}
@@
 \author{Alberto Lumbreras \\Jouve B., Velcin J., Guégan, M.}
-\date{April 8, 2016}
+\date{May 3, 2016}
 \maketitle
@@
 	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}complexsystems}}\\
+	\href{www.reddit.com/r/gameofthrones}{www.reddit.com/r/{\color{red}gameofthrones}} $\leftarrow$ in this presentation\\
 	\href{www.reddit.com/r/podemos}{www.reddit.com/r/{\color{red}podemos}} $\leftarrow$ in this presentation\\
@@
 
-\begin{frame}{Graph representations}{Which one is better?}
+\begin{frame}{Graph representations}{Conversations are trees}
 
 \begin{itemize}
-\item Depends on the task!
-\item One might choose multiple representations (multi-level analysis)
-\end{itemize}
-\vfill
-My choice: 
-\begin{itemize}
-\item Mostly tree representation
-\item Because it explicitly represents discussions (and their evolution).
+\item Explicit structure.
+\item Dynamic (order, time)
 \end{itemize}
@@
 \end{figure}
-And sometimes:
-\begin{itemize}
-\item SNA representation of single conversations.
-\end{itemize}
 \end{frame}
@@
 \textit{Hypothesis}: different individuals have tendency towards different types of conversations and these types are reflected in the structure of their interactions.
-\vfill
-These conversational structures might be observed at two levels (at least):
-\begin{itemize}
-\item Social graph.
-\item Posts graph (tree)
-\end{itemize}
+
 \end{frame}
@@
 
-\begin{frame}{Order-based neighbourhoods}{Census}
-We found 129 different motifs:
-	\begin{figure}
-		\centering
-		\includegraphics[width=1\textwidth]{neighbourhoods_2_4_1}
-	\end{figure}
-\end{frame}
-
 \begin{frame}{Time-based neighbourhoods}{Definition}
@@
 	\item 1. Extract neighbourhood of post $i$ with radius $r$.
-	\item 2. Detect changes of speed (vertical/horizontal changepoints)
+	\item 2. Detect changes of speed (vertical/horizontal \textit{changepoints}) (PELT algo)
 	\item 3. From $i$, get the posts around until a changepoint is found.
@@
 
-\begin{frame}{Time-based neighbourhoods}{Census}
-We found 165 different motifs:
+\begin{frame}{Colouring}
+Label special nodes:
+\begin{itemize}
+	\item Red: ego.
+	\item Yellow: parent of ego (and posts of same author)
+	\item Orange: other posts by ego author
+	\item White: root
+\end{itemize}
 	\begin{figure}
 		\centering
-		\includegraphics[width=1\textwidth]{neighbourhoods_time_1}
+		\includegraphics[width=0.3\textwidth]{colouring}
 	\end{figure}
+
+\end{frame}
+\begin{frame}{Some frequent neighbourhoods}{}
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{dictionary1}
+	\end{figure}
+\end{frame}
+\begin{frame}{Some frequent neighbourhoods}{}
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{dictionary2}
+	\end{figure}
+\end{frame}
+
+\subsection{Comparing neighbourhoods}
+\begin{frame}{Frequency distribution}{}
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.8\textwidth]{compare_census_distribution}
+	\end{figure}
+Time-based (black) reduces the space w.r.t structure-based (blue)
+\end{frame}
+
+\begin{frame}{Frequency distribution}{}
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{compare_census_distribution_cum}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Discrepancies}{Time-based vs Order-based neighbourhood}
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{confusion_time_order}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Discrepancies}{Basic neighbourhood vs Time-based neighbourhood}
+	\begin{figure}
+		\centering
+		\includegraphics[width=1\textwidth]{confusion_struct_time}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Structure-based vs Order-based vs Time-based}
+Structure-based:
+\begin{itemize}
+\item  too big (and too many) neighbourhoods.
+\end{itemize}
+Order-based:
+\begin{itemize}
+\item  Dominance of monoid hides richer conversational structures.
+\end{itemize}
+Time-based:
+\begin{itemize}
+\item Space more reduced than simple structure-based.
+\item Criteria to choose radius dynamically (r = until conversation slows down)
+\end{itemize}
 \end{frame}
@@
 
-\begin{frame}{Conversation-based clustering}{Order-based}
+\begin{frame}{Conversation-based clustering}{Time-based}
 	\begin{figure}
 		\centering
-		\includegraphics[width=1\textwidth]{PCA_cluster_order_based}
+		\includegraphics[width=1\textwidth]{cluster_whiskers}
+	\end{figure}	
+\end{frame}
+
+\begin{frame}{Conversation-based clustering}{Interpretation}
+	Greens:
+	\begin{figure}
+		\centering
+		\includegraphics[height=0.17\textwidth]{motif8}
+		\includegraphics[height=0.17\textwidth]{motif4}
+		\includegraphics[height=0.17\textwidth]{motif15}
+	\end{figure}	
+	Reds:
+	\begin{figure}
+		\centering
+		\includegraphics[height=0.17\textwidth]{motif20}
+		\includegraphics[height=0.17\textwidth]{motif3}
+		\includegraphics[height=0.17\textwidth]{motif19}	
+		\includegraphics[height=0.17\textwidth]{motif14}	
+		\includegraphics[height=0.17\textwidth]{motif13}	
+		\includegraphics[height=0.17\textwidth]{motif18}	
+	\end{figure}	
+	Blues (avoid these motifs):
+	\begin{figure}
+		\centering
+		\includegraphics[height=0.17\textwidth]{motif3}
+		\includegraphics[height=0.17\textwidth]{motif19}
 	\end{figure}	
@@
 	\end{itemize}
-	\textbf{Future work:}
-	\begin{itemize}
-		\item Prune time-based neighbourhoods to reduce dimensionality.
-		\item Do users jump from cluster to cluster (paths of roles)
-	\end{itemize}	
 \end{frame}

### 2016-05-05 doc/IEEEtran/paper_IEEEformat.tex

@@
 % in the abstract
-\begin{abstract}Users of social networks are often characterised by extracting some relevant features from the graphs associated to that network.
-The structural dynamic of the conversation are usually forgotten due to a lack of proper tools.
-We present a purely graph-based method to cluster users in online forums.
+\begin{abstract}Social network analysis provides tools to analyse the global and the local structure of graphs when these graphs represent users interactions.
+In the context of online conversations, usually represented by trees, there is a lack of tools to analyse the local structure of these conversations.
+In this paper, we introduced two novel definitions of neighbourhood adapted to conversation trees.
+We analyse them in two Reddit forum and, finally, we show that they can successfully detect different types of conversationalists in online forums.
 \end{abstract}
@@
 Another common graph is a tree graph where vertices represent posts (or e-mails) and edges from one vertex to another indicate that the first is a reply to the second.
-Given the different discussion trees of a forum (a forest) the global analyses include depth distribution, branching factors, or even the time between two posts \cite{Bhatt2012}.
+Given the different discussion trees of a forum (a forest) the global analyses include depth distribution, branching factors, or even the time between two posts \cite{Bhatt2012, Gaumont2016,Dorat2007}.
+
 Studying discussion trees from the point of view of the users means, since trees are discussions, studying the user at a conversational level.
-Unfortunately, there is a lack of structural tools to perform this analysis.
+Unfortunately, there is a lack of tools to perform an structural analysis of the conversation trees at a local level.
 In this paper, we propose to fill this gap through the concept of \textit{structural-temporal neighbourhood}.
-%El esquema debe ser;
-%- SNA: 
-%*** comunidad: degree distribution, density, clustering coefficient etc
-%*** usuario: centrality
-%- Tree/forest:
-%*** global properties: degree distribution, depth distribution, inter-message times,...
-%** Individual properties: ...
 
@@
 We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees.
-Then we introduce our two definitions of structural-temporal neighbourhoods.
-To illustrate the kind of neighbourhoods that we obtain in a real data, we analyse the neighbourhood census of a Reddit forum.
-Finally, we apply both neighbourhood definitions to detect clusters of users that tend to appear in the same neighbourhoods.
+Then we introduce our two new definitions of structural-temporal neighbourhoods.
+To illustrate the kind of neighbourhoods that we obtain in a real data, we analyse two Reddit forums.
+Finally, we apply our time-based neighbourhood definition to detect clusters of users that tend to appear in the same neighbourhoods.
 We close the paper with some suggestions of future research.
@@
 
-Given a type of neighbourhood (structure-based, order-based or time-based) we extract the neighbourhood around every post of a forum as follows.
+For each method (structure-based, order-based or time-based) we extract the neighbourhood around every post of a forum as follows.
 A new label is assigned to every neighbourhood structure the first time it is detected.
@@
 Otherwise we create a new label for the current neighbourhood.
+The set of neighbourhoods found and their labels constitutes a \textit{dictionary}.
 
-%We identified every type of neighbourhood with a unique integer (label). Two neighbourhoods have the same label if their coloured graphs are isomorphic.
+Once the six extractions are finished, we merge the dictionaries into a global one and re-label the neighbourhoods by frequency (1 for the most frequent and so forth).
+Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.
 
-%In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings), we re-label the structures after the extraction for every post is finished.
-We will call this final labelling a \textit{dictionary} of neighbourhoods.
-
-The set of all neighbourhoods $\mathcal{N}_i(r)$, $\mathcal{N}_i^O(r,n)$ and  $\mathcal{N}_i^T(r)$ for the two forums constitutes our \textit{dictionary}.
-
-Once the six extractions are finished, we merge the dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth).
-Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.
 % Computational cost
@@
 \subsection{Comparing neighbourhood methods}
-In this section we compare different aspects of our three classes of neighbourhoods, namely (a) the size and frequency distribution of the neighbours for each class, (b) the discrepancies between methods when they extract the neighbourhood of a same post, and (c) the neighbourhood census obtained for each class in the same forum.
+In this section we compare different aspects of the three methods, namely (a) the size and frequency distribution of the neighbours for each method, (b) the discrepancies between methods when they extract the neighbourhood of a same post.
 
 \subsubsection{Size and frequency distribution}
-We detected in the the whole dataset 3,791 different structure-based neighbourhoods, 174 different order-based and 1,488 different time-based, which means that the structure-based captures a much wider set of structures.
+We detected in the whole dataset 3,791 different structure-based neighbourhoods, 174 different order-based and 1,488 different time-based, which means that the structure-based captures a much wider set of structures.
 Of course, this is not necessarily an advantage since these extra structures may be just spurious and only occur a very small number of times.
@@
 This is the main handicap of the order-based neighbourhood since, as we will see later, it avoids finding richer structures.
+Table~{\ref{tab:table}} shows the most frequent neighbourhoods for the three methods in each forum.
+The results are quite sensitive to the choice of the neighbourhood.
+Note that the order-based detects less differences among the most frequent neighbourhoods in both forums.
+Besides, it seems that when the time-based detects the neighbourhood 4, the structure-based detects the neighbourhood 7.
+This illustrate the effect of the changepoint: the presence of an horizontal changepoint in 7 (many posts replying to the root) converts it, from the point of view of the time-based neighbourhood, to a 4, which indicates that the ego has at most two siblings in the same period of time.
 
@@
 
+%\begin{figure}
 \begin{table}
@@
  \end{center}
- \caption{Frequence this is an example}
+ \caption{Most frequent neighbourhoods (and cummulative frequence) for the three methods if Podemos (POD) and Game of Thrones (GOT) forums.}
 \end{table}
+%\end{figure}
 
 \subsubsection{Discrepancies}
-The divergent number of neighbourhoods between structure-based, order-based and times-based suggests that some neighbourhoods, for instance, time-based, might be systematically mapped to the same neighbourhoods in order-based.
+The divergent number of non-isomorphic neighbourhoods between structure-based, order-based and times-based methods suggests that some neighbourhoods, for instance, time-based, might be systematically mapped to the same neighbourhoods in order-based.
 Or similarly, that some structure-based neighbourhoods are systematically mapped into the same time-based neighbourhoods.
@@
 In general, the structures with a bigger agreement are among the most frequent.
-This is comprehensible since the most frequent tend to be also the most simple structures with less room for discrepancy.
+This is probably due to the fact that a high amount of threads are small and therefore there is less room for discrepancy.
 There is a 30.75\% of the posts that are assigned the same neighbourhood structure by the three methods.
@@
 
-It is clear from the last sections that the order-based does not do a good job on capturing all the richness of the conversational structures.
+It is clear from the this analysis that the order-based does not do a good job on capturing all the richness of the conversational structures due to the dominance of neighbourhood 1.
 There are two other reasons to choose the time-neighbourhood over the structure based.
@@
 \subsection{Conversation-based clustering of users}
-In this section, we cluster users in the Podemos dataset based on the neighbourhoods of their posts.
+In this section, we apply the time-based neighbourhoods to cluster users in the Podemos dataset based on the neighbourhoods of their posts.
 Our intuition is that some users like participating in some kind of discussion rather than other.
@@
 Since the set of feature dimension is relatively large, we show in Figure \ref{fig:whiskers} how the users in a cluster are distributed for each feature.
-The features shown are the ones that have an bigger absolute value (in mean) for each cluster.
+The features shown are the ones that have an bigger absolute value (in mean) for each cluster (above the percentile 0.95).
 
@@
 \begin{itemize}
-\item The green cluster (attraction for X, X, repulsion XXX) correspond to users that...
-\item The red cluster...
-\item The blue cluster...
+\item Green cluster: these users are prominent  in conversations 8, 4, 15, which correspond to replies to the root in threads where the root post attracted some attention.
+
+\item Red cluster: these users tend to appear in 20, 3, 19, 14, 13, and 18, which are almost all cascades, sometimes between only two users (20, 18).
+
+\item Blue cluster: these users seems to avoid cascades (3, 19) while they do not limit themselves to reply to the root, but instead the participate deeper in the discussion (6).
 \end{itemize}
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{whiskers}
+	\includegraphics[width=0.5\textwidth]{cluster_whiskers}
 	\caption{Whiskers plots in the most relevant features (time-based)}
@@
 
-% TODO: Paragraph to explain similarity between clusters assignments
-
-% TODO: Paragraph to explain which motifs are prefered for every cluster
-
-%Interpretation
-
 \section{Conclusions}
-The goal of is paper was to shed some light on what would be a proper method to characterise the discussion in which a post is embedded.
-For that, we proposed three different kinds of neighbourhood applicable to tree structures by which conversation threads are represented.
+The goal of is paper was to shed some light on what would be a proper method to characterise discussions online forums.
+For that, we analysed three different kinds of neighbourhood applicable to tree structures by which conversation threads are represented.
 We named this three types of neighbourhood structure-based, order-based and time-based.
@@
 
-We also showed that the main contribution of the time-based neighbourhood with respect to the most basic structure-based is that, thanks to the use of changepoints to set limits of the neighbourhoods, the space of neighbourhoods is considerably reduced and with less occasional neighbourhoods that occur only once.
-Besides, it prevents neighbourhoods from being too big, even though some post-pruning is still necessary.
+In the context of online discussions, time is critical and must be considered in the definition of the neighbourhood motifs.
+Our choice has been to use \textit{changepoints}, variations in the pace of the discussion, to decide where a neighbourhood ends.
+Besides explicitly considering time, this method gives in practice relatively small neighbourhoods, which guarantees its scalability since isomorphisms are considerably fast to compute in small graphs.
 
-We analysed the time .......
+We also showed that, thanks to the use of changepoints to set limits of the neighbourhoods, the space of neighbourhoods is considerably reduced with respect to the structural neighbourhood.
 
- the  it is common in the literature to use the order rather than the real time. The
-
-We introduced three different  a method to characterise conversations of users in online threads.
-Due to the tree nature of online threads, traditional patters such as triads are not able to capture much of relevant dynamics of a conversation.
-Our defined order-based and temporal-based neighbourhoods
-are able to capture a very rich variety of structures.
-We used this neighbourhoods to characterise users in terms of the structure of the conversations they participate in and showed that, indeed, there are different types of structural conversationalists
+Finally, we used this neighbourhoods to characterise users in terms of the structure of the conversations they participate in and showed that, indeed, there are different types of structural conversationalists.
 
@@
 The concept of structural-temporal neighbourhood opens the door to some interesting paths of research.
-One might wonder whether other pruning are more pertinent than the proposed here.
-Also, even after pruning some neighbourhoods suggest the same type of conversation, so a manual merge might be convenient.
-Maybe sociologists can help to merge the found neighbourhoods in a meaningful way.
-
-\section{Future work}
-In this paper we focused on the comparative analysis between three kind of neighbourhoods in order to demonstrate that the time-based neighbourhoods are able to capture relevant structures that other definitions of neighbourhood cannot.
-We also showed that it can be used to characterise users according to the kind of conversation in which they participate.
-
-We can see many applications of this kind of neighbourhood.
-
-\begin{itemize}
-\item  Caracterise reactions to user posts: instead of capturing the neighbours all around a post, we can limit we neighbourhood to the descendants.
-This would allow to cluster users from a slightly (but maybe important) point of view.
-\item The boundaries of our time-based neighbourhood are decided by the changepoint algorithm PELT. How sensible is time-based neighbourhood to the choice of the algorithm?  
-\item Merge more neighbourhoods with sociological criteria.
-\item Use different coulours to represent different phenomena.
-\end{itemize}
+The space of neighbourhoods is still quite large and, besides using pruning strategies more aggressive, it would be interesting to analyse which neighbourhoods, despite of being non-isomorphic, may represent the same type of conversation.
+Such analyse would allow to merge some of the neighbourhoods resulting in an even more reduced and meaningful space.
+Also, other kind of neighbourhoods can also be studied to analyse different phenomena.
+For instance, we might take only those neighbours which are descendants of the ego post.
+Such kind of neighbourhood may be used to classify users in terms in the reaction they trigger in the discussion.
+The inverse technique, that is, take all the neighbours that are not descendants of the ego, may help to understand the kind of discussion that a users tends to be attracted to.
+This might be useful for posts recommendations or even to improve current generative models that try to reproduce the way how a thread grows \cite{Gomez2012, Kumar2010}.
 
